### 2014-07-04 papers/termination/table.tex
Reran benchmarks

loop1.c & 6 &  & true & true & false & false & 1 & 5 & term & 2.4s & 1\\

loop2.c & 7 &  & true & true & false & false & 1 & 5 & term & 2.5s & 3\\

loop3.c & 8 & true & true & true & true & false & 1 & 5 & term & 2.8s & 3\\

loop4.c & 13 &  & true & true & false & false & 2 & 5 & term & 3.1s & 17\\

loop5.c & 13 &  & true & true & false & false & 2 & 5 & term & 2.9s & 8\\

loop6.c & 6 &  & false & true & false & false & 1 & 5 & term & 2.4s & 1\\

loop7.c & 7 &  & true & unk & false & false & unk & 5 & term & 18.2s & 10\\

loop8.c & 6 &  & true & true & true & true & 1 & 5 & term & 2.4s & 1\\

loop9.c & 11 & true & true & false & false & false & 1 & 5 & term & 3.6s & 3\\

loop10.c & 6 &  & true & true & false & false & 1 & 5 & term & 2.5s & 1\\

loop43.c & 9 &  & true & false & false & false & ? & 5 & term & 3.7s & 4\\

loop44.c & 9 & false & true & false & false & false & ? & 5 & nonterm & 2.6s & 4\\

loop45.c & 15 &  & true & false & false & false & ? & 5 & term & 3.4s & 19\\

loop46.c & 9 &  & true & false & false & false & ? & 5 & term & 2.7s & 4\\

loop47.c & 6 &  & true & false & false & false & ? & 5 & term & 3.6s & 4\\

loop48.c & 7 &  & true & false & false & false & ? & 5 & term & 2.5s & 3\\

loop49.c & 6 &  &  &  &  &  &  & 5 & ? & T/O & 12\\

svcomp1.c & 17 &  & true & true & false & false & 1 & 5 & term & 2.8s & 5\\

svcomp2.c & 13 &  & false & true & false & false & 1 & 5 & term & 2.6s & 6\\

svcomp3.c & 15 &  & true & true & false & false & unk & 5 & term & 36.7s & 36\\

svcomp4.c & 14 & false & true & unk & true & false & unk & 5 & nonterm & 50.5s & 75\\

svcomp5.c & 18 &  & true & false & true & false & unk & 5 & ? & T/O & 46\\

svcomp6.c & 11 &  & true & true & false & false & 1 & 5 & ? & T/O & 16\\

svcomp7.c & 9 & true & true & true & true & false & 1 & 5 & term & 9.7s & 7\\

svcomp8.c & 9 &  & true & unk & false & false & unk & 5 & ? & T/O & 38\\

svcomp9.c & 12 &  & true & unk & false & false & unk & 5 & term & 2.8s & 11\\

svcomp10.c & 15 &  & true & true & false & false & 1 & 5 & term & 3.4s & 19\\

svcomp11.c & 16 &  & true & true & false & false & 1 & 5 & term & 5.5s & 27\\

svcomp12.c & 13 &  & true & true & false & false & 1 & 5 & term & 5.9s & 28\\

svcomp13.c & 11 &  & true & true & false & false & 1 & 5 & ? & T/O & 30\\

svcomp14.c & 18 &  & true & true & false & false & 1 & 5 & ? & T/O & 66\\

svcomp15.c & 30 &  & true & true & false & false & 1 & 5 & ? & T/O & 28\\

svcomp16.c & 17 &  & true & true & false & false & 1 & 5 & term & 2.7s & 3\\

svcomp17.c & 18 &  & true & true & false & false & 1 & 5 & ? & T/O & 45\\

svcomp18.c & 19 &  & true & true & false & false & 1 & 5 & ? & T/O & 27\\

svcomp25.c & 12 &  & true & true & false & false & 1 & 5 & ? & T/O & 15\\

svcomp26.c & 23 &  & true & true & false & false & 1 & 5 & term & 2.9s & 11\\

svcomp27.c & 13 & false & true & true & false & false & 1 & 5 & ? & T/O & 76\\

svcomp28.c & 18 &  & true & true & false & false & 1 & 5 & ? & T/O & 82\\

svcomp29.c & 9 &  & true & true & false & false & 1 & 5 & ? & T/O & 1\\

svcomp37.c & 22 &  & true & true & false & false & 1 & 5 & term & 2.8s & 14\\

svcomp38.c & 17 & true & true & false & true & false & 1 & 5 & ? & T/O & 48\\

svcomp39.c & 56 &  & true & true & false & false & 1 & 5 & ? & T/O & 1\\

svcomp40.c & 12 &  & true & true & false & false & 1 & 5 & ? & T/O & 20\\

svcomp41.c & 12 &  & true & true & false & false & 1 & 5 & ? & T/O & 26\\

svcomp42.c & 22 &  & true & true & false & false & 1 & 5 & ? & T/O & 84\\