### 2014-07-06 papers/termination/paper.tex
various edits

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{multicol}
\usepackage{appendix}
\usepackage{caption,subcaption}
\DeclareCaptionType{copyrightbox}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{framed}
\usepackage{tikz}
\usetikzlibrary{arrows, automata, shapes, calc}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\newcommand{\xmark}{\ding{55}}
\newcommand{\todo}[1]{{\bf TODO:} #1}
\newcommand{\newC}{C$^-$\xspace}

\lstset{language=c++}

\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

\newcommand*\Let[2]{\State #1 $\gets$ #2}

\newcommand{\bv}[2]{\mathcal{BV}(#1, #2)}

\makeatletter
\pgfdeclareshape{datastore}{
  \inheritsavedanchors[from=rectangle]
  \inheritanchorborder[from=rectangle]
  \inheritanchor[from=rectangle]{center}
  \inheritanchor[from=rectangle]{base}
  \inheritanchor[from=rectangle]{north}
  \inheritanchor[from=rectangle]{north east}
  \inheritanchor[from=rectangle]{east}
  \inheritanchor[from=rectangle]{south east}
  \inheritanchor[from=rectangle]{south}
  \inheritanchor[from=rectangle]{south west}
  \inheritanchor[from=rectangle]{west}
  \inheritanchor[from=rectangle]{north west}
  \backgroundpath{
    %  store lower right in xa/ya and upper right in xb/yb
    \southwest \pgf@xa=\pgf@x \pgf@ya=\pgf@y
    \northeast \pgf@xb=\pgf@x \pgf@yb=\pgf@y
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@ya}}
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@yb}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
 }
}
\makeatother

\begin{document}

%\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{POPL '15}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{2015}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\title{Deciding Bitvector Termination with Program Synthesis}

\authorinfo{Cristina David\and Daniel Kroening\and Matt Lewis}
           {Oxford University}
           {firstname.lastname@cs.ox.ac.uk}

\maketitle

\begin{abstract}
%
Proving program termination is typically done by finding a well-founded
\emph{ranking function} for the program states.  Existing termination
provers typically find ranking functions using either linear algebra or
templates.  As such they are often restricted to finding linear ranking
functions over mathematical integers.  This class of ranking functions is
insufficient for prove termination for many terminating programs, and
furthermore a termination argument for a program operating on mathematical
integers does not always lead to a termination argument for the same program
operating on fixed-width machine integers.

We present a reduction from program \emph{termination} to program
\emph{synthesis}.  This reduction allows us to generate nonlinear,
lexicographic ranking functions that are correct for fixed-width machine
arithmetic and floating-point arithmetic.  We use this reduction to build a
sound and complete procedure for the termination of fixed-width and
floating-point arithmetic programs.
%
\end{abstract}

%\category{CR-number}{subcategory}{third-level}

\keywords
Termination, Program Synthesis, Lexicographic Ranking Functions, Bitvector Ranking Functions,
Floating-Point Ranking Functions.

\section{Introduction}\label{sec:intro}

The halting problem has been of central interest to computer scientists
since it was first considered by Turing in 1936~\cite{turing}.  Informally,
we can say that the halting problem is concerned with answering the question
``does this program run forever, or will it eventually terminate?''

Proving program termination is typically done by finding a \emph{ranking
function} for the program states, i.e.  a monotone map from the program's
state space to a well-ordered set.  Historically, the search for ranking
functions has been constrained in various syntactic ways, leading to
incompleteness, and is performed over abstractions that do not soundly
capture the behaviour of physical computers.  In this paper, we present a
sound and complete method for deciding whether a program terminates.

%% As this definition of a ranking function is very general, research is often limited to some
%% convenient and tractable form of ranking functions, most frequently \emph{linear ranking functions} (see Section~\ref{sec:ranking.functions}). 
%% In doing so, an analysis ensures that such a ranking function will be found if one  exists, 
%% while failing to prove termination for terminating programs whose ranking functions fall short of the considered restriction.

When surveying the area of program termination chronologically, we observe
an initial focus on monolithic approaches based on a single measure shown to
decrease over all program
paths~\cite{DBLP:conf/vmcai/P04,DBLP:conf/cav/BradleyMS05}, followed by more
recent techniques that use termination arguments based on Ramsey's
theorem~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
Briefly, the transition relation is shown to be disjunctively well founded
by finding a set of simple functions, each of which proves well foundedness
for part of the transition relation.  The main benefit of this approach is
the simplicity of local termination measures in contrast to global ones. 
For instance, there are cases in which linear arithmetic suffices when using
local measures, while corresponding global measures require non-linear
functions or lexicographic orders.

%This explains why approaches based on Ramsey's theorem are restricted to searching for linear termination arguments.

One drawback of the Ramsey-based approach is that the validity of the
termination argument relies on checking the \emph{transitive closure} of the
program, rather than a single step.  As such, there is experimental evidence
that most of the effort is spent in reachability
analysis~\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/KroeningSTW10},
requiring the support of powerful safety checkers: there is a tradeoff
between the complexity of the termination arguments and that of checking
their validity.

As Ramsey-based approaches are limited by the state of the art in safety
checking, recent research shifts back to more complex termination arguments
that are easier to
check~\cite{DBLP:conf/cav/KroeningSTW10,DBLP:conf/tacas/CookSZ13,DBLP:conf/cav/HeizmannHP14}. 
Following the same trend, we investigate its extreme: \emph{unrestricted}
termination arguments.  This means that our ranking functions may involve
non-linearity and lexicographic orders: we do not commit to any particular
syntactic form, and do not use templates.

Figure~\ref{fig:handletable} summarises the related work with respect to the
restrictions they impose on the transition relations as well as the form of
the ranking functions computed.  While it supports the observation that the
majority of existing termination analyses are designed for linear programs
and linear ranking functions, it also highlights another simplifying
assumption made by most state-of-the-art termination provers: that
bit-vector semantics and integer semantics give rise to the same termination
behaviour.  Thus, most of the techniques treat bit-vectors and IEEE floats
as mathematical integers and reals,
respectively~\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/vmcai/BradleyMS05,DBLP:conf/cav/KroeningSTW10}.

By assuming bit-vector semantics to be identical to integer semantics, these
techniques ignore the wrap-around behaviour caused by overflows, which can
be unsound.  In Section~\ref{sec:motivation}, we show that integers and
bit-vectors exhibit incomparable behaviours with respect to termination,
i.e.  programs that terminate for integers may \emph{not} terminate for
bit-vectors and vice versa.  Thus, abstracting bit-vectors to integers may
give rise to {\em unsound} and {\em incomplete} analyses.

%% \noindent {\bf Bit-vectors (machine-level integers) vs. mathematical integers.}
%% The abstraction of bit-vectors to mathematical integers ignores
%% the wrap-around behaviour caused by under/over-flows in bit-vector arithmetic, resulting in
%% incomparable behaviours with respect to termination.

%% Programs that terminate for integers may \emph{not} terminate for bit-vectors. For illustration, consider the following loop:
%% \begin{lstlisting}[language=C]
%% while (x > 0) x -= 2;
%% \end{lstlisting}
%% The loop always terminates for unbounded integers as the value of \texttt{x} does eventually become negative, 
%% whereas, with bit-vector arithmetic, \texttt{x} underflows before becoming negative and goes back to being positive causing the loop to never terminate.

%% Similarly, programs that terminate for bit-vectors may \emph{not} terminate for integers. One such situation is illustrated next:  
%%  \begin{lstlisting}[language=C]
%%  while(x > 0) x++;
%%  \end{lstlisting}
%% This loop is  terminating for bit-vectors since \texttt{x}
%% will eventually over-flow and become negative. Conversely, the same program is non-terminating using integer
%% arithmetic since the loop condition stays always true for any initial \texttt{x} at least 1.

%% %Similarly, programs that terminate for bit-vectors may \emph{not} terminate for integers. One such situation is illustrated next:

%% \begin{lstlisting}[language=C]
%% while (x > 0) x -= 2;
%% \end{lstlisting}

%% \begin{lstlisting}[language=C]
%%  while(x > 0) x++;
%%  \end{lstlisting}

%% %\noindent {\bf Floats vs. reals.} 
%% A scenario similar to the one for bit-vectors happens if floats were to be abstracted to unbounded reals by termination provers \cite{}. 
%% Consequently, these provers ignore potential rounding errors, under- and over-flows, which are precisely what makes  reasoning about
%% floating-point inherently difficult.
%% This approximation may lead to erroneous diagnosis of a program's terminating behaviour as illustrated by the loops in Figure~\ref{} and Figure~\ref{}.
%% While the former does not terminate for reals, but does for floats, the latter 
%% terminates for reals, but does not for floats.

%% \begin{lstlisting}
%% while (x > 0.0) x *= 0.5;
%% \end{lstlisting}

%% \begin{lstlisting}
%% while (x > 0.0) x -= 1.0;
%% \end{lstlisting}
%% \todo{explain the reasons for non-termination with figures.}\\

%% \noindent {\bf Linear programs and linear ranking functions.} As visible in Figure~\ref{fig:handletable}, 
%% most termination techniques assume \emph{restrictive transition relations}, e.g. linear programs,  and are only able to compute
%% \emph{linear ranking functions} \cite{}. To better understand this restriction, we computed the probability of a random terminating program having a linear ranking
%% function (see Section ?). This probability proved to be very low, indicating that the linearity assumption for the ranking functions is indeed prohibitive.

We propose a general framework that uniformly computes lexicographic,
non-linear ranking functions supported by non-linear inductive invariants
for loops with arbitrary guards and transitions over bit-vectors and floats. 
It also computes non-linear recurrence sets, which comprise non-termination
proofs.  Our encoding treats multiple loops and nested loops uniformly,
without having to enumerate lassos.
%
%Our technique is {\emph complete} and can handle programs with non-linear operations, e.g. logical and.
%--In our design, the termination problem becomes as hard as finding ranking functions, rather than as hard as
%checking the validity of a termination argument. 
%We combine the generation of ranking functions with the generation of invariants such that the necessary 
%supporting invariants and the ranking function are discovered at the same time.

 For this purpose, we phrase the termination and non-termination problems as
second-order satisfaction problems.  We identify a fragment of second-order
logic that is expressive enough to capture our formulation, which we call
the \emph{synthesis fragment}.  Subsequently, we propose a method for
checking the satisfiability of a formula in the synthesis fragment by
solving an isomorphic program synthesis problem
(Section~\ref{sec:synthesis}).  Our method is sound and semi-complete, i.e. 
if a pair of a ranking function and a supporting invariant exist it will be
found, but if the program does not terminate we may not find a recurrence
set.  However, if the program's transition relation is deterministic, our
method is complete.

With respect to the performance of our technique, we show that its runtime is dominated by the size of 
the shortest termination proof.
%While investigating the search space and the distribution of solutions (Section~\ref{}), we identify factors 
%influencing our method's performance and
%find that our technique  
%
This means that our procedure is not directly dependent on the structure of
the analysed loop (e.g.~the number of variables or the number of lassos),
but on the \emph{Kolmogorov complexity of its minimal termination argument}. 
We argue that the length of a program's termination proof is indicative of
how easy to understand the program is, and thus programmers tend to write
programs that have relatively short termination proofs.  We empirically show
that this claim holds in practice, causing our technique to perform well.

%% Our approach to proving termination has two distinct features over previous work. 
%% First, it is \emph{general}, i.e. does not make any assumption about the form of the termination argument.   
%% Second, it correctly handles bit-vectors and floats without abstracting them to integers and reals, respectively.

\begin{figure*}
\centering
 \begin{tabular}{|ll||c|c|c|c|c|c|c|c|}
 \hline
  & & \multicolumn{8}{c|}{Program} \\
  & & \multicolumn{2}{c|}{Integers} & \multicolumn{2}{c|}{Reals} & \multicolumn{2}{c|}{Bit-vectors} & \multicolumn{2}{c|}{Floats} \\
  & & L & NL & L & NL & L & NL & L & NL \\
  \hline
  \hline
  \multirow{4}{*}{Ranking} & Linear lexicographic &  \cite{DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13,DBLP:conf/vmcai/P04} && & &\checkmark&\checkmark&\checkmark&\checkmark\\
   & Linear non-lexicographic & \cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/vmcai/BradleyMS05,DBLP:conf/cav/KroeningSTW10} & \cite{DBLP:conf/vmcai/BradleyMS05} & && \checkmark~ \cite{DBLP:conf/tacas/CookKRW10} &\checkmark~ \cite{DBLP:conf/tacas/CookKRW10}&\checkmark&\checkmark\\
   & Non-linear lexicographic &  &  & &&\checkmark&\checkmark&\checkmark&\checkmark\\
   & Non-linear non-lexicographic & \cite{DBLP:conf/vmcai/BradleyMS05} &  \cite{DBLP:conf/vmcai/BradleyMS05} & &&\checkmark&\checkmark&\checkmark&\checkmark\\
   \hline
 \end{tabular}

 \caption{Legend: \checkmark = we can handle\label{fig:handletable}}
\end{figure*}

%We empirically observe that most programs have
%ranking functions with low Kolmogorov complexity (see Section ?).

%Church was BIG into program synthesis~\cite{church-synth}, so you know it's good stuff.
%Something, something, Curry-Howard Isomorphism, something, something, programs-as-proofs.

The main contributions of our work can be summarised as follows:
%
\begin{itemize}

\item We designed a bit-level accurate technique for computing ranking
functions that correctly accounts for the wrap-around behavior caused by
under- and overflows in bit-vector and floating-point arithmetic.  Our
technique is not restricted to finding linear ranking functions, but can
also compute lexicographic non-linear ones.

\item  We rephrased the termination and non-termination problems as
second-order satisfaction problems and used program synthesis techniques to
solve them.  In contrast to previous approaches to program synthesis, we
found that using genetic programming was very effective at generating
candidate termination and non-termination proofs.

\item Our procedure does not require the enumeration of lassos, or directly
reasoning about the transitive closure of the transition relation.  To
verify our proofs, we only require a single call to a SAT solver.

%\item We present a formulation that allows us to prove termination without an expensive reachability check (as we don't use Ramsey-based termination arguments).  In particular,
%we only need a bounded model checker that performs a single unwinding of a loop.

%% \item Our technique is able to uniformly handle conditionally-terminating loops, as well as programs with
%% multiple loops.

\item Our procedure is based on a series of source-to-source
transformations.  As well as leading to a concise implementation, this has
the side effect of ensuring that our proof is correct for the exact binary
produced by a particular compiler, rather than for an abstraction of the
program.

\item We implemented our technique and tried it on a selection of programs
handling both bit-vectors and floats.

%\item We propose a non-trivial formulation of the (non-)termination problem inside a
% second-order logic fragment with only one quantifier alternation.

\end{itemize}
% unified approach to termination and non-termination
%. only non-strict inequalities can be transformed using Farkas’ Lemma
%--Our approach to termination analysis has two distinct features over previous works
%-- a lexicographic ranking function imposes a lexicographic ordering on among the ranking function components.

% The main advantage of our approach is its unitary nature. We do not require any initial assumption regarding the form of the termination  
% argument, as this does not influence our search process. We identify the factors influencing the performance of our design and 
% show that these factors are independent on the linearity of the ranking function or the number of lexicographic components...

%semantic approach

%given that the search space size is constant, the probability of finding a solution 
%increases when considering all the possible solutions. There is no point in restricting the form of the solution..
%to read: constraint-based.., integers vs rationals.

%This paper is organised as follows:

{\bf Limitations.}
Our algorithm decides termination of transition systems with finite state
spaces.  The (non-)termination proofs take the form of ranking functions and
program invariants that are expressed in a quantifier free language.  This
formalism is powerful enough to handle a large fragment of C, but is not
rich enough to analyse code using any of the following features:

\begin{itemize}
\item Recursion (could probably do this).
\item Heap (Thor).
\item Arrays.
\end{itemize}

Additionally, if the program under analysis has a non-deterministic transition relation
and is non-terminating, our procedure is not guaranteed to find a non-termination proof.

\section{Motivating Examples} \label{sec:motivation}
%- see phase change programs.

Figure~\ref{fig:handletable} illustrates the most common simplifying
assumptions made by existing termination analyses:
%
\begin{itemize}
\item[(i)] programs use only linear arithmetic.
\item[(ii)] terminating programs have termination arguments expressible in linear arithmetic.
\item[(iii)] the semantics of bit vectors and mathematical integers are equivalent.
\item[(iv)] the semantics of IEEE floating-point numbers and mathematical reals are equivalent.
\end{itemize}

%In the previous sections, we have discussed some of the assumptions made by existent termination analyses (see Figure~\ref{fig:handletable}).
%
To show how these assumptions are violated by even simple programs, we draw
the reader's attention to the programs in Figure~\ref{fig:motivation} and
their curious properties:
%
%We next discus some intricate examples that prove challenging for existent termination analysis.
%
%\subsection{Program linearity vs. the linearity of its ranking function.}
%
%Figure~\ref{fig:handletable} shows that the most common assumption existent termination techniques make is that \emph{programs are linear}.
%
\begin{itemize}

\item Program (a) breaks assumption (i) as it makes use of the bit-wise $\&$ operator.
%
%As such, Figure~\ref{} denotes a program with non-linear operations. Given
%that many of the existing techniques commit to linear programs, they cannot
%handle this situation, although a linear ranking function does exist (see
%Figure~\ref{fig:handletable}).  In order to find a ranking function for
%this example, it is necessary to take into account the semantics of the
%bit-wise AND operator.
%
Our technique finds that an admissible ranking function is the linear
function $R(x) = x$, whose value decreases with every iteration, but cannot
decrease indefinitely as it is bounded from below.  This example also
illustrates the lack of a direct correlation between the linearity of a
program and that of its termination arguments.

\item Program (b) breaks assumption (ii), in that it has no linear ranking
function.  We prove that this loop terminates by finding the non-linear
ranking function $R(x) = |x|$.

\item Program (c) breaks assumption (iii).  This loop is terminating for
bit-vectors since $x$ will eventually over-flow and become negative. 
Conversely, the same program is non-terminating using integer arithmetic
since $x > 0 \rightarrow x+1 > 0$ for any integer $x$.

\item Program (d) also breaks assumption (iii), but ``the other way'': it
terminates for integers but not for bit-vectors.  If each of the variables
is stored in an unsigned $k$-bit word, the following entry state will lead
to an infinite loop:
%
$$ M = 2^k - 1,\quad N = 2^k - 1,\quad i = M,\quad j = N-1 $$

\item Program (e) breaks assumption (iv): it terminates for reals but not
for floats.  If $x$ is sufficiently large, rounding error will cause the
subtraction to have no effect.

\item Program (f) breaks assumption (iv) ``the other way'': it terminates
for floats but not for reals.  Eventually $x$ will become sufficiently small
that the nearest representable number is $0.0$, at which point it will be
rounded to $0.0$ and the loop will terminate.

\end{itemize}

Up until this point, we considered examples that are not soundly treated by
existing techniques as they don't fit in the range of programs addressed by
these techniques.  Next, we look at some programs that are handled by
existing termination tools via dedicated analyses.  We show that our method
handles them uniformly, without the need for any special treatment.
%
\begin{itemize}

\item Program (g) is a linear program that is shown
in~\cite{DBLP:conf/tacas/CookSZ13} not to admit (without prior manipulation)
a lexicographic linear ranking function.  With our technique we can find the
non-linear ranking function $R(x) = |x|$.

%As with linear lexicographic ranking functions, a state is mapped to a tuple of values such that the
%loop transition leads to a decrease with respect to the lexicographic
%ordering for this tuple. Therefore no function may increase unless a function of
%a lower index decreases. Additionally, at every step, there must be at least one
%function that decreases.

\item Program (h) illustrates conditional termination.  When proving program
termination we are simultaneously solving two problems: the search for a
termination argument, and the search for a supporting
invariant~\cite{DBLP:conf/cav/BrockschmidtCF13}.  For this loop, we find the
ranking function $R(x) = x$ together with the supporting invariant $y=1$.

\item In the terminology of \cite{DBLP:conf/tacas/LeikeH14}, program (i)
admits a \emph{multiphase} ranking function, computed from a multiphase
ranking template.  Multiphase ranking templates are targeted at programs
that go through a finite number of phases in their execution.  Each phase is
ranked with an affine-linear function and the phase is considered to be
completed once this function becomes non-positive.

In our setting this type of programs does not need special treatment, as we
can find a non-linear lexicographic ranking function $R(x, y, z) = (x < y,
z)$.\footnote{This termination argument is somewhat subtle, but can be understood as
follows: The Boolean values $\mathit{false}$ and $\mathit{true}$ are
interpreted as 0 and 1, respectively.  The Boolean $x < y$ thus eventually
decreases, that is to say once a state with $x \geq y$ is reached, $x$ never
again becomes greater than $y$.  This means that as soon as the ``else''
branch of the if statement is taken, it will continue to be taken in each
subsequent iteration of the loop.  Meanwhile, if $x < y$ has not decreased
(i.e., we have stayed in the same branch of the ``if''), then $z$ does
decrease.  Since a Boolean only has two possible values, it cannot decrease
indefinitely.  Since $z > 0$ is a conjunct of the loop guard, $z$ cannot
decrease indefinitely, and so $R$ proves that the loop is well founded.}
\iffalse
\item Program (j) illustrates a nested loop.  This construction can cause difficulties for provers based on
enumerating lassos
\fi
\end{itemize}
%
As with all of the termination proofs presented in this paper, the ranking
functions above were all found completely automatically.

\begin{figure*}
\centering
 \begin{tabular}{ccc}

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x > 0) {
  x = (x - 1) & x;
}
\end{lstlisting}
\caption{Taken from~\cite{DBLP:conf/tacas/CookKRW10}.}
 \label{fig:motivation.a}
\end{subfigure}%

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x != 0) {
  x = -x / 2;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.b}
\end{subfigure}%

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}[language=C]
while(x > 0) {
  x++;
}
 \end{lstlisting}
\caption{}
 \label{fig:motivation.c}
\end{subfigure} \\

\hline

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (i < M || j < N) {
  i = i + 1;
  j = j + 1;
}
\end{lstlisting}
\caption{Taken from~\cite{DBLP:conf/sigsoft/Nori013}}
 \label{fig:motivation.d}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
float x;

while (x > 0.0) {
  x -= 1.0;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.e}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
float x;

while (x > 0.0) {
  x *= 0.5;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.f}
\end{subfigure} \\
\hline

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x != 0) {
  if (x > 0)
    x--;
  else
    x++;
}
\end{lstlisting}
\caption{Taken from \cite{DBLP:conf/tacas/CookSZ13}}
 \label{fig:motivation.g}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
y = 1;

while (x > 0) {
  x = x - y;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.h}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x > 0 && y > 0 && z > 0) {
  if (y > x) {
    y = z;
    x = nondet();
    z = x - 1;
  } else {
    z = z - 1;
    x = nondet();
    y = x - 1;
  }
}
\end{lstlisting}
\caption{Taken from~\cite{BA:mcs}}
 \label{fig:motivation.i}
\end{subfigure} 
% 
% \\
% 
% \hline
% 
% &
% 
% \begin{subfigure}[b]{0.3\textwidth}
% \begin{lstlisting}
% while (i < n) {
%   j = 0;
%   while (j <= i) {
%     j = j + 1;
%   }
%   i = i + 1;
% }
% \end{lstlisting}
% \caption{Taken from~\cite{DBLP:conf/cav/BrockschmidtCF13}}
% \end{subfigure}

\end{tabular}
\caption{Motivational examples, mostly taken from the literature.\label{fig:motivation}}
\end{figure*}

%\subsection{Differences in the termination behaviour for integers and bit-vectors.}
%We have collected a number of motivational examples from other termination papers that treat bit-vectors as mathematical integers \cite{DBLP:conf/tacas/LeikeH14,DBLP:conf/tacas/CookSZ13}. 
%We show that the termination arguments computed by such techniques do not directly apply when considering the bit-vector semantics.
%these programs actually exhibit different terminating behaviours for bit-vectors and integer

%% \begin{lstlisting}
%%   while (x != m) {
%%     if (x > m)
%%       x = 0;
%%     else
%%       x++;
%%   }
%% \end{lstlisting}

%% The program in Figure~\ref{} is taken from \cite{DBLP:conf/tacas/CookSZ13}, where
%% $m$ and $x$ start as any integers with $m$ positive. If $x$ is greater than
%% $m$, $x$ is set to 0. Otherwise, $x$ increases until it equals $m$, upon which
%% the loop terminates. While a disjunctive well-founded termination argument does exist for the loop, 
%% e.g. ($x < oldx$ and $0 \leq oldx$) or ($m-x < oldm-oldx$ and $0 \leq oldm-oldx$), the loop does not 
%% terminate under the bit-vector semantics. The reason is ..

%\subsection{Multi-phase ranking functions}
%\subsection{Lexicographic ranking function with strict inequalities.}
%Approaches based on Farkas’ Lemma can only handle non-strict inequalities \cite{DBLP:conf/cav/BradleyMS05,DBLP:conf/vmcai/P04}.

\section{Preliminaries}

Given a program, we first formalise its termination argument as a ranking
function (Section~\ref{sec:ranking.functions}).  Subsequently, we discus
bit-vector semantics and illustrate differences between machine arithmetic
and integer arithmetic that show that the abstraction of bit vectors to
mathematical integers unsound (Section~\ref{sec:machine.arith}).

\subsection{Termination and Ranking Functions} \label{sec:ranking.functions}

A programs $P$ is represented as a transition systems with state space $X$,
transition relation $T \subseteq X \times X$, and set of variables $V$.  A
state of the program is a valuation of the variables from $V$.  For a state
$x \in X$ with $T(x,x')$ we say $x'$ is the successor of $x$ under $T$.

\begin{definition}[Unconditional termination]
%
A transition system is said to be \emph{unconditionally terminating} if
there is no infinite sequence of states $x_1, x_2, \ldots \in X$ with
$\forall i.~T(x_i, x_{i+1})$.
%
\end{definition}

We can prove that the transition system is unconditionally terminating by
finding a ranking function for its transition relation.
%
\begin{definition}[Ranking function]
%
An injective function ${R:X\to Y}$ is a \emph{ranking function} for the
transition relation $T$ if $Y$ is a well-founded set with order $<$ and 
$R$ is monotonically decreasing with respect to $T$.  That is
to say:
$$\forall x, x' \in X. T(x, x') \Rightarrow R(x) < R(x')$$
%
\end{definition}

\begin{definition}[Linear ranking function]
A \emph{linear ranking function} $R: X \to Y$ 
with $\dim(X) = n$ and $\dim(Y) = m$ is of the form: $$f(\vec{x}) = M\vec{x}$$ where
$M$ is an $n \times m$ matrix.
\end{definition}

In the case that $\dim(Y) = 1$, this reduces to the inner product
%
$$f(\vec{x}) = \vec{\lambda} \cdotp \vec{x} + c \;.$$

\begin{definition}[Lexicographic ranking function]
For $Y = Z^m$, we say that a ranking function $R: X \to Y$ is \emph{lexicographic}
if it maps each state in $X$ to a tuple of values such that the loop transition leads to a decrease with
respect to the lexicographic ordering for this tuple.
The total order imposed on $Y$ is the lexicographic ordering
induced on tuples of $Z$'s. 
\end{definition}

For lexicographic ranking functions, no function may increase unless a function of a lower index decreases.
Additionally, at every step, there must be at least one function that decreases.
We note that some termination arguments require lexicographic ranking functions, or equivalently, ranking functions
whose co-domain is the ordinals, rather than just $\mathbb{N}$.
%In the rest of the paper, we consider any ranking function $R: X \to Z^m$ to be lexicographic, and the situation when $m=1$ to be a special case.

%% \begin{definition}{Supportive invariant}
%% \end{definition}

\subsection{Machine Arithmetic and Bit Vectors} \label{sec:machine.arith}

Physical computers have bounded storage, which means they are unable to
perform calculations on mathematical integers.  For example, if $A$ is the
Ackermann function and $G$ is Graham's number, a physical computer capable
of computing $A(G, G)$ would contain (much!) more matter than is believed to
exist in the universe.  Fortunately, it is rare for a programmer to need
such a large number and so modern computers do their arithmetic over
fixed-width binary words, otherwise known as bit vectors.  For the remainder
of this section, we will say that the bit vectors we are working with are
$k$-bits wide, which means that each word can hold one of $2^k$ bit
patterns.  Typical values for $k$ are 32 and 64.

Machine words can be interpreted as ``signed'' or ``unsigned'' values. 
Signed values can be negative, while unsigned values cannot.  The encoding
for signed values is two's complement, where the most significant bit
$b_{k-1}$ of the word is a ``sign'' bit, whose weight is $-(2^k - 1)$ rather
than $2^k - 1$.  Two's complement representation has the property that
$\forall x .  -x = (\mathord{\sim} x) + 1$, where $\mathord{\sim}(\bullet)$
is bitwise-negation.  Two's complement also has the property that addition,
multiplication and subtraction are defined identically for unsigned and
signed numbers.

Bit-vector arithmetic is performed modulo $2^k$, which is the source of many
of the differences between machine arithmetic and Peano
arithmetic.\footnote{ISO C requires that unsigned arithmetic is performed
modulo $2^k$, whereas the overflow case is undefined for signed arithmetic. 
In practice, the undefined behaviour is implemented just as if the
arithmetic had been unsigned}.  To give an example, $(2^k - 1) + 1 \equiv 0
\pmod {2^k}$ provides a counterexample to the statement $\forall x. 
x + 1 > x$, which is a theorem of Peano arithmetic but not of modular
arithmetic.  When an arithmetic operation has a result greater than $2^k$,
it is said to ``overflow''.  If an operation does not overflow, its
machine-arithmetic result is the same as the result of the same operation
performed on integers.

The final source of difference between integer arithmetic and bitvector
arithmetic stems from width conversions.  Many programming languages allow
numeric variables of different types, which can be represented using words
of different widths.  In C, a \texttt{short} might occupy 16 bits, while an
\texttt{int} might occupy 32 bits.  When a $k$-bit variable is assigned to a
$l$-bit variable with $l < k$, the result is truncated $\mod 2^l$.  For
example, if $x$ is a 32-bit variable and $y$ is a 16-bit variable, $y$ will
hold the value $0$ after the following code is executed:
%
\begin{lstlisting}
x = 65536;
y = x;
\end{lstlisting}

This gives us a counterexample to the statement $\forall x, y . x = y
\Rightarrow (x + 1) = (y + 1)$, which is a theorem of Peano arithmetic.

As well as machine arithmetic differing from Peano arithmetic on the
operators they have in common, computers have several ``bitwise'' operations
that are not taken as primitive in the theory of integers.  These operations
include the Boolean operators \texttt{and, or, not, xor} applied to each
element of the bit vector.  Computer programs often make use of these
operators, which are non-linear when interpreted in the standard model of
Peano arithmetic\footnote{Some of these operators can be seen as
linear in a different algebraic structure, e.g.~\texttt{xor} corresponds to
addition in the Galois field $\mathrm{GF}(2^k)$.}.

\iffalse
\section{Combinatorics of Finite Termination}
In this section, we fix a model of computation, describe its semantics and
define the syntax of a language we will work over.

\subsection{Syntax and Semantics}

\begin{itemize}
 \item Our transition relation is $T(x, x') \subseteq X \times X$.
 \item Our loop condition is $L(x) \subseteq X$.
 \item Our ranking function is $R(x) : X \to Y$.
 \item Our state space has size $\| X \| = M = 2^k$.
 \item Our ranking co-domain has size $\| Y \| = N = 2^j$.
 \item The number of looping states is $\| L \| = l$.
 \item Our transition relation is deterministic and parititioned into chains of length $c_i$, with $l = \sum c_i$.
\end{itemize}

\subsection{Counting Programs}
\begin{itemize}
 \item There are a TON of programs (way more than you'd expect).
 \item There are a TON of terminating programs, and for our model of computation we can count
  how many (the Chaitin constant).
 \item There are a TON of ranking functions (way more than you'd expect, but not many as a
  fraction of programs).
 \item There are not many linear functions.
 \item Most terminating programs don't admit linear ranking functions.
 \item The Curry-Howard isomorphism
 \item Kolmogorov complexity is relevant for understanding termination proofs.
\end{itemize}

\begin{theorem}
 A random function $f : X \to Y$ is a ranking function for $(T, L)$ with probability

 $$N^{-l} \times \prod {{N-1} \choose c_i}$$
\end{theorem}

\begin{proof}
 Combinatorics.
\end{proof}

\begin{corollary}
 This number is really small (e.g. $10^{-193}$ for a 64-bit program with 1 variable and 10 looping states.
 Randomly sampling functions \& hoping they're ranking functions is not going to work.
\end{corollary}

\begin{conjecture}
 The probability that a random program $(T, L)$ is terminating (the Chaitin constant)
 is $0.7$.
\end{conjecture}

\begin{conjecture}
 The probability that a random program $(T, L)$ admits a linear ranking function is
 $0.1$.
\end{conjecture}

\begin{conjecture}
 The probability that a random, terminating program $(T, L)$ admits a linear ranking function
 is $0.2$.
\end{conjecture}

\begin{corollary}
 Most terminating programs do not have linear ranking functions.
\end{corollary}
\fi

\section{Termination as Second-Order Satisfaction} \label{sec:second.order}
%Program model? Programs: transition systems? Constraint generation? No CFG?

%% One difference from previous approaches is that our method is fully semantic. As a consequence, 
%% we do not care about the CFG (we do not need to encode the CFG 
%% through program counter in the second-order constraints).

The problem of program verification can be reduced to the problem of finding
solutions to a second-order
constraint~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08}. 
Our intention is to apply this approach to termination analysis.  In this
section we show how several variations of both the termination and the
non-termination problem can be defined in second-order logic.
%
%The second-order unknowns in this constraint are the unknown program invariants
%that are inductive and strong enough to prove the desired assertions. In this section we describe the conversion of programs to constraints

In fact, we identify a fragment of second-order logic with a constrained use
of quantification that is expressive enough to encode both termination and
non-termination.  We will suggestively refer to the fragment as the
\emph{synthesis fragment}:

% The synthesis fragment is motivated by practical requirements in software verification.

\begin{definition}[Synthesis Fragment]
A formula is in the \emph{synthesis fragment} iff it is of the form
%
 \[
  \exists \vec{P},~ \vec{x} . ~\forall~ \vec{y} . \sigma(\vec{P}, \vec{x}, \vec{y})
 \]
%
where $\vec{P}$ ranges over sets, while $\vec{x}$ and $\vec{y}$ range over
ground terms.  Function $\sigma: (X^n \times Y^m \to Z^k) \times X^n \times
Y^m \to \mathbb{B}$ can be seen as a specification function that returns
true iff the functions $\vec{P}$ compute appropriate outputs when fed the
inputs $\vec{x}$ and $\vec{y}$.
%
\end{definition}
%
Checking satisfiability of a formula in the synthesis fragment corresponds to program synthesis and
%Solving the synthesis problem 
amounts to finding witnesses for the unknowns $\vec{P}$ and $\vec{x}$ that meet the specification
for all $\vec{y}$. 
If a pair $(\vec{P_0}, \vec{x})$ is a solution to the synthesis problem, then we write $(\vec{P_0}, \vec{x}) \models \sigma$.
For the remainder of the presentation, we drop the $\vec{x}$ and instead just write $x$, with the understanding
that all variables range over vectors.

In the rest of this section, we show that the synthesis fragment 
is expressive enough to encode both termination and non-termination. %allow the formulation of both the termination and the non-termination problem. 
Following our formulation as second-order satisfaction, %the framing of the halting problem as a program synthesis problem,
in Section~\ref{sec:synthesis} we present a solver for the synthesis fragment that models bit-accurate semantics.

%% Following the framing of the halting problem as a program synthesis problem, 
%% we show how we solve this problem in Section~\ref{sec:synthesis}.

\subsection{An Isolated Loop}

We will begin our discussion by showing how to encode the
\mbox{(non-)termination} of a program consisting of a single loop in the
synthesis fragment.  For the time being, a loop is a pair ${L= \langle G,T
\rangle}$, where the states $x$ satisfying the loop's guard are given by the
predicate $G(x)$.  The body of the loop is encoded as the transition
relation $T(x, x')$, meaning that state $x'$ is reachable from state $x$ via
a single iteration of the loop body.  For example, the loop in
Figure~\ref{fig:motivation.a} is encoded as:
%
\begin{align*}
G(x) & = \{ x \mid x>0 \} \\
T(x,x') &= \{ \langle x, x' \rangle \mid x' = (x - 1) \, \& \, x \}
\end{align*}
We will abbreviate this with the notation:
\begin{align*}
G(x) & \triangleq x > 0 \\
T(x, x') & \triangleq x' = (x - 1) \, \& \, x
\end{align*}

\begin{figure*}
\begin{framed}
\begin{definition}[Unconditional Termination Formula {\bf [UT]}]
\label{def:UT}
\begin{align*}
 \exists R . \forall x, x' . & G(x) \wedge T_o(x, x') \rightarrow R(x) > 0 \wedge R(x) > R(x')
\end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula -- Open Recurrence Set  {\bf [ONT]}]
\label{def:ont}
 \begin{align*}
  \exists N, x_0 . \forall x . \exists x' . & N(x_0) ~ \wedge ~ N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \rightarrow T_u(x, x') \wedge N(x') 
 \end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula -- Closed Recurrence Set {\bf [CNT]}]
\label{def:cnt}
 \begin{align*}
  \exists N, x_0 . \forall x, x' . & N(x_0) ~ \wedge ~ N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T_u(x, x') \rightarrow N(x') 
 \end{align*}
\end{definition}

%% \begin{definition}[Non-deterministic Non-Termination Formula - Closed Recurrence Set {\bf [Nondet-CNT]}]
%% \label{def:nondet-cnt}
%%  \begin{align*}
%%   \exists N, x_0 . \forall x, x' . \exists y_0. & N(x_0) ~ \wedge ~ N(x) \rightarrow G(x) ~ \wedge \\
%% 							& N(x) \wedge T_u(x, x') \wedge A(x') \rightarrow N(x')~
%%  \end{align*}
%% \end{definition}

\end{framed}
\end{figure*}

\iffalse
Many loops do not terminate for all starting states, but are contained in programs that guarantee the loop will
terminate.  Traditional termination provers have difficulty reasoning about such conditionally-terminating loops.
We are able to handle such loops by computing \emph{termination invariants}.  This mechanism also allows us to
prove that programs with multiple loops terminate, even if the termination of some loop depends on the states
reachable after leaving a previous loop.

Our method for ranking function synthesis can be stated as follows:
discuss what spec is used (non-lexicographic vs lexicographic) + the completeness claims.
Any termination guarantees?  
\fi

\noindent {\bf Unconditional termination.}
To show that $L$ is unconditionally terminating (i.e. it eventually terminates regardless of the
state it starts in), we try to find a ranking function for it. When defining $L$'s termination 
formula, we need only consider the over-approximation of its transition relation $T$, which we call $T_o$.
Note that this distinction is only important if $L$'s body contains nested loops, for which we'll have to 
consider the over-approximation of their transition relation's transitive closure
as shown in Section~\ref{sec:env}. For any other statements, $T_o$ and $T$ are equivalent.

\begin{theorem}
\label{thm:ut}
 The loop $L(G, T_o)$ terminates from every start state iff formula {\bf [UT]} (Definition~\ref{def:UT}) is
 satisfiable.
\end{theorem}

\begin{proof}
 Then the codomain of $R$ is a well founded set and $R$ is an order homomorphism.

 (Do this proof properly).
\end{proof}

As the existence of a ranking function is equivalent to the satisfiability
of the formula {\bf [UT]}, a satisfiability witness is also a ranking
function and thus a proof of $L$'s unconditional termination.

Continuing the example of the program in Figure~\ref{fig:motivation.a}, we
can see that the corresponding synthesis formula {\bf [UT]} is satisfiable,
as witnessed by the function $R(x) = x$.  Thus, $R(x) = x$ constitutes a
proof that the program in Figure~\ref{fig:motivation.a} is unconditionally
terminating.

Note that different formulations for unconditional termination are possible. 
We are aware of a proof rule based on transition invariants, i.e.  supersets
of the transition relation's transitive
closure~\cite{DBLP:conf/pldi/GrebenshchikovLPR12}.  This formulation assumes
that the second-order logic has a primitive predicate for disjunctive
well-foundedness.  Conversely, our formulation in Definition~\ref{def:UT}
does not require a transitive closure operator, nor a disjunctive
well-foundedness predicate.  \\

\noindent{\bf Non-termination.}
Dually to termination, we might want to consider the non-termination of a program.  If a program terminates,
we can prove this by finding a ranking function %and supporting invariant 
witnessing the satisfiability of formula {\bf[UT]}.  What then would a proof of non-termination look like?

Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterize non-termination
of a transition relation $T$ by the existence of an \emph{(open) recurrence
set}, i.e.  a nonempty set of states $N$ such that for each $s \in N$ there
exists a transition to some $s'\in N$.
For our case, the notion of open recurrent set is encoded by formula {\bf [ONT]}. %Definition~\ref{def:nonterm-formula}.
Dual to the termination problem, when formulating $L$'s non-termination, we only need to consider an under-approximation of its transition
relation~$T_u$. 
%$L = \langle G,T \rangle$
Again, this distinction is mostly important when dealing with nested loops,
as illustrated in Section~\ref{sec:env}.

\begin{theorem}
\label{thm:ont}
 The loop $L(G, T_u)$ has an infinite execution iff Defintion~\ref{def:ont} is satisfiable.
\end{theorem}

\begin{proof}
 Then $x_0$ is an underapproximation to the reachable states on entry to $L$, and $N$ is an open recurrence set.

 (Do this proof properly).
\end{proof}

If this formula is satisfiable, $N$ is an open recurrence set for $L$, which proves
$L$'s non-termination. The issue with this formula is the additional level of quantifier alternation as compared to the synthesis fragment
(it is an $\exists \forall \exists$ formula). %This means that we will not be able to use the solver for the synthesis fragment to solve it.

In order to formulate the non-termination problem inside the synthesis
fragment, we make use of the notion of \emph{closed recurrence set}
introduced by Chen et al.~in~\cite{DBLP:conf/tacas/ChenCFNO14}.  As opposed
to an open recurrence set, this is an $\exists \forall$ property: for each
state in the recurrence set $N$, all of its successors must be in $N$.  We
note that if $T$ is deterministic, every open recurrence set is also a
closed recurrence set (since each state has at most one successor).  Thus,
the non-termination problem for deterministic transition systems is
equivalent to the satisfiability of formula {\bf [CNT]}.

%Definition~\ref{def:deterministic-nonterm-formula}.  As our formulation is
%with respect to a given loop $L=\langle I,G,T \rangle$, we require the
%recurrence set to include the loop's guard, $N(x) \rightarrow G(x)$.  In
%other words, the loop's guard needs to be satisfied in order for the
%infinite execution to happen.

\begin{theorem}
\label{thm:cnt}
 If $T_u$ is deterministic, i.e. each state $x$ for which the guard $G(x)$ holds has exactly one successor $x'$ such that $T_u(x, x')$, then
 Definition~\ref{def:ont} is equivalent to Definition~\ref{def:cnt}.
\end{theorem}

\begin{proof}
 If $T_u$ is deterministic, then each $x$ satisfying $G(x)$ has a unique successor $S(x)$, so $\forall x, x' . G(x) \wedge T_u(x, x') \leftrightarrow x' = S(x)$.

 (Finish this proof).
\end{proof}

Note that by defining $T_u$'s determinism as the existence of exactly one
successor for any state satisfying $L$'s guard, we eliminate the possibility
of $T_u$ being $\mathit{false}$, which would otherwise constitute a
counterexample for Theorem~\ref{thm:cnt}.

Given that our program's state space is finite, a transition relation
induces a infinite execution if and only if some state is visited infinitely
often, or equivalently the state is visited twice in the course of the
execution.  This gives us an alternative way of expression a non-termination
proof: we can can try to find a witness $x$ to the satisfiability of the
formula
%
$$ \exists x . T^+(x, x)$$
%
However, deciding satisfiability of this formula directly requires a logic
that includes a transitive closure operator, which ours does not have.\\

\noindent {\bf Non-determinism.} A non-termination diagnosis procedure based
on checking the satisfiability of formula {\bf [CNT]} may fail to find a
closed recurrence set for a loop $L=(G, T_u)$ when $T_u$ is
non-deterministic.  To understand the issue, imagine the following loop with
a non-deterministic assignment to \lstinline!y!:

\begin{lstlisting}[language=C]
while(x != 0) {
  y = nondet();
  x = x-y;
}
\end{lstlisting}

Only for certain choices of \texttt{y} the loop is non-terminating.  In
order to restrict the values that may be assigned to \texttt{y}, we need an
assumption associated with the non-deterministic choice.  However, such an
assumption may cause a reachable state at the non-deterministic assignment
to not have any successor.  In such a case the termination formula would be
satisfiable, although the execution halts.  This situation can be avoided by
checking the existence of at least one successor for each reachable state at
the non-deterministic assignment.  Although we did put a reasonable amount
of effort into it, we failed to come up with a formulation for the
non-termination of such non-deterministic transition relations inside the
synthesis fragment.
%
%formula {\bf [Nondet-CNT]} in Definition~\ref{def:nondet-cnt}.

\iffalse
\noindent {\bf Non-determinism.} A non-termination diagnosis procedure searching for a closed recurrence set under-approximates the transition relation, i.e. 
the closed recurrence set together with the under-approximation characterise non-termination \cite{DBLP:conf/tacas/ChenCFNO14}.
A potential difficulty with such a procedure lies in the handling of non-deterministic assignments. For illustration, consider the following loop:

\begin{lstlisting}[language=C]
while(x != 0) {
  y = nondet();
  x = x-y;
}
\end{lstlisting}

Only for certain values of $y$ the loop is non-terminating.
In order to restrict the values that may be assigned to $y$, we replace the non-deterministic choice with a fresh variable $a$ and use formula {\bf [CNT]}
to compute a closed recurrence set $N[a]$ parametrised by $a$. Now $N[a]$ becomes an assertion over $x \cup \{a\}$.
%, i.e. we under-approximate the transition relation. 
For this example, our technique computes $N[a](x,y,a) = ...$. \\%, where $a$ corresponds to the non-deterministic choice.  \\
%The issue with the closed recurrence set is that it yields an incomplete  for loops 
%that have non-deterministic assignments 
%parametric recurrence set

\fi

%% According to \cite{DBLP:conf/popl/GuptaHMRX08,DBLP:conf/tacas/ChenCFNO14}, a transition relation is non-terminating iff it has a recurrence set of states.
%% Candidate transition relations are 
%% \emph{lassos} in the program's CFG.
%% A lasso at a cutpoint location consists of two sequences
%% of transitions, the stem and the loop.
%% The stem is a finite path from an initial location to the cutpoint location. 
%% The loop is a finite path that starts and ends at the cutpoint
%% location by following a cyclic path through the CFG. 
%% A lasso induces an infinite execution if the infinite path obtained by traversing the stem and then unrolling
%% the loop infinitely many times is feasible. In this case, the lasso is said to be non-terminating.
%% We observe that in the case that $T$ is deterministic (that is, each state has
%% exactly one successor),  the non-termination problem is equivalent to the
%% satisfiability of Definition~\ref{def:deterministic-nonterm-formula}.

%% The method is incomplete as not all non-terminating
%% program executions are induced by lassos. An infinite behaviour
%% that is not periodic is illustrated by the following program:

%Our method is purely semantic.

\begin{figure*}
 \begin{framed}

\begin{definition}[Conditional Termination Formula {\bf [CT]}]
\label{def:ct}
 \begin{align*}
  \exists R, W . \forall x, x' . & I(x) \wedge G(x) \rightarrow W(x) ~ \wedge \\
                                 & G(x) \wedge W(x) \wedge T_o(x, x') \rightarrow W(x') \wedge R(x) > 0 
  \wedge R(x) > R(x')
 \end{align*}
\end{definition}

%%  \begin{definition}[Sequential Loops Termination Formula {\bf [ST]}]
%%   \begin{align*}
%% \label{def:multi-termination-formula}
%%   \exists R, Inv_1,..., Inv_n . \forall x_0,...,x_n, x_1',...,x_n'.  & P_0(x_0,x_1) \rightarrow Inv_1(x_1) ~  \\
%%  & \bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) ~  \\  
%%  & \bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) \rightarrow Inv_{i+1}(x_{i+1})) ~ \wedge \\
%%  & Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge G_n(x_n) \rightarrow Inv_n(x_n) ~ \wedge \\
%%  & Inv_n(x_n) \wedge G_n(x_n) \wedge T(x_n, x_n') \rightarrow Inv_n(x_n') \wedge R(x_n) > 0 \wedge R(x_n) > R(x_n')
%%  \end{align*}
%% \end{definition}

%% \begin{definition}[Nested Loops Termination Formula {\bf [NT]}]
%% \label{def:nested-term-formula}
%%  \begin{align*}
%%   \exists R, S . \forall x, x' . & G_1(x) \wedge P_1(x,x') \rightarrow S(x',x') ~ \wedge \\
%%                                 & G_2(x') \wedge S(x,x') \wedge T(x',x'')\rightarrow S(x,x'') ~ \wedge \\
%% 				& G_1(x) \wedge P_1(x,x') \wedge \neg G_2(x') \wedge S(x',x'') \wedge 
%%                                 P_2(x'',x''') \wedge 
%%                                   R(x) > 0 \wedge R(x) > R(x''') 
%%  \end{align*} 
%% \end{definition}

%% \begin{definition}[Sequential Loops Non-Termination Formula {\bf [SNT]}]
%% \label{def:multi-termination-formula}
%%  \begin{align*}
%%   \exists R, Inv_1,..., Inv_n . \forall x_0,...,x_n, x_1',...,x_n'.  & P_0(x_0,x_1) \rightarrow Inv_1(x_1) ~  \\
%%  & \bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) ~  \\  
%%  & \bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) \rightarrow Inv_{i+1}(x_{i+1})) ~ \wedge \\
%%  & Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge N(x_n)  ~ \wedge \\
%%                                                         & N(x_n) \rightarrow G_n(x_n) ~ \wedge 
%% 							 N(x_n) \wedge T_n(x_n, x_n') \rightarrow N(x_n') 
%%  \end{align*}
%% \end{definition}
 \end{framed}

\end{figure*}

\subsection{Composing a Loop with its Environment} \label{sec:env}

Sometimes the termination behaviour of a loop depends on its environment, i.e. the rest of the program.  That is to say,
the loop may not terminate if started in some particular state, but the rest of the program
ensures that such a state cannot be reached on entry to the loop.  The program as a whole
terminates, but if the loop were considered in isolation we would not be able to prove that
it terminates. We must therefore encode a loop's interaction with its environment %the rest of the program 
in order to do a sound termination analysis.\\

\iffalse
Let us assume that we have done some preprocessing of our program which has identified
loops, straight line code blocks and the control flow between these.  In particular,
the control flow analysis has determined which order these code blocks execute in,
and the nesting structure of the loops.
\fi

\noindent {\bf Conditional termination.}
Given a loop $L=(G,T_o)$, where $T_o$ is an over-approximation of its transition relation, 
if $L$'s termination depends on the state it begins
executing in, we say that $L$ is \emph{conditionally terminating}.
The information we require of the environment is a predicate $I$ which
overapproximates the set of states that $L$ may begin executing in.
That is to say, for each state $x$ that is reachable on entry to $L$,
we have $I(x)$.

Conditional termination is then equivalent to the
formula {\bf [CT]} of Defintion~\ref{def:ct}.

\begin{theorem}
\label{thm:ct}
 The loop $L(G, T_o)$ terminates when started in any state satisfying $I(x)$ iff
 Defintion~\ref{def:ct} is satisfiable.
\end{theorem}

\begin{proof}
 Do this proof.
\end{proof}

If this formula is satisfiable, two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

$W$ is called a \emph{supporting invariant} for $L$ and $R$ proves termination relative to $W$.
We require that $I \wedge G$ is strong enough to establish the base case of $W$'s inductiveness.

Conditional termination is illustrated by the program in Figure~\ref{fig:motivation.h},
which is encoded as:
\begin{align*}
            I(\langle x, y \rangle) & \triangleq y = 1 \\
            G(\langle x, y \rangle) & \triangleq x > 0 \\
            T(\langle x, y \rangle, \langle x', y' \rangle) & \triangleq x' = x - y \wedge y' = y 
\end{align*}
If the initial states $I$ are ignored this loop cannot be shown to terminate, since any state with $y = 0$ and $x > 0$
would lead to a non-terminating execution.

However, formula {\bf [CT]} is satisfiable, as witnessed by the ranking function
$R(\langle x,y\rangle) = x$ and supporting invariant $W(\langle x, y \rangle ) \triangleq y = 1$.
This constitutes a proof that the program as a whole terminates, since the loop always begins
executing in a state that guarantees its termination.\\

%synthesizing environment abstractions: underapp for termination and underapp for non-termination.
% \cite{} addresses the problem of automatically underapproximating weakest preconditions.

%\noindent {\bf Nested loops.}
%% If $L$'s body contains another loop $L'$, we cannot directly apply either formula {\bf [UT]} or
%% {\bf [CT]} or prove that $L$ terminates.  This is because we don't have direct access to the
%% single step transition relation for $L$'s body.  Instead, we consider $L'$ to be part of
%% $L$'s environment, which we model with a relation $S$.  We require that $S$ overapproximates
%% $L'$ in the sense that it satisfies the following formula:
%% \begin{align*}
%%  \forall x, x'. G(x) \wedge W(x) \wedge T'^*(x, x') \wedge \lnot G'(x') \rightarrow S(x, x')
%% \end{align*}
%% Where $G$ and $G'$ are respectively $L$ and $L'$'s guards, and $T'^*$ is the reflexive transitive closure of
%% the inner loop ($L'$)'s transition relation.

%% To give an example, if we wish to prove that the outer loop in Figure~\ref{fig:nested} terminates,
%% it suffices to summarise the inner loop with the relation:
%% \begin{align*}
%% S(\langle x, y, z \rangle, \langle x', y', z' \rangle) & \triangleq x' = x \wedge y = 1
%% \end{align*}
%% Then the ranking function $R(x, y, z) = x$ proves termination of the outer loop $L(G, S)$.\\

%% \begin{figure}
%% \begin{lstlisting}
%% while (x > 0) {
%%   z = x, y = z + 1;
%%   while (z > 0) {
%%     z--, y--;
%%   }
%%   x -= y;
%% }
%% \end{lstlisting}
%% \caption{A nested loop\label{fig:nested}}
%% \end{figure}

\noindent {\bf Over-approximating the environment for termination.}
%% While proving that loop $L$ terminates, we create constraints on the environment in the
%% form of the predicate $I$ and the over-approximating summary relation $T_o$:
%% In the presence of nested loops, 
%% for each loop $L'$ nested in the body
%% of $L$, we must generate constraints saying that $T_o$ overapproximates $L'$, i.e. it is an over-approximation of the transitive closure of $L'$'s transition relation.
%% Similarly, for each loop $L''$ appearing before $L$ in the program's control flow, we
%% must add constraints saying that $I$ over-approximates the states reachable via $L''$.
%The details of how to create these constraints are standard, so
%rather than presenting the algorithm for generating the constraints, we direct the
%reader's attention to Figure~\ref{fig:environment-model}.  
%
%\noindent {\bf Environment abstractions.}
For a loop $L=(G,T)$, the formulae {\bf [CT]} %and {\bf [CNT]} 
from Definition~\ref{def:ct} 
%and Definition~\ref{def:cnt} 
is in terms of 
$T_o$ and $I$, which over-approximate the transition relation~$T$ and the set of states that $L$ may begin executing in, respectively. 
The reason for using over-approximations [instead of exact relations] %when formulating the (non-)termination of an isolated loop 
is so that we can use the same formulae regardless of the environment the loop is executing in, or 
the complexity of its body, e.g.~if it contains nested loops.

All we need to do in order to use {\bf [CT]} in any context %environment and regardless of the complexity of the loops's body 
is generate constraints over-approximating %abstracting 
the environment and the loop's body. %: for termination, we need over-approximating abstractions, whereas for non-termination 
%we need under-approximations. 
%
As the generation of such constraints is standard and covered by several
other
works~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08},
we will not provide the full algorithm, but rather illustrate it through the
example in Figure~\ref{fig:environment-model}.  Let us assume that we have
already proven the termination of loops $L_1$ and $L_3$, and we are now
trying to prove that $L_2$ terminates, which we encode as the conjunction of
constraints on the right-hand side of the figure:
%
\begin{itemize}
\item $W_1$, $W_2$ and $W_3$ are (over-approximating) inductive invariants for $L_1$, $L_2$ and $L_3$, respectively.
\item $S$ is a summary of $L_3$ that over-approximates the transitive closure of its transition relation.
\item If the formula is satisfiable, then $R$ is a ranking function for $L_2$.
 \end{itemize}

%the lines marked as $*$ over-approximate the states in which loop 2 can start execute.
%% contains a non-trivial program
%% along with the constraint system
%% encoding the termination of the outer loop.  Since the program contains sequential composition
%% of loops and nested loops, the constraint
%% system involves termination invariants $W_1, W_2, W_3$ and the inner loop summary $S$.

Note that despite its intricacy, the constraint system is still a formula in the synthesis fragment.
Since the outer loop of this program does indeed terminate, the system is satisfiable, as witnessed by:
\begin{align*}
W_1(\langle v, x, y, z \rangle ) & \triangleq x \leq y + 1 \\
W_2(\langle v, x, y, z \rangle ) & \triangleq x = y + 1 \\
S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) & \triangleq x ' = x \wedge y' = y \wedge \\
& z' + v' = x' \wedge v' \leq y' \\
R(x) & = x
\end{align*}

\begin{figure*}
\begin{framed}
\begin{minipage}{0.17\textwidth}
\begin{lstlisting}[mathescape=true]
x = 0;
$L_1:$
while(x <= y)
  x++;

$L_2:$
while(x > 0){
  z = x;
  v = 0;

$L_3:$
  while(v < y){
    z--;
    v++;
  }

  x -= z;
  y -= z;
}
\end{lstlisting}
\end{minipage}
\vline
\begin{minipage}{0.85\textwidth}
\begin{align*}
 \exists W_1, W_2, R, S . \forall v, x, y, z, v', x', y', z' . \\
   x = 0 & \rightarrow W_1(v, x, y, z) \, \wedge \\
   x \leq y \wedge W_1(v, x, y, z) & \rightarrow W_1(v, x+1, x, y, z) \, \wedge \\
   x > 0 \wedge x > y \wedge W_1(v, x, y, z) & \rightarrow W_2(v, x, y, z) \, \wedge \\
   %x > 0 \wedge W_2(v, x, y, z) & \rightarrow W_3(0, x, y, x) \, \wedge \\
   %v < y \wedge W_3(v, x, y, z) & \rightarrow W_3(v+1, x, y, z-1) \, \wedge \\
   %v \geq y \wedge W_3(v, x, y, z) & \rightarrow W_2(v, x-z, y-z, z) \, \wedge \\
   S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) \wedge v' < y' & \rightarrow S(\langle v, x, y, z \rangle, \langle v'+1, x', y', z'-1 \rangle) \, \wedge \\
   x > 0 \wedge W_2(v, x, y, z) & \rightarrow S(\langle v, x, y, z \rangle, \langle 0, x, y, x \rangle) \, \wedge \\
   x > 0 \wedge W_2(v, x, y, z) \wedge S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) \wedge v' \geq y' & \rightarrow \\
   & W_2(v', x' - z', y' - z', z') \, \wedge \\
   & R(v, x, y, z) > 0 \, \wedge \\
   & R(v, x, y, z) > R(v', x' - z', y'-z', z')
\end{align*}
\end{minipage}
\end{framed}
\caption{A non-trivial program and its constraint system\label{fig:environment-model}}
\end{figure*}

\noindent {\bf Under-approximating the environment for non-termination.}
Dual to termination, when proving non-termination, we need to under-approximate the environment and the loop's body and apply
formula {\bf [CNT]}. %As opposed to over-approximations 
%this is not as straightforward as when computing over-approximations in the termination case described above.
Recall that there is one more constraint that we must obey when doing so: we must stay inside the synthesis fragment! 
This becomes more challenging now then when dealing with over-approximations due to the existential nature of under-approximations: they require an additional existential quantifier, which might result in an additional quantifier alternation, i.e. $\exists \forall \exists$.

This issue does not arise when abstracting the environment as its under-approximation and the loop under consideration make use of existential quantification at the same nestedness level. However, when dealing with nested loops, we need to compute the under-approximation of the inner loop, which requires a nested existential quantifier, 
resulting in the $\exists \forall \exists$ alternation. 
%
%% computing the under-approximation of the loop's body requires an extra quantifier alternation. 
%% The origin of the problem comes from the fact that an under-approximation is an existential property and therefore
%% requires an extra existential quantifier, resulting in the $\exists \forall \exists$ alternation. This situation is similar to the one encountered when proving non-termination of non-deterministic loops. 
%
To address this limitation, we treat nested loops as a special case and give
a separate formula for them in Figure~\ref{fig:nonterm-nested}.  If the
formula on the right of the figure is satisfiable, then $L_1$ is
non-terminating as witnessed by the recurrence set $N_1$ and the initial
state $x_0$ in which the program begins executing.  There are two possible
scenarios for $L_2$'s termination:
%
\begin{itemize}

\item If $L_2$ is terminating, then $N_2$ is an inductive invariant that
reestablished $N_1$ after $L_2$ stops executing: $\lnot G_2(x) \wedge N_2(x)
\wedge P_2(x,x') \rightarrow N_1(x') $.

\item If $L_2$ is non-terminating, then $N_2$ is its recurrence set.

\end{itemize}

%% The issue appears when we try to under-approximate the body of the loop. 
%% Here we encounter a similar problem with the one for non-determinism mentioned above.

%% Consider the generic example in Figure~\ref{fig:nonterm-environment}, where we are attempting to prove non-termination
%% of the innermost loop $L_3$ by finding a closed recurrence set $N_3$. 
%% -- We will assume that the other two loops iterate only a finite number of times.

%% We now explain the constraints on the right of the figure.
%% Given that we prove non-termination, we try to  under-approximate the states from which $L_3$ can start executing. 
%% We do this in several steps:

%% $W_1$, $W_2$ and $W_3$ are inductive invariants corresponding to $L_1$, $L_2$ and $L_3$, respectively, such that:

%% - $W_1$ is an inductive invariant of $L_1$ that is strong enough to establish $W_2$ if $L_1$ exits (which it does by assumption).
%% - $W_2$ is an inductive invariant of $L_2$ that is strong enough to establish either $W_3$ or $N_3$ on entry to $L_3$, depending on whether $L_3$ terminates or not.
%% - Recall that $N_3$ is a recurrence set for $L_3$, so if a state $x \in N_3$ is reached, the loop $L_3$ will execute forever. Otherwise, $W_3$ holds on entry to $L_3$.
%% - $W_3$ is an inductive invariant of $L_3$ that is strong enough to re-establish $W_2$ when $L_3$ exits.

%% If the non-termination formula is satisfiable, $x_0$ is some initial state in which the program begins executing.

%% First, we under-approximate the states in which $L_2$ can start executing through $W_2$.
%% Then $N_1$ is an inductive invariant of $L_1$ that is strong
%% enough to establish $N_2$ if $L_1$ exits (which it does by assumption).  Now, $N_2$ is an inductive invariant
%% of $L_2$ that is strong enough to establish either $N_3$ or $N_4$ on entry to $L_3$.  Recall that $N_4$ is
%% a recurrence set for $L_3$, so if a state $x \in N_4$ is reached, the loop $L_3$ will execute forever.
%% Otherwise, $N_3$ holds on entry to $L_3$.  $N_3$ is an inductive invariant of $L_3$ that is strong enough to
%% re-establish $N_2$ when $L_3$ exits.

%% If the constraint system is satisfiable, $x_0$ is some initial
%% state in which the program begins executing.

%%   Since we assumed that $L_2$ iterated only a finite number of times,
%% $N_3$ must be established only a finite number of times, which means that $N_4$ will eventually be reached,
%% at which point $L_3$ will loop forever since $N_4$ is a recurrence set for $L_3$.

%% \begin{figure*}
%% \begin{framed}
%%  \begin{minipage}{0.19\textwidth}
%% \begin{lstlisting}
%% L1: while (c1) {
%%      B1;
%%     }

%% L2: while (c2) {
%%      P;

%% L3:  while (c3) {
%%        B3;
%%      }
%%     }
%% \end{lstlisting}
%% \end{minipage}
%% \vline
%% \begin{minipage}{0.82\textwidth}
%% \begin{align*}
%%  \exists W_1, W_2, W_3, N_3, x_0 . \forall x, x' . \\
%%   W_1(x_0) \, \wedge \\
%%   c_1(x) \wedge W_1(x) \wedge B_1(x, x') & \rightarrow W_1(x') \, \wedge \\
%%   \lnot c_1(x) \wedge W_1(x) & \rightarrow W_2(x) \, \wedge \\
%%   W_2(x) & \rightarrow c_2(x) \, \wedge \\
%%   W_2(x) \wedge P(x, x') & \rightarrow W_3(x') \vee N_3(x') \, \wedge \\
%%   c_3(x) \wedge W_3(x) \wedge B_3(x, x') & \rightarrow W_3(x') \, \wedge \\
%%   \lnot c_3(x) \wedge W_3(x) & \rightarrow W_2(x) \, \wedge \\
%%   N_3(x) & \rightarrow c_3(x) \, \wedge \\
%%   N_3(x) \wedge B_3(x, x') & \rightarrow N_3(x')
%% \end{align*}
%% \end{minipage}
%% \end{framed}

%% \caption{The constraint system for non-termination\label{fig:nonterm-environment}}
%% \end{figure*}

\begin{figure*}
\begin{framed}
 \begin{minipage}{0.19\textwidth}
\begin{lstlisting}[mathescape=true]
$L_1$:
while ($G_1$) {
  $P_1$;

$L_2$:
  while ($G_2$) {
    $B_2$;
  }

  $P_2$;
}
\end{lstlisting}
\end{minipage}
\vline
\begin{minipage}{0.82\textwidth}
\begin{align*}
 \exists N_1, N_2, x_0 . \forall x, x' . \\
  N_1(x_0) & \, \wedge \\
  N_1(x) & \rightarrow G_1(x) \, \wedge \\
  N_1(x) \wedge P_1(x,x') & \rightarrow N_2(x') \, \wedge \\
  G_2(x) \wedge N_2(x) \wedge B_2(x,x') & \rightarrow N_2(x') \, \wedge \\
  \lnot G_2(x) \wedge N_2(x) \wedge P_2(x,x') & \rightarrow N_1(x') 
\end{align*}
\end{minipage}
\end{framed}

\caption{The non-termination formula for nested loops \label{fig:nonterm-nested}}
\end{figure*}

\iffalse

\noindent {\bf Sequential loops.}
Conditional termination allows us to
prove that programs with multiple loops terminate, even if the termination of some loop depends on the states
reachable after leaving a previous loop.
%The conditional termination formula becomes more complex in the presence of multiple loops.
For illustration, consider the following program with $n$ sequential loops $\forall i=1..n. ~L_i=(x_i,G_i,T_i)$ and
the assertions $\forall i=1..n. ~P_i$ modelling the program statements in between loops.

\begin{lstlisting}[mathescape=true]
$P_1$
while ($G_1$) { $T_1$ }
$P_2$

while ($G_2$) { $T_2$ }
...
$P_n$
while ($G_n$) { $T_n$ }
\end{lstlisting}

In order to prove that the $n^{th}$ loop terminates, we need to apply the {\bf [CT]} formula for $L_n$.
The challenge is determining the initial states  $I(x_n)$ at the loop's entry point.  
The termination problem for $L_n$ is thus equivalent to the satisfiability of
Definition~\ref{def:multi-termination-formula}.  If this formula is satisfiable:
\begin{itemize}
\item $Inv_i$ is an inductive invariant of $L_i$ for $i=1..{n-1}$ as established by the conjuncts:
$$\bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) $$

\item Each loop invariant $Inv_i$ with $i=1..{n-2}$ is strong enough to establish $Inv_{i+1}$:
$$\bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) {\rightarrow} Inv_{i+1}(x_{i+1})) ~$$

\item When $L_n$'s guard holds, $Inv_{n-1}$ 
is strong enough to establish $Inv_n$, i.e. $Inv_n$ holds for the program states after $L_n$'s entry point:
$$ Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge G_n(x_n) $$
$$\qquad\qquad\rightarrow Inv_n(x_n)$$

\item $Inv_n$ is a supporting termination invariant of $L_n$ and $R$ is a ranking function relative to $Inv_n$:
$$Inv_n(x_n) \wedge G_n(x_n) \wedge T(x_n, x_n') \rightarrow Inv_n(x_n')$$
$$ \qquad \qquad \wedge R(x_n) > 0 \wedge R(x_n) > R(x_n')$$
\end{itemize}

\todo{add example?}\\

\noindent{\bf Nested loops.}
%
We illustrate our approach to handling nested loops for only two loops for
sake of brevity, but the approach can be easily generalised to any number of
nested loops:
%
\begin{lstlisting}[mathescape=true]
while ($G_1$) { 
  $P_1$ 
  while ($G_2$) { $T$ }
  $P_2$
}
\end{lstlisting}

The termination of the outer loop is equivalent to the formula in
Definition~\ref{def:nested-term-formula}.  The formula requires an auxiliary
assertion $S(x,x')$ representing a binary relation between the initial
states at the inner loop's entry point and their successors according to the
transition relation $T$.  If this formula is satisfiable, then $R$ is a
ranking function of the outer loop.

In order to find a termination argument for the inner loop, we use {\bf [CT]} where the initial states are given by 
$\neg G_1(x) \wedge P_1(x,x') \rightarrow I(x')$.

\noindent{\bf Sequential loops.}
Discuss {\bf [SNT]}.\\

\noindent{\bf Nested loops.}
\fi

\subsection{Generalised synthesis formula.}
Our design: for each loop always search for lexicographic ranking function + supporting invariant OR recurrence set.

Note: given a program with multiple loops (sequential and nested), we could potentially compute termination/non-termination 
arguments for all of them in the same formula. Currently, our tool generates a separate formula for each loop.

- when analysing a loop, we assume that the nested ones and the one preceding it in program order terminate.\\
%% \subsection{Program Synthesis as Second-Order Satisfaction}

%% With the exception of Definition~\ref{def:nonterm-formula}, each of the second-order formulae above
%% belongs to the synthesis fragment.  We use this name because the program synthesis problem can be
%% described as finding a satisfying assigment to the following formula, which is the general case of
%% the synthesis fragment:

%% \begin{definition}[Synthesis Formula]
%%  \label{def:synth-formula}
%%  $$\exists P, x . \forall y . \sigma(P, x, y)$$
%% \end{definition}

\section{Satisfiability of a Formula in the Synthesis Fragment} \label{sec:synthesis}
%
\iffalse
\subsection{The Search Space and the Kolmogorov Complexity of a Ranking Function}
\begin{conjecture}
The lexicographic ordering does not influence the size of the search space. Thus, finding potentially lexicographic termination arguments is not more difficult than finding non-lexicographic ones.   
\end{conjecture}

As a consequence of the aforementioned conjecture, we do not need to commit
to a specific form of the termination argument.  While other approaches,
e.g.  \cite{DBLP:conf/tacas/LeikeH14}, conduct independent searches for each
possible form of the ranking function with some of them futile whenever the
assumed constrained ranking function does not exist, we have a single search
that aims at finding the ranking function with the lowest Kolmogorov
complexity.
%
\fi
%
%

%\todo{add related work for synthesis.}

In this section we present a solver for the synthesis fragment:
 \[
  \exists P,~ x . ~\forall~ y . \sigma(P, x, y)
 \]
%
where $P$ ranges over sets, while $x$ and $y$ range over ground terms (as
mentioned before, we drop the vector notation and write $x$ for $\vec{x}$).

By viewing $\sigma: (X^n \times Y^m \to Z^k) \times X^n \times Y^m  \to
\mathbb{B}$ as a specification function that returns true iff $P$ computes
appropriate outputs for the inputs $x$ and $y$, we can regard the
satisfaction problem for the synthesis fragment as \emph{program synthesis},
i.e.  the mechanised construction of software that provably satisfies a
given specification.  Research in the area of program synthesis has been
particularly fruitful beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}, and
continuing with works such as {\sc Brahma}~\cite{brahma} and program
sketching~\cite{lezama-thesis,sketch,modular-sketch}.

% 
%We can solve the synthesis problem using any off-the-shelf program synthesiser.  If you don't
%have a program synthesiser kicking around, we present the design of one synthesiser
%that can be used to solve the problem.
%We start by recalling the synthesis formula:
% $$\exists P, x . \forall y . \sigma(P, x, y)$$

%% Our program encoding generates an $\exists\forall$ formula that is \emph{linear} in the length of the shortest program
%% satisfying the user's specification.  This formula is then checked for satisfiability
%% using the CEGIS algorithm.

In the context of (non-)termination analysis, a witnesses to the
satisfiability of a formula in the synthesis fragment is a
\mbox{(non-)termination} proof.  As mentioned earlier in the paper, our
analysis must ensure precise treatment of bit-wise operations and arithmetic
modulo fixed widths by modelling bit-accurate semantics.  The synthesiser
must therefore be able to compute (non-termination) proofs over machine
integers and floating-point numbers from the specifications detailed in
Section~\ref{sec:second.order}.

%\subsection{\newC}
%\label{sec:logic}

\subsection{The Synthesis Specification}

The logic we use to express the synthesis formula, i.e.~the
\mbox{(non-)termination} specifications, is a subset of C that we call
\newC.  The characteristic property of a \newC program is that safety can be
decided for it using a single query to a Bounded Model Checker.  A~\newC
program is just a C program with the following syntactic restrictions:
 all loops in the program must have a constant bound;
 all recursion in the program must be limited to a constant depth;
 all arrays must be statically allocated (i.e. not using \texttt{malloc}),
 and be of constant size.
Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.

The (non-)termination proofs are arbitrary expressions over the grammar given in Figure~\ref{fig:l-language}.  
%written in a simple RISC-like language $\mathcal{L}$, whose syntax is given in Fig.~\ref{fig:l-language}.  
\todo{update the figure and make it look more like a grammar.}

\begin{algorithm*}
 \caption{Abstract refinement algorithm
 \label{alg:cegis}}
 \vspace{-1.5em}
 \begin{multicols}{2}
 \begin{algorithmic}[1]
 \Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\columnbreak
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{\Call{synth}{inputs}}
    \If{candidate = UNSAT}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{\Call{verif}{candidate}}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
 \end{multicols}
\end{algorithm*}

\begin{figure}
 \begin{center}
 \resizebox{\linewidth}{!}{%
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,
 semithick, initial text=]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=1cm, minimum width=1.5cm},
          row sep=2cm, column sep=2.5cm, ampersand replacement=\&] {
   \node (synth) {Synthesise};
   \&
   \node (verif) {Verify}; %\\
   %\node[draw=none] {};
   \&
   \node[ellipse] (done) {Done}; \\
  };

   \path
    (synth) edge [bend left] node {Candidate program} (verif)
    (verif) edge [bend left] node {Counterexample input} (synth)
    (verif) edge node {Valid} (done);
 \end{tikzpicture}
 }
 \end{center}

 \caption{Abstract synthesis refinement loop
 \label{fig:abstract-refinement}}
\end{figure}

\begin{figure}
{\small
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

\begin{tabular}{llll}
 \verb|add a b| & \verb|sub a b| & \verb|mul a b| & \verb|div a b| \\
 \verb|neg a| &   \verb|mod a b| & \verb|min a b| & \verb|max a b|
\end{tabular}

\medskip

Bitwise logical and shift instructions:

\begin{tabular}{lll}
 \verb|and  a b| & \verb|or   a b| & \verb|xor a b| \\
 \verb|lshr a b| & \verb|ashr a b| & \verb|not a|
\end{tabular}

\medskip

Unsigned and signed comparison instructions:

\begin{tabular}{lll}
 \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| \\
 \verb|slt a b| & \verb|eq  a b| & \verb|neq  a b|
\end{tabular}

Miscellaneous logical instructions:

\begin{tabular}{lll}
 \verb|implies a b| & \verb|ite a b c| & 
\end{tabular}

Floating-point arithmetic:

\begin{tabular}{llll}
 \verb|fadd a b| & \verb|fsub a b| & \verb|fmul a b| & \verb|fdiv a b| 
\end{tabular}

\end{center}
}
 \caption{The language $\mathcal{L}$}
 \label{fig:l-language}
\end{figure}

\subsection{The Synthesis Algorithm}
As safety of a \newC program is efficiently 
%the quantifier free first-order fragment of the logic is efficiently 
decidable and a termination proof is expressible as a ground term of the \newC logic, we use
Counterexample Guided Inductive Synthesis (CEGIS)~\cite{lezama-thesis,sketch} to
check the validity of the synthesis formula. 
The core of the CEGIS algorithm is the refinement loop shown in Fig.~\ref{fig:abstract-refinement} and
detailed in Algorithm~\ref{alg:cegis}.  
The algorithm is divided into two
procedures: {\sc synth} (see Figure~\ref{fig:synth-dfd}) and {\sc verif}, which interact via
a finite set of test vectors {\sc inputs}, which is initialised to $\emptyset$.

The {\sc synth} procedure tries to find existential witnesses $(P, x_0)$
that satisfy the partial specification:
%
\[
 \exists P, x_0 . \forall x \in \text{\sc inputs} . \sigma(P, x_0, x)
\]

If {\sc synth} succeeds in finding a witness $(P, x_0)$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verif} which determines whether the candidate does satisfy
the specification on all inputs by checking satisfiability of the
verification formula:
%
\[
 \exists x . \lnot \sigma(P, x_0, x)
\]

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set and the loop iterates again.

Concretely, {\sc synth} is implemented as shown in
Figure~\ref{fig:synth-dfd}.  We start by taking the termination
specification and generating a C program which takes as input a candidate
$(P, x_0)$ and asserts that the candidate fails to meet the specification on
at least on of the elements of {\sc inputs}.  So this program is unsafe iff
there is some candidate $(P, x_0)$ satisfying the specification for all the
elements of {\sc inputs}.  Finding a new candidate solution is therefore
reduced to model checking this C program, which by construction is loop
free.  The model checking is done using a combination of bounded model
checking, explicit state enumeration and genetic programming (GP).  The
latter two techniques involve combining the previously generated C program,
along with code that searches for candidate solutions.  Finally, if any of
the model checkers finds a candidate solution it returns it.  The form of
this candidate solution is yet another program, this time written in the
proof language $\mathcal{L}$.

As well as efficiency, compiling the original source code of the program
under analysis has the advantage of guaranteeing that we are verifying the
exact semantics of the program as understood by a compiler, rather than some
ad-hoc abstraction of its semantics.

Similarly to {\sc synth}, the {\sc verif} procedure is implemented by
creating a C program that is unsafe iff some input exists on which the
candidate solution fails to meet the specification.  Again, symbolic and
explicit state model checking are used to decide the satisfiability of the
verification formula.

%If the input set is finite, this procedure is guaranteed to terminate.

%% In some cases, it is faster to use an explicit-state model checker rather than a Bounded Model Checker.  
%% This is particularly true for the {\sc verif} component, 
%% where we have observed that incorrect programs tend
%% to be incorrect on a large fraction of the input space.  Counterexamples
%% are then very easy to find by explicit enumeration of a few inputs.
%% Since \newC is a fragment of C, we can generate an explicit-state
%% model checker using the same source files that we pass to {\sc cbmc}
%% and adding a small function to enumerate possible inputs.

%% ---  We begin with the program under analysis (the PUA).  From this we produce a specification (in
%% second-order logic)
%% encoding the termination of the PUA.  From the logical specification, we produce a program SYNTH.
%% SYNTH takes as input a proof and asserts that the proof is incorrect on at least one of the inputs being tracked so far.
%% Therefore, SYNTH is unsafe iff there is some proof that is correct for all of the tracked inputs.  Now we use testing
%% techniques to try to find a proof that causes SYNTH to crash.  The testing techniques we use are: model checking,
%% exhaustive search and genetic programming.  For the latter 2 techniques, we combine SYNTH with another program
%% (GP or EXPLICIT respectively) to create a program that searches for test vectors (proofs) that cause SYNTH
%% to crash.

%% In the {\sc synth} component, in order to find a candidate proof $P$ satisfying
%% $\sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$
%% for the tracked inputs $(i_1, \ldots, i_N)$, 
%% we attempt a safety proof.
%In order to determine the validity of the {\sc synth} formula, we can
%check the {\sc synth} program for safety.

\begin{figure*}
\begin{center}
\tikzstyle{file}=[draw, text width=7.0em, text centered,
  minimum height=1.5em]
\tikzstyle{process} = [draw, minimum height=3em, circle]
\tikzstyle{line} = [draw, color=black, -latex']

\def\Divide#1#2{%
 \coordinate(a) at ($(#1.east) !.5! (#2.west)$);
 \coordinate(b) at (a |- 0,-3);
 \draw[dotted] (b) -- ++(0, 4.5);
}

%\resizebox{\linewidth}{!}{
\begin{tikzpicture}[font=\sffamily]

\node [file] (pua) {input.c};

\path (pua.east)+(2,0) node [file] (spec) {\sc Termination Specification};
\path (spec.south)+(0.0, -1) node [file] (inputs) {\sc Tracked Inputs};
%% \path (synth.south)+(0.0, -0.5) node [file] (tests) {\sc tests.c};
%% \path (tests.south)+(0.0, -0.5) node [file] (interpreter) {\sc interpreter.c};
%% \path (interpreter.south)+(0.0, -0.5) node [file] (spec) {\sc spec.c};

%% \path (spec.east)+(2.0, -0.25) node [process] (merged) {merge};
\path (spec.east)+(2.0, -0.75) node [file] (file) {synth.c};

\path (file.east)+(2.0, 1.5) node [process] (cbmc) {\sc cbmc};
\path (file.east)+(2.0, 0.0) node [process] (gcc) {\sc Search};
\path (file.east)+(2.0, -1.5) node [process] (gp) {\sc GP};

\path (gcc.east)+(2.5, 0.0) node [file] (out) {Candidate Proof};

%\path [line] (spec) -- (merged);

\path [line] (pua) -- (spec);

\path [line] (spec) -- (file);
\path [line] (inputs) -- (file);

\path [line] (file) -- (cbmc);
\path [line] (file) -- (gcc);
\path [line] (file) -- (gp);

\path [line] (cbmc) -- (out);
\path [line] (gcc) -- (out);
\path [line] (gp) -- (out);

\Divide{pua}{spec}{puaspec}
%\Divide{spec}{file}{specfile}
\Divide{file}{gcc}{filegcc}
\Divide{gcc}{out}{gccout}

\draw (pua |- 0,-3) node [align=center] {Input program \\ (written in C)};

\coordinate (midspec) at ($(spec) !.5! (file)$);
\draw (midspec |- 0,-3) node [align=center] {Automatically generated specification \\ (written in C)};
%\draw (file |- 0,-3) node {Spec};
\draw (gcc |- 0,-3) node [align=center] {Model checker};
\draw (out |- 0,-3) node [align=center] {Proof \\ (written in $\mathcal{L}$)};

\end{tikzpicture}
%}
\end{center}

\caption{Schematic diagram of {\sc synth}}
\label{fig:synth-dfd}
\end{figure*}

\iffalse
\subsection{Parameterising the Program Space}

In order to search the space of candidate programs, we parametrise
the language~$\mathcal{L}$ by program length, word width and number of constants,
inducing a lattice of progressively
more expressive languages.  We start by attempting to synthesise
a program at the lowest point on this lattice and increase the
parameters of~$\mathcal{L}$ until we reach a point at which
the synthesis succeeds.

As well as giving us an automatic search procedure, this parametrisation
greatly increases the efficiency of our system since languages
low down the lattice are very easy to decide safety for.  If a program
can be synthesised in a low-complexity language, the whole procedure
finishes much faster than if synthesis had been attempted in a
high-complexity language.

We introduce a novel parametrisation of the programming language
used to express our synthesised programs.
This parametrisation allows us to efficiently explore the program space
without relying on human guidance and also ensures that our programs
are of minimal length.

Our tool is the first we are aware of that is able to effectively
synthesise floating-point programs. We demonstrate this by
synthesising {\sc Fast2Sum} using Knuth's {\sc 2Sum}~\cite{taocp2} as
a specification.

---Our exploration of the program space ensures that our programs are minimal in size.
--- We parametrise the space of programs in such a way that it can be explored automatically, rather than asking a human for hints.

We consider the following parameters:
\begin{itemize}
\item{Program Length: $l$}
The first parameter we introduce is program length, denoted by $l$.
At each iteration we synthesise programs of length exactly $l$.
We start with $l = 1$ and increment $l$ whenever we determine
that no program of length $l$ can satisfy the specification.  When we do
successfully synthesise a program, we are \emph{guaranteed that it
is of minimal length} since we have previously established that no
shorter program is correct.
\item{Word Width: $w$}
An $\mathcal{L}$-program runs on a virtual machine (the $\mathcal{L}$-machine) that
has its own set of parameters.  The only relevant parameter is
the \emph{word width} of the $\mathcal{L}$-machine, that is, the number of bits
in each internal register and immediate constant.  This parameter is denoted by
$w$.  The size of the final SAT problem generated by {\sc cbmc} scales
polynomially with $w$, since each intermediate C variable corresponds
to $w$ propositional variables.

\item{Number of Constants: $c$}
Instructions in $\mathcal{L}$ take either one or two operands.
Since any instruction whose operands are all constants can always be
eliminated (since its result is a constant), we know that a loop-free program
of minimal length will not contain any instructions with two constant
operands.  Therefore the number of constants that can appear in
a minimal program of length $l$ is at most $l$.  By minimising the number
of constants appearing in a program, we are able to use a particularly
efficient program encoding that speeds up the synthesis procedure
substantially.  The number of constants used in a program is the parameter $c$.

\end{itemize}
\subsubsection{Searching the Program Space}

The key to our automation approach is to come up with a sensible way in which to
adjust the $\mathcal{L}$-parameters in order to cover all possible programs.
After each round of {\sc synth}, we may need to adjust the parameters.  
The
logic for these adjustments is shown as a tree in Fig.~\ref{fig:paramsflow}.
Whenever {\sc synth} fails, we consider which parameter might have caused the
failure.  
There are two possibilities: either the program length $l$ was too small,
or the number of allowed constants $c$ was.  If $c < l$, we just increment $c$ and
try another round of synthesis, but allowing ourselves an extra program constant.
If $c = l$, there is no point in increasing $c$ any further.  This is because
no minimal $\mathcal{L}$-program has $c > l$, for if it did there would
have to be at least one instruction with two constant operands.  This
instruction could be removed (at the expense of adding its result as
a constant), contradicting the assumed minimality of the program.  So
if $c = l$, we set $c$ to 0 and increment $l$, before attempting
synthesis again.

If {\sc synth} succeeds but {\sc verif} fails, we have a candidate
program that is correct for some inputs but incorrect on at least
one input.  However, it may be the case that the candidate program
is correct for \emph{all} inputs when run on an $\mathcal{L}$-machine
with a small word size.  For example, we may have synthesised a
program which is correct for all 8-bit inputs, but incorrect for
some 32-bit input.  If this is the case (which we can determine
by running the candidate program through {\sc verif} using the smaller
word size), we may be able to produce a correct program for
the full $\mathcal{L}$-machine by using the constant extension rules
shown in Fig.~\ref{fig:generalize}.  If constant generalization
is able to find a correct program, we are done.  Otherwise,
we need to increase the word width of the $\mathcal{L}$-machine
we are currently synthesising for.

\section{Conclusion}

By expressing the program synthesis problem as a safety property for a
program interpreter, we have been able to harness the power of
state-of-the-art program analysis tools and reuse them in a new problem
domain.  We have implemented our algorithm as a freely downloadable tool
whose performance compares favourably to a recent program synthesiser. 
Finally, we have taken advantage of the expressiveness of our specification
language to make an initial step towards practical synthesis of
floating-point programs.

\fi

\subsection{Candidate Generation Strategies}
\noindent{\bf Explicit Proof Search} The simplest strategy for finding candidates
is to just exhaustively enumerate them all, starting with the shortest and
progressively increasing the number of instructions.  Since the set of
$\mathcal{L}$-programs is recursively enumerable, this procedure is complete.
\\

\noindent{\bf Symbolic Bounded Model Checking} Another complete method for generating
candidates is to simply use BMC on the {\sc synth.c} program.  As with explicit
search, we must progressively increase the length of the solution we search for
in order to get a complete search procedure.
\\

\noindent {\bf Genetic Programming and Incremental Evolution} \label{sec:gp}
To generate candidate (non-)termination proofs in the ``Synthesise'' phase
of the refinement loop, we make use of genetic programming
(GP)~\cite{langdon:fogp} as an alternative to model checking and exhaustive
search.  Genetic programming provides an intelligent and adaptive way of
searching through the space of computer programs for an individual that is
highly fit in solving the problem at hand by using the Darwinian principle
of survival and reproduction of the fittest.  Thus, the program that emerges
from the genetic programming paradigm is a consequence of fitness.

We begin with a population of random programs, then iteratively evolve this
population by applying the genetic operators {\sc crossover} and {\sc
mutate}.  The {\sc crossover} operator takes two programs and combines their
code in some way to create a new program.  The {\sc mutate} operator takes a
single program and randomly changes parts of its code, again creating a new
program.  Both these operators are applied to programs selected from the
population in proportion to their observed fitness, and a program is said to
be fitter than another if it meets the specification on more test cases.

In the setting of program synthesis, genetic programming can be very
usefully combined with the concept of incremental evolution.  The idea is
that in order to evolve a program that is adequate for some difficult task,
it is often helpful to first evolve a population that as an ensemble is good
at some simpler variation of the task.  This principle leads us to the
observation that each round of the CEGIS loop can be thought of as creating
a more difficult problem for the GP synthesiser to solve.  Therefore, if GP
finds a candidate program which is correct for the $k$ inputs so far
observed, but is not correct for the full specification, that candidate
program came from a population that was ``good'' for the specification
restrict to those $k$ inputs.  The next time round the CEGIS loop, the
problem GP must solve is slightly harder -- any solution to this new problem
must be correct for the previous $k$ inputs and also whatever new input was
added by the previous iteration of {\sc verif}.  It therefore makes sense to
begin the $k+1$st round of evolution with the population at the end of the
$k$-th round.

%% The general idea is to start with a population
%% of random programs and evolve a good one.  ``Good'' in this case means that the program
%% satisfies our specification.  To do this evolution, we compute a fitness for each program
%% and repeatedly apply genetic operators to the population based on this fitness.  The operators
%% used are {\sc crossover} and {\sc mutate}.  The crossover operator takes two programs
%% and combines their code in some what to create a new program.  The mutate operator takes a single
%% program and randomly changes parts of its code, again creating a new program.  We arrange
%% things so that the fitter a program is, the more likely it is to be selected as a parent
%% that will be crossover'd to produce a child in the next generation.  Our fitness function
%% is just the number of test cases on which the program met the specification.

\iffalse
We begin by observing that the asymptotic complexity of all of our synthesis
backends are equal, assuming $P \neq NP$.  This complexity is:

$$O\left(2^{K(f)}\right)$$

Where $K(f)$ is the Kolmogorov complexity of $f$, which is $O(\log Y^X) = O(X)$
so the complexity is doubly exponential in the width of $X$.

\begin{theorem}
 Fitness landscapes form a lattice.  Adding test vectors corresponds to abstraction refinement on this
 lattice, which is why incremental GP works well.
\end{theorem}

\begin{proof}
 Trivial.
\end{proof}

\begin{conjecture}
 A single fitness landscape isn't really very smooth (e.g. small changes in program representation
 don't correspond to small changes in fitness), so GP probably shouldn't work very well.

 But it does.
\end{conjecture}
\fi

\section{Soundness, Completeness and Complexity}

Termination for bitvector programs is known to be
PSPACE-complete~\cite{DBLP:conf/tacas/CookKRW10}.  We first record the
expected results for soundness and completeness of our algorithm for
deciding it.
%
\begin{theorem}\label{thm:synth-sound}
Algorithm~\ref{alg:cegis} is sound -- if it terminates with witness $P$, then
$P \models \sigma$.
\end{theorem}

\begin{proof}
 The procedure {\sc RefinementLoop} terminates only if {\sc Verif} returns ``valid''.  In that
 case, $\exists x . \lnot \sigma(x, P)$ is unsatisfiable and so $\forall x . \sigma(x, P)$ holds.
\end{proof}

\begin{theorem}
 \label{thm:synth-semicomplete}
 If the existential first-order theory used to express the specification $\sigma$ is decidable,
 algorithm~\ref{alg:cegis} is semi-complete -- if a program $P \models \sigma$
 exists then algorithm~\ref{alg:cegis} will terminate.  However, if no program
 satisfies the specification, the algorithm may not terminate.  Furthermore if the domain of inputs $X$
 is finite, the algorithm is guaranteed to terminate.
\end{theorem}

\begin{proof}
 Since the {\sc ExplicitSearch} routine enumerates all programs (as can be seen by induction on
 the program length $l$), it will eventually enumerate a program that meets the specification
 in whatever set of inputs are currently being tracked.  Since the first-order theory is
 decidable, the query in {\sc Verif} will succeed for this program, causing the algorithm to terminate.
 The set of correct programs is therefore recursively enumerable and algorithm~\ref{alg:cegis}
 enumerates this set, so it is semi-complete.

 If the domain $X$ is finite then the loop in procedure {\sc RefinementLoop} can only
 iterate $\| X \|$ times, since by this time all of the elements of $X$ would have been
 added to the inputs set.  Therefore if the {\sc Synth} procedure always terminates (which it does
 by assumption of decidability), the whole algorithm terminates.
\end{proof}

\begin{theorem}
\label{thm:sound}
 {\sc Headshot} is sound -- if it returns a verdict of termination, the program under
 analysis terminates on all inputs.  Similarly if it returns a verdict of non-termination,
 the program under analysis has at least one infinite execution.
\end{theorem}

\begin{proof}
 This is a direct consequence of Theorems~\ref{thm:ut}, \ref{thm:ct}, \ref{thm:ont}, \ref{thm:cnt} and \ref{thm:synth-sound}.
\end{proof}

\begin{theorem}
\label{thm:complete-termination}
 {\sc Headshot} is complete for terminating programs -- if a program terminates, {\sc Headshot} will find a proof that it does.
\end{theorem}

\begin{proof}
 If the program under analysis $P$ has variables $v_1, \ldots, v_k$ each of which is $b$ bits wide, its state space $\mathcal{S}$ is of size $2^{bk}$.
 A ranking function $R: \mathcal{S} \to \mathcal{D}$ for $P$ exists iff $P$ terminates.  Since $R$ is injective, we have that
 $\| \mathcal{D} \| \geq \| \mathcal{S} \|$.  If $\| \mathcal{D} \| > \| \mathcal{S} \|$, we can construct a function $R': \mathcal{S} \to \mathcal{D'}$
 with $ \| \mathcal{D'} = \| \mathcal{S} \|$ by just setting $R' = R|_\mathcal{S}$, i.e. $R'$ is just the restriction of $R$ to $\mathcal{S}$
 Since $\mathcal{S}$ already comes equipped with a natural well ordering we can also construct $R'' = \iota \circ R'$
 where $\iota: \mathcal{D'} \to \mathcal{S}$ is the unique order isomorphism from $\mathcal{D'}$ to $\mathcal{S}$.
 So assuming that $P$ terminates, there is some ranking function $R''$ that is just a permutation of $\mathcal{S}$.
 If the number of variables $k > 1$ then in general the ranking function will be lexicographic with dimension $\leq k$
 and each co-ordinate of the output being a single $b$-bit value.

 It is fairly easy to see that any finite permutation is computed by a finite program in the proof language generated
 by {\sc Kalashnikov}.  A very inefficient construction which computes the permutation $R(x) = x+1$ for one variable
 is:
 \begin{verbatim}
t1 = 0
t2 = v1 == 0
t3 = ITE(t2, 1, t1)
t4 = v1 == 1
t5 = ITE(t4, 2, t3)
...
 \end{verbatim}
%
It is easy to see how this generalises to computing an arbitrary permutation.
Thus, every permutation is computed a program of size at most $b \times
2^{bk + 1}$.  Since {\sc Kalashnikov} enumerates all programs in order of
length, it will eventually check a program computing each permutation of
$\mathcal{S}$ and so will eventually check a program computing the ranking
function $R''$.  Checking that a single program satisfies the specification
is a decidable problem, and the search as a whole terminates.
%
\end{proof}

\begin{theorem}
 {\sc Headshot} is complete for deterministic programs -- {\sc Headshot} will terminate when run on
 any program with a deterministic transition relation.
\end{theorem}

\begin{proof}
 If the program under analysis is terminating, then by Theorem~\ref{thm:complete-termination} {\sc Headshot}
 will terminate having computed a ranking function.

 Conversely if the program is non-terminating, {\sc Headshot} will terminate
 having computed a recurrence set.  Since, by assumption, the program under analysis is determinstic,
 Definition~\ref{def:ont} is equivalent to Definition~\ref{def:cnt}.
 That is to say, each open recurrence set is also a closed recurrence set.  Since the specification for a predicate
 to encode a closed recurrence set is in the synthesis fragment, {\sc Kalashnikov} will find a program computing
 such a predicate if one exists.  The proof is similar to the proof of Theorem~\ref{thm:complete-termination}.
\end{proof}

We will now show that our time complexity is asymptotically a function of
the Kolmogorov complexity of the (non-)termination proof, and argue
that this gives our procedure various desirable qualities in practical
appliations.
%
\begin{theorem}
 {\sc Kalashnikov} has time complexity $NP^{NP}(K(\sigma))$.
\end{theorem}

\begin{theorem}
 {\sc Headshot} uses a single call to {\sc Kalashnikov} and uses an encoding with $O(1)$
 overhead.  It therefore has time complexity $NP^{NP}(K(\tau(P))$ where $\tau(P)$ is the
 termination specification for program $P$.
\end{theorem}

In Section~\ref{sec:synthesis} we will describe a procedure for solving the synthesis formula.
This procedure has the property that it will find the shortest program meeting the
specification, and as a consequence the runtime of the program synthesiser is dominated
by the size of this shortest program.  To frame the discussion of the runtime of our
termination proving procedure, we first recall the definition of the Kolmogorov complexity
of a function $f$:

\begin{definition}[Kolmogorov complexity]
 The Kolmogorov complexity $K(f)$ is the length of the shortest program that
 computes~$f$.
\end{definition}

We can extend this definition slightly to talk about the Kolmogorov complexity of a
synthesis problem in terms of its specification:

\begin{definition}[Kolmogorov complexity of a synthesis problem]
 The Kolmogorov complexity of a program specification $K(\sigma)$ is the length of the shortest
 program $P$ such that $P \models \sigma$.
\end{definition}

We will later show that the time complexity of our synthesis procedure is $NP^{NP}(K(\sigma))$,
which means that for even moderately large $K(\sigma)$ our procedure is intractably slow.  Conversely
if $K(\sigma)$ is small we will be able to find a termination proof, regardless of other parameters
such as the size of the program under analysis.  To put this another way, under the assumption
that $K(\sigma)$ is small for a program's termination problem, we will find a ranking function
quite rapidly.  We will now investigate the properties of functions with low Kolmogorov complexity,
and show that the assumption of a low-Kolmogorov ranking function is much weaker than the assumption
of a linear ranking function.

\begin{theorem}
 Linear functions have low Kolmogorov complexity.
\end{theorem}

\begin{proof}
 The function $f: X \to Y$ can be computed with a program consisting of
 $2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$ instructions (1 multiplication and 1 addition for
 each cell in the matrix representing $f$).  Therefore,
 $K(f) \leq 2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$.
\end{proof}

\begin{theorem}
Assuming a reasonable program encoding, the probability that a random program of length $l$ computes
a linear function is $O(2^{-l})$.
\end{theorem}

\begin{proof}
 (Do this proof properly).

 Otherwise, we'd need to have an exponential number of programs computing the same function.
 That would be an unreasonable program encoding.
\end{proof}

Therefore the number of non-linear functions $f$ with $K(f) \leq \kappa$
grows exponentially with $\kappa$, while the number of linear functions
grows only linearly.  This shows that assuming the existence of a linear
ranking function is much stronger than assuming the existence of a short
termination proof.  In addition to being a much weaker assumption,
we argue that low Kolmogorov complexity is a more natural assumption than
linearity.  Humans tend to write programs they can understand.  It's hard
to quantify exactly what ``understandable'' means, but we think that
``having a short proof'' is closer to the mark than ``having a linear proof''.

\todo{do something with this?}
Since each loop is bounded by a constant, and each recursive function call is
limited to a constant depth, a \newC program necessarily terminates and in
fact does so in $O(1)$ time.  If we call the largest loop bound~$k$, then
a Bounded Model Checker with an unrolling bound of $k$ will be a complete
decision procedure for the safety of the program.  For a \newC program of
size $l$ and with largest loop bound~$k$, a Bounded Model Checker will
create a SAT problem of size $O(lk)$.  Conversely, a SAT problem
of size $s$ can be converted trivially into a loop-free \newC program
of size $O(s)$.  The safety problem for \newC is therefore NP-complete,
which means it can be decided fairly efficiently for many practical
instances.

\iffalse
\begin{corollary}
 LKC is a weaker assumption than linearity.
\end{corollary}

\begin{theorem}
 LKC programs do not always have LKC ranking functions.
\end{theorem}

\begin{proof}
 This would solve the halting problem, Goldbach conjecture, Collatz conjecture.
\end{proof}

\begin{conjecture}
 High-Kolmogorov-complexity (HKC) programs often have LKC ranking functions.
\end{conjecture}

\begin{theorem}
 \textsc{Headshot} is biased towards finding ranking functions with
 low-Kolmogorov-complexity (LKC).
\end{theorem}

\begin{proof}
 Trivial.
\end{proof}

\subsection{The linearity assumption vs. the LKC one for ranking functions}

\begin{conjecture}
 Most LKC programs compute non-linear functions, but linear functions are LKC.
 So LKC is a weaker assumption than linearity.
\end{conjecture}

\begin{corollary}
 \textsc{Headshot} can often prove termination where linear methods cannot.
\end{corollary}
\fi

\section{Experiments}

We implemented our algorithm in the {\sc Headshot} tool, which uses the {\sc
Kalashnikov} program synthesiser for termination.  We ran the tool on 40
benchmarks taken from the termination category of SVCOMP'15.  The
experiments were performed on a 4-core 3.3\,GHz i5-2500k with 8\,GB of RAM.

\iffalse
We proved termination for a bunch of programs, see Fig.~\ref{fig:linear} and Fig.~\ref{fig:nonlinear}.

\begin{figure*}
\centering
\begin{tabular}{|l|r|r||r|r|r|r|}
\hline
    & LOC & \shortstack{Rank function \\ size} & \textsc{T2} & \textsc{ARMC} & \textsc{Headshot} & \textsc{Headshot-Linear} \\
    \hline
    \hline
 P1 & 10 & 3 & 100s & 70s & 0.1s & \bf{0.01s} \\
 P2 & 10 & 3 & 100s & 70s & 0.1s & \bf{0.01s} \\
 P3 & 10 & 3 & 100s & 70s & 0.1s & \bf{0.01s} \\
 \hline
\end{tabular}
\caption{Termination for linear programs with disjunctive, linear ranking functions\label{fig:linear}}
\end{figure*}
\fi

\begin{figure*}
\centering
\footnotesize
\begin{tabular}{|l|r|c|c|c|c|c|r|r|c|r|r|}
\hline
    & LOC & Terminates? & \shortstack{Linear \\ program?} & \shortstack{Linear ranking \\ function?}  & Conditional? & Float? & Dimension & \shortstack{Ranking \\ program size} & Res & Time (s) & Iterations\\
    \hline
    \hline
\input{table}
    \hline
\end{tabular}

\caption{\textsc{Headshot} termination for nonlinear programs with nonlinear ranking functions\label{fig:nonlinear}}
 \end{figure*}

 \section{Conclusions and Related Work}
%\subsection{Termination analysis}
Automated program termination is a research topic that has received a fair amount of attention from the software verification community.
In order to compare our technique to the rest of the area, 
Figure~\ref{fig:handletable} summarises the related works with respect to the restrictions they impose on the transition relations as well as the form of the computed ranking functions.

Somehow expected, most of the techniques are specialised in the synthesis of linear ranking functions for linear programs over integers (or rationals) \cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13}. 
Among them, 
Lee et al. make use of transition predicate abstraction, algorithmic learning, and decision procedures to compute transition
invariants as proofs for the termination of linear programs \cite{DBLP:conf/cav/LeeWY12}.
Leike and Heizmann present a new method for the constraint-based synthesis
of termination arguments for linear loop programs based on
linear ranking templates \cite{DBLP:conf/tacas/LeikeH14}.
Linear ranking functions supported by inductive linear invariants for loops with linear guards and transitions: \cite{DBLP:conf/cav/BradleyMS05}. 
Learning: \cite{DBLP:journals/corr/HeizmannHP14}

%A program is lasso-shaped if it is composed by a stem followed by a single loop without branching, i.e. there is only one path. 
%Consequently, the termination argument can be very simple. There are numerous techniques specialised in
%proving termination of lasso-shaped programs efficiently \cite{DBLP:conf/popl/Ben-AmramG13,DBLP:conf/cav/BradleyMS05,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/vmcai/P04}.

While the synthesis of termination arguments for linear programs over
integers is indeed well-covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.  present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

Given the lack of research in handling \emph{not-linear programs}, as well
as \emph{programs over bit vectors and floats}, our work focuses on covering
these areas.  One of the obvious conclusions that can be reached from
observing Figure~\ref{fig:handletable}, is that most of the works tend to
specialise on a certain aspect of termination proving that they can solve
efficiently.  Conversely to this view, we aim for \emph{generality}, as we
do not restrict the form of the synthesised ranking functions, nor the form
of the input programs.

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution follows through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
In an attempt to reduce the complexity of checking the validity of the
termination argument, Cook et al present an iterative termination proving
procedure that searches for lexicographic termination
arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et al. 
strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  For this purpose, several
tools alternate between calls to a safety prover and calls to a ranking
function synthesis tool \cite{}.  In order to share more information about
the state of the termination proofs between the two tools, Brockschmidt et
al.  use a single representation of the state of the termination proof
search that both tools operate over, resulting in performance improvements
\cite{DBLP:conf/cav/BrockschmidtCF13}.  In \cite{DBLP:conf/cav/BradleyMS05},
Bradley et al.  combine the generation of ranking functions with the
generation of invariants to form one constraint solving problem such that
the necessary supporting invariants for the ranking function are discovered
on demand.  In our setting, both the ranking function and the supporting
invariant are iteratively constructed through the same refinement loop.

%In our setting, a candidate termination argument is iteratively constructed. 
%A candidate termination argument is composed out of a ranking function 
%and a supporting invariant.

%% The safety prover proves or disproves the
%% validity of the current argument via the search for invariants. Refinement of the
%% current termination argument is performed using the output of a rank function
%% synthesis tool when applied to counterexamples found by the safety prover.

%The existing work on termination is primarily divided into structural (syntactic) vs. semantic approaches.  
%For example, terminationg of term rewriting systems is largely structural, whereas DWF analysis (cf. Terminator) is more semantic. While it deals
%with transition relations, it also uses some amount of syntax in that it identifies loops \& looping paths.

%-- Usually the termination argument for the program LOOPS
%on Figure 1 is based on a lexicographic combination
%of well-founded orderings.

While program termination has been extensively studied, much less research has been conducted in the area of proving nontermination.
% A counterexample to termination is an infinite program execution. 
Gupta et al. dynamically enumerate lasso-shaped candidate
paths for counterexamples, and then statically prove their feasibility \cite{DBLP:conf/popl/GuptaHMRX08}. 
In \cite{DBLP:conf/tacas/ChenCFNO14}, Chen et al. prove nontermination via
reduction to safety proving. Their iterative algorithm uses counterexamples to a fixed safety property
to refine an underapproximation of a program.

% program analysis as constraint solving:
%% how the constraint-based approach can be used to model a wide spectrum of program analyses in an 
%%expressive domain containing disjunctions and conjunctions of linear inequalities.

\bibliographystyle{abbrvnat}
\bibliography{synth}{}

\end{document}