### 2014-07-06 papers/termination/table.tex
Change results table

loop1.c & ? & \tick & 0.06s & \tick & 1.3s & 1\\

loop2.c & ? & \tick & 0.06s & \tick & 1.4s & 4\\

loop3.c & \tick & \tick & 0.06s & \tick & 1.8s & 8\\

loop4.c & ? & \tick & 0.12s & \tick & 2.9s & 18\\

loop5.c & ? & \tick & 0.12s & \tick & 5.3s & 18\\

loop6.c & ? & \tick & 0.05s & \tick & 1.2s & 1\\

loop7.c & ? & \xmark & 0.05s & \tick & 8.3s & 17\\

loop8.c & ? & \xmark & 0.06s & \tick & 1.3s & 1\\

loop9.c & \tick & \tick & 0.11s & \tick & 1.6s & 1\\

loop10.c & ? & \xmark & 0.05s & \tick & 1.3s & 1\\

loop43.c & ? & \tick & 0.07s & \tick & 1.5s & 1\\

loop44.c & \xmark & \xmark & 0.05s & \xmark & 10.5s & 9\\

loop45.c & ? & \tick & 0.12s & \tick & 4.3s & 18\\

loop46.c & ? & \xmark & 0.05s & \tick & 1.5s & 2\\

loop47.c & ? & \tick & 0.10s & \tick & 1.8s & 1\\

loop48.c & ? & \tick & 0.06s & \tick & 1.4s & 4\\

loop49.c & ? & \xmark & 0.05s & ? & T/O & 206\\

svcomp1.c & ? & \tick & 0.11s & \tick & 2.3s & 2\\

svcomp2.c & ? & \tick & 0.05s & \tick & 1.5s & 9\\

svcomp3.c & ? & \tick & 0.15s & \tick & 146.4s & 43\\

svcomp4.c & \xmark & \xmark & 0.09s & \xmark & 2.1s & 7\\

svcomp5.c & ? & \tick & 0.38s & ? & T/O & 46\\

svcomp6.c & ? & ? & T/O & \tick & 29.1s & 28\\

svcomp7.c & \tick & \tick & 0.09s & \tick & 5.5s & 17\\

svcomp8.c & ? & \xmark & 0.05s & ? & T/O & 37\\

svcomp9.c & ? & \tick & 0.10s & \tick & 1.5s & 8\\

svcomp10.c & ? & \tick & 0.11s & \tick & 4.5s & 18\\

svcomp11.c & ? & \tick & 0.20s & \tick & 14.6s & 33\\

svcomp12.c & ? & ? & T/O & \tick & 10.9s & 24\\

svcomp13.c & ? & \xmark & 0.07s & \tick & 35.1s & 62\\

svcomp14.c & ? & ? & T/O & \tick & 30.8s & 26\\

svcomp15.c & ? & \xmark & 0.12s & ? & T/O & 191\\

svcomp16.c & ? & \tick & 0.06s & \tick & 2.2s & 3\\

svcomp17.c & ? & \tick & 0.05s & ? & T/O & 32\\

svcomp18.c & ? & \xmark & 0.27s & ? & T/O & 26\\

svcomp25.c & ? & \xmark & 0.05s & ? & T/O & 26\\

svcomp26.c & ? & \tick & 0.26s & \tick & 3.2s & 23\\

svcomp27.c & \xmark & \tick & 0.11s & ? & T/O & 67\\

svcomp28.c & ? & \tick & 0.13s & ? & T/O & 79\\

svcomp29.c & ? & \xmark & 0.05s & ? & T/O & 1\\

svcomp37.c & ? & \tick & 0.16s & \tick & 2.1s & 10\\

svcomp38.c & \tick & \tick & 0.10s & ? & T/O & 1\\

svcomp39.c & ? & \tick & 0.25s & ? & T/O & 1\\

svcomp40.c & ? & \xmark & 0.07s & ? & T/O & 30\\

svcomp41.c & ? & \xmark & 0.07s & ? & T/O & 27\\

svcomp42.c & ? & \tick & 0.22s & ? & T/O & 85\\