### 2014-07-06 papers/termination/paper.tex
Remove comments

\documentclass[preprint]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{listings}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{pgf}
\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{multicol}
\usepackage{appendix}
\usepackage{caption,subcaption}
\DeclareCaptionType{copyrightbox}
\usepackage{subfig}
\usepackage{multirow}
\usepackage{framed}
\usepackage{tikz}
\usetikzlibrary{arrows, automata, shapes, calc}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\newcommand{\xmark}{\ding{55}}
\newcommand{\tick}{\checkmark}
\newcommand{\todo}[1]{{\bf TODO:} #1}
\newcommand{\newC}{C$^-$\xspace}

\lstset{language=c++}

\algrenewcommand\algorithmicrequire{\textbf{Precondition:}}
\algrenewcommand\algorithmicensure{\textbf{Postcondition:}}

\newcommand*\Let[2]{\State #1 $\gets$ #2}

\newcommand{\bv}[2]{\mathcal{BV}(#1, #2)}

\makeatletter
\pgfdeclareshape{datastore}{
  \inheritsavedanchors[from=rectangle]
  \inheritanchorborder[from=rectangle]
  \inheritanchor[from=rectangle]{center}
  \inheritanchor[from=rectangle]{base}
  \inheritanchor[from=rectangle]{north}
  \inheritanchor[from=rectangle]{north east}
  \inheritanchor[from=rectangle]{east}
  \inheritanchor[from=rectangle]{south east}
  \inheritanchor[from=rectangle]{south}
  \inheritanchor[from=rectangle]{south west}
  \inheritanchor[from=rectangle]{west}
  \inheritanchor[from=rectangle]{north west}
  \backgroundpath{
    %  store lower right in xa/ya and upper right in xb/yb
    \southwest \pgf@xa=\pgf@x \pgf@ya=\pgf@y
    \northeast \pgf@xb=\pgf@x \pgf@yb=\pgf@y
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@ya}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@ya}}
    \pgfpathmoveto{\pgfpoint{\pgf@xa}{\pgf@yb}}
    \pgfpathlineto{\pgfpoint{\pgf@xb}{\pgf@yb}}
 }
}
\makeatother

\begin{document}

%\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{POPL '15}{Month d--d, 20yy, City, ST, Country} 
\copyrightyear{2015}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

\title{Deciding Bitvector Termination with Program Synthesis}

\authorinfo{Cristina David\and Daniel Kroening\and Matt Lewis}
           {University of Oxford}
           {firstname.lastname@cs.ox.ac.uk}

\maketitle

\begin{abstract}
%
Proving program termination is typically done by finding a well-founded
\emph{ranking function} for the program states.  Existing termination
provers typically find ranking functions using either linear algebra or
templates.  As such they are often restricted to finding linear ranking
functions over mathematical integers.  This class of ranking functions is
insufficient for proving termination of many terminating programs, and
furthermore a termination argument for a program operating on mathematical
integers does not always lead to a termination argument for the same program
operating on fixed-width machine integers.

We present a reduction from program \emph{termination} to program
\emph{synthesis}.  This reduction allows us to generate nonlinear,
lexicographic ranking functions that are correct for fixed-width machine
arithmetic and floating-point arithmetic.  We use this reduction to build a
sound and complete procedure for the termination of fixed-width and
floating-point arithmetic programs.
%
\end{abstract}

%\category{CR-number}{subcategory}{third-level}

\keywords
Termination, Program Synthesis, Lexicographic Ranking Functions, Bitvector Ranking Functions,
Floating-Point Ranking Functions.

\section{Introduction}\label{sec:intro}

The halting problem has been of central interest to computer scientists
since it was first considered by Turing in 1936~\cite{turing}.  Informally,
we can say that the halting problem is concerned with answering the question
``does this program run forever, or will it eventually terminate?''

Proving program termination is typically done by finding a \emph{ranking
function} for the program states, i.e.  a monotone map from the program's
state space to a well-ordered set.  Historically, the search for ranking
functions has been constrained in various syntactic ways, leading to
incompleteness, and is performed over abstractions that do not soundly
capture the behaviour of physical computers.  In this paper, we present a
sound and complete method for deciding whether a program terminates.

When surveying the area of program termination chronologically, we observe
an initial focus on monolithic approaches based on a single measure shown to
decrease over all program
paths~\cite{DBLP:conf/vmcai/P04,DBLP:conf/cav/BradleyMS05}, followed by more
recent techniques that use termination arguments based on Ramsey's
theorem~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
Briefly, the transition relation is shown to be disjunctively well founded
by finding a set of simple functions, each of which proves well foundedness
for part of the transition relation.  The main benefit of this approach is
the simplicity of local termination measures in contrast to global ones. 
For instance, there are cases in which linear arithmetic suffices when using
local measures, while corresponding global measures require non-linear
functions or lexicographic orders.

One drawback of the Ramsey-based approach is that the validity of the
termination argument relies on checking the \emph{transitive closure} of the
program, rather than a single step.  As such, there is experimental evidence
that most of the effort is spent in reachability
analysis~\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/KroeningSTW10},
requiring the support of powerful safety checkers: there is a tradeoff
between the complexity of the termination arguments and that of checking
their validity.

As Ramsey-based approaches are limited by the state of the art in safety
checking, recent research shifts back to more complex termination arguments
that are easier to
check~\cite{DBLP:conf/cav/KroeningSTW10,DBLP:conf/tacas/CookSZ13,DBLP:conf/cav/HeizmannHP14}. 
Following the same trend, we investigate its extreme: \emph{unrestricted}
termination arguments.  This means that our ranking functions may involve
non-linearity and lexicographic orders: we do not commit to any particular
syntactic form, and do not use templates.

Figure~\ref{fig:handletable} summarises the related work with respect to the
restrictions they impose on the transition relations as well as the form of
the ranking functions computed.  While it supports the observation that the
majority of existing termination analyses are designed for linear programs
and linear ranking functions, it also highlights another simplifying
assumption made by most state-of-the-art termination provers: that
bit-vector semantics and integer semantics give rise to the same termination
behaviour.  Thus, most of the techniques treat bit-vectors and IEEE floats
as mathematical integers and reals,
respectively~\cite{DBLP:conf/pldi/CookPR06,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/vmcai/BradleyMS05,DBLP:conf/cav/KroeningSTW10}.

By assuming bit-vector semantics to be identical to integer semantics, these
techniques ignore the wrap-around behaviour caused by overflows, which can
be unsound.  In Section~\ref{sec:motivation}, we show that integers and
bit-vectors exhibit incomparable behaviours with respect to termination,
i.e.  programs that terminate for integers may \emph{not} terminate for
bit-vectors and vice versa.  Thus, abstracting bit-vectors to integers may
give rise to {\em unsound} and {\em incomplete} analyses.

We propose a general framework that uniformly computes lexicographic,
non-linear ranking functions supported by non-linear inductive invariants
for loops with arbitrary guards and transitions over bit-vectors and floats. 
It also computes non-linear recurrence sets, which comprise non-termination
proofs.  Our encoding treats multiple loops and nested loops uniformly,
without having to enumerate lassos.

 For this purpose, we phrase the termination and non-termination problems as
second-order satisfaction problems.  We identify a fragment of second-order
logic that is expressive enough to capture our formulation, which we call
the \emph{synthesis fragment}.  Subsequently, we propose a method for
checking the satisfiability of a formula in the synthesis fragment by
solving an isomorphic program synthesis problem
(Section~\ref{sec:synthesis}).  Our method is sound and semi-complete, i.e. 
if a pair of a ranking function and a supporting invariant exist it will be
found, but if the program does not terminate we may not find a recurrence
set.  However, if the program's transition relation is deterministic, our
method is complete.

With respect to the performance of our technique, we show that its runtime is dominated by the size of 
the shortest termination proof.

This means that our procedure is not directly dependent on the structure of
the analysed loop (e.g.~the number of variables or the number of lassos),
but on the \emph{Kolmogorov complexity of its minimal termination argument}. 
We argue that the length of a program's termination proof is indicative of
how easy to understand the program is, and thus programmers tend to write
programs that have relatively short termination proofs.  We empirically show
that this claim holds in practice, causing our technique to perform well.

\begin{figure*}
\centering
 \begin{tabular}{|ll||c|c|c|c|c|c|c|c|}
 \hline
  & & \multicolumn{8}{c|}{Program} \\
  & & \multicolumn{2}{c|}{Integers} & \multicolumn{2}{c|}{Reals} & \multicolumn{2}{c|}{Bit-vectors} & \multicolumn{2}{c|}{Floats} \\
  & & L & NL & L & NL & L & NL & L & NL \\
  \hline
  \hline
  \multirow{4}{*}{Ranking} & Linear lexicographic &  \cite{DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13,DBLP:conf/vmcai/P04} && & &\checkmark&\checkmark&\checkmark&\checkmark\\
   & Linear non-lexicographic & \cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/vmcai/BradleyMS05,DBLP:conf/cav/KroeningSTW10} & \cite{DBLP:conf/vmcai/BradleyMS05} & && \checkmark~ \cite{DBLP:conf/tacas/CookKRW10} &\checkmark~ \cite{DBLP:conf/tacas/CookKRW10}&\checkmark&\checkmark\\
   & Non-linear lexicographic &  &  & &&\checkmark&\checkmark&\checkmark&\checkmark\\
   & Non-linear non-lexicographic & \cite{DBLP:conf/vmcai/BradleyMS05} &  \cite{DBLP:conf/vmcai/BradleyMS05} & &&\checkmark&\checkmark&\checkmark&\checkmark\\
   \hline
 \end{tabular}

 \caption{Legend: \checkmark = we can handle\label{fig:handletable}}
\end{figure*}

The main contributions of our work can be summarised as follows:
%
\begin{itemize}

\item We designed a bit-level accurate technique for computing ranking
functions that correctly accounts for the wrap-around behavior caused by
under- and overflows in bit-vector and floating-point arithmetic.  Our
technique is not restricted to finding linear ranking functions, but can
also compute lexicographic non-linear ones.

\item  We rephrased the termination and non-termination problems as
second-order satisfaction problems and used program synthesis techniques to
solve them.  In contrast to previous approaches to program synthesis, we
found that using genetic programming was very effective at generating
candidate termination and non-termination proofs.

\item Our procedure does not require the enumeration of lassos, or directly
reasoning about the transitive closure of the transition relation.  To
verify our proofs, we only require a single call to a SAT solver.

\item Our procedure is based on a series of source-to-source
transformations.  As well as leading to a concise implementation, this has
the side effect of ensuring that our proof is correct for the exact binary
produced by a particular compiler, rather than for an abstraction of the
program.

\item We implemented our technique and tried it on a selection of programs
handling both bit-vectors and floats.

\item We propose a non-trivial formulation of the (non-)termination problem inside a
 second-order logic fragment with only one quantifier alternation.

\end{itemize}

{\bf Limitations.}
Our algorithm decides termination of transition systems with finite state
spaces.  The (non-)termination proofs take the form of ranking functions and
program invariants that are expressed in a quantifier free language.  This
formalism is powerful enough to handle a large fragment of C, but is not
rich enough to analyse code using any of the following features:

\begin{itemize}
\item Recursion (could probably do this).
\item Heap (Thor).
\item Arrays.
\end{itemize}

Additionally, if the program under analysis has a non-deterministic transition relation
and is non-terminating, our procedure is not guaranteed to find a non-termination proof.

\section{Motivating Examples} \label{sec:motivation}

Figure~\ref{fig:handletable} illustrates the most common simplifying
assumptions made by existing termination analyses:
%
\begin{itemize}
\item[(i)] programs use only linear arithmetic.
\item[(ii)] terminating programs have termination arguments expressible in linear arithmetic.
\item[(iii)] the semantics of bit vectors and mathematical integers are equivalent.
\item[(iv)] the semantics of IEEE floating-point numbers and mathematical reals are equivalent.
\end{itemize}

To show how these assumptions are violated by even simple programs, we draw
the reader's attention to the programs in Figure~\ref{fig:motivation} and
their curious properties:

%
\begin{itemize}

\item Program (a) breaks assumption (i) as it makes use of the bit-wise $\&$ operator.
%
Our technique finds that an admissible ranking function is the linear
function $R(x) = x$, whose value decreases with every iteration, but cannot
decrease indefinitely as it is bounded from below.  This example also
illustrates the lack of a direct correlation between the linearity of a
program and that of its termination arguments.

\item Program (b) breaks assumption (ii), in that it has no linear ranking
function.  We prove that this loop terminates by finding the non-linear
ranking function $R(x) = |x|$.

\item Program (c) breaks assumption (iii).  This loop is terminating for
bit-vectors since $x$ will eventually over-flow and become negative. 
Conversely, the same program is non-terminating using integer arithmetic
since $x > 0 \rightarrow x+1 > 0$ for any integer $x$.

\item Program (d) also breaks assumption (iii), but ``the other way'': it
terminates for integers but not for bit-vectors.  If each of the variables
is stored in an unsigned $k$-bit word, the following entry state will lead
to an infinite loop:
%
$$ M = 2^k - 1,\quad N = 2^k - 1,\quad i = M,\quad j = N-1 $$

\item Program (e) breaks assumption (iv): it terminates for reals but not
for floats.  If $x$ is sufficiently large, rounding error will cause the
subtraction to have no effect.

\item Program (f) breaks assumption (iv) ``the other way'': it terminates
for floats but not for reals.  Eventually $x$ will become sufficiently small
that the nearest representable number is $0.0$, at which point it will be
rounded to $0.0$ and the loop will terminate.

\end{itemize}

Up until this point, we considered examples that are not soundly treated by
existing techniques as they don't fit in the range of programs addressed by
these techniques.  Next, we look at some programs that are handled by
existing termination tools via dedicated analyses.  We show that our method
handles them uniformly, without the need for any special treatment.
%
\begin{itemize}

\item Program (g) is a linear program that is shown
in~\cite{DBLP:conf/tacas/CookSZ13} not to admit (without prior manipulation)
a lexicographic linear ranking function.  With our technique we can find the
non-linear ranking function $R(x) = |x|$.

%As with linear lexicographic ranking functions, a state is mapped to a tuple of values such that the
%loop transition leads to a decrease with respect to the lexicographic
%ordering for this tuple. Therefore no function may increase unless a function of
%a lower index decreases. Additionally, at every step, there must be at least one
%function that decreases.

\item Program (h) illustrates conditional termination.  When proving program
termination we are simultaneously solving two problems: the search for a
termination argument, and the search for a supporting
invariant~\cite{DBLP:conf/cav/BrockschmidtCF13}.  For this loop, we find the
ranking function $R(x) = x$ together with the supporting invariant $y=1$.

\item In the terminology of \cite{DBLP:conf/tacas/LeikeH14}, program (i)
admits a \emph{multiphase} ranking function, computed from a multiphase
ranking template.  Multiphase ranking templates are targeted at programs
that go through a finite number of phases in their execution.  Each phase is
ranked with an affine-linear function and the phase is considered to be
completed once this function becomes non-positive.

In our setting this type of programs does not need special treatment, as we
can find a non-linear lexicographic ranking function $R(x, y, z) = (x < y,
z)$.\footnote{This termination argument is somewhat subtle.  The Boolean
values $\mathit{false}$ and $\mathit{true}$ are interpreted as 0 and 1,
respectively.  The Boolean $x < y$ thus eventually decreases, that is to say
once a state with $x \geq y$ is reached, $x$ never again becomes greater
than $y$.  This means that as soon as the ``else'' branch of the if
statement is taken, it will continue to be taken in each subsequent
iteration of the loop.  Meanwhile, if $x < y$ has not decreased (i.e., we
have stayed in the same branch of the ``if''), then $z$ does decrease. 
Since a Boolean only has two possible values, it cannot decrease
indefinitely.  Since $z > 0$ is a conjunct of the loop guard, $z$ cannot
decrease indefinitely, and so $R$ proves that the loop is well founded.}

%\item Program (j) illustrates a nested loop.  This construction can
%cause difficulties for provers based on enumerating lassos

\end{itemize}
%
As with all of the termination proofs presented in this paper, the ranking
functions above were all found completely automatically.

\begin{figure*}
\centering
 \begin{tabular}{ccc}

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x > 0) {
  x = (x - 1) & x;
}
\end{lstlisting}
\caption{Taken from~\cite{DBLP:conf/tacas/CookKRW10}.}
 \label{fig:motivation.a}
\end{subfigure}%

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x != 0) {
  x = -x / 2;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.b}
\end{subfigure}%

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}[language=C]
while(x > 0) {
  x++;
}
 \end{lstlisting}
\caption{}
 \label{fig:motivation.c}
\end{subfigure} \\

\hline

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (i < M || j < N) {
  i = i + 1;
  j = j + 1;
}
\end{lstlisting}
\caption{Taken from~\cite{DBLP:conf/sigsoft/Nori013}}
 \label{fig:motivation.d}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
float x;

while (x > 0.0) {
  x -= 1.0;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.e}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
float x;

while (x > 0.0) {
  x *= 0.5;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.f}
\end{subfigure} \\
\hline

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x != 0) {
  if (x > 0)
    x--;
  else
    x++;
}
\end{lstlisting}
\caption{Taken from \cite{DBLP:conf/tacas/CookSZ13}}
 \label{fig:motivation.g}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
y = 1;

while (x > 0) {
  x = x - y;
}
\end{lstlisting}
\caption{}
 \label{fig:motivation.h}
\end{subfigure}

&

\begin{subfigure}[b]{0.3\textwidth}
\begin{lstlisting}
while (x > 0 && y > 0 && z > 0) {
  if (y > x) {
    y = z;
    x = nondet();
    z = x - 1;
  } else {
    z = z - 1;
    x = nondet();
    y = x - 1;
  }
}
\end{lstlisting}
\caption{Taken from~\cite{BA:mcs}}
 \label{fig:motivation.i}
\end{subfigure} 
% 
% \\
% 
% \hline
% 
% &
% 
% \begin{subfigure}[b]{0.3\textwidth}
% \begin{lstlisting}
% while (i < n) {
%   j = 0;
%   while (j <= i) {
%     j = j + 1;
%   }
%   i = i + 1;
% }
% \end{lstlisting}
% \caption{Taken from~\cite{DBLP:conf/cav/BrockschmidtCF13}}
% \end{subfigure}

\end{tabular}
\caption{Motivational examples, mostly taken from the literature.\label{fig:motivation}}
\end{figure*}

\section{Preliminaries}

Given a program, we first formalise its termination argument as a ranking
function (Section~\ref{sec:ranking.functions}).  Subsequently, we discus
bit-vector semantics and illustrate differences between machine arithmetic
and integer arithmetic that show that the abstraction of bit vectors to
mathematical integers unsound (Section~\ref{sec:machine.arith}).

\subsection{Termination and Ranking Functions} \label{sec:ranking.functions}

A programs $P$ is represented as a transition system with state space $X$ and
transition relation $T \subseteq X \times X$.  For a state
$x \in X$ with $T(x,x')$ we say $x'$ is a successor of $x$ under $T$.
If each state $x$ has at most one successor $x'$, then we say that $T$ is
deterministic.

\begin{definition}[Unconditional termination]
%
A transition system is said to be \emph{unconditionally terminating} if
there is no infinite sequence of states $x_1, x_2, \ldots \in X$ with
$\forall i.~T(x_i, x_{i+1})$.
%
\end{definition}

We can prove that the transition system is unconditionally terminating by
finding a ranking function for its transition relation.
%
\begin{definition}[Ranking function]
%
An injective function ${R:X\to Y}$ is a \emph{ranking function} for the
transition relation $T$ if $Y$ is a well-founded set with order $<$ and 
$R$ is monotonically decreasing with respect to $T$.  That is
to say:
$$\forall x, x' \in X. T(x, x') \Rightarrow R(x) < R(x')$$
%
\end{definition}

\begin{definition}[Linear ranking function]
A \emph{linear ranking function} $R: X \to Y$ 
with $\dim(X) = n$ and $\dim(Y) = m$ is of the form: $$f(\vec{x}) = M\vec{x}$$ where
$M$ is an $n \times m$ matrix.
\end{definition}

In the case that $\dim(Y) = 1$, this reduces to the inner product
%
$$f(\vec{x}) = \vec{\lambda} \cdotp \vec{x} + c \;.$$

\begin{definition}[Lexicographic ranking function]
For $Y = Z^m$, we say that a ranking function $R: X \to Y$ is \emph{lexicographic}
if it maps each state in $X$ to a tuple of values such that the loop transition leads to a decrease with
respect to the lexicographic ordering for this tuple.
The total order imposed on $Y$ is the lexicographic ordering
induced on tuples of $Z$'s. 
\end{definition}

For lexicographic ranking functions, no function may increase unless a function of a lower index decreases.
Additionally, at every step, there must be at least one function that decreases.
We note that some termination arguments require lexicographic ranking functions, or equivalently, ranking functions
whose co-domain is the ordinals, rather than just $\mathbb{N}$.

\subsection{Machine Arithmetic and Bit Vectors} \label{sec:machine.arith}

Physical computers have bounded storage, which means they are unable to
perform calculations on mathematical integers.  For example, if $A$ is the
Ackermann function and $G$ is Graham's number, a physical computer capable
of computing $A(G, G)$ would contain (much!) more matter than is believed to
exist in the universe.  Fortunately, it is rare for a programmer to need
such a large number and so modern computers do their arithmetic over
fixed-width binary words, otherwise known as bit vectors.  For the remainder
of this section, we will say that the bit vectors we are working with are
$k$-bits wide, which means that each word can hold one of $2^k$ bit
patterns.  Typical values for $k$ are 32 and 64.

Machine words can be interpreted as ``signed'' or ``unsigned'' values. 
Signed values can be negative, while unsigned values cannot.  The encoding
for signed values is two's complement, where the most significant bit
$b_{k-1}$ of the word is a ``sign'' bit, whose weight is $-(2^k - 1)$ rather
than $2^k - 1$.  Two's complement representation has the property that
$\forall x .  -x = (\mathord{\sim} x) + 1$, where $\mathord{\sim}(\bullet)$
is bitwise negation.  Two's complement also has the property that addition,
multiplication and subtraction are defined identically for unsigned and
signed numbers.

Bit-vector arithmetic is performed modulo $2^k$, which is the source of many
of the differences between machine arithmetic and Peano
arithmetic.\footnote{ISO C requires that unsigned arithmetic is performed
modulo $2^k$, whereas the overflow case is undefined for signed arithmetic. 
In practice, the undefined behaviour is implemented just as if the
arithmetic had been unsigned.}.  To give an example, $(2^k - 1) + 1 \equiv 0
\pmod {2^k}$ provides a counterexample to the statement $\forall x. 
x + 1 > x$, which is a theorem of Peano arithmetic but not of modular
arithmetic.  When an arithmetic operation has a result greater than $2^k$,
it is said to ``overflow''.  If an operation does not overflow, its
machine-arithmetic result is the same as the result of the same operation
performed on integers.

The final source of difference between integer arithmetic and bitvector
arithmetic stems from width conversions.  Many programming languages allow
numeric variables of different types, which can be represented using words
of different widths.  In C, a \texttt{short} might occupy 16 bits, while an
\texttt{int} might occupy 32 bits.  When a $k$-bit variable is assigned to a
$l$-bit variable with $l < k$, the result is truncated $\mathrm{mod}~2^l$.  For
example, if $x$ is a 32-bit variable and $y$ is a 16-bit variable, $y$ will
hold the value $0$ after the following code is executed:
%
\begin{lstlisting}
x = 65536;
y = x;
\end{lstlisting}

This gives us a counterexample to the statement $\forall x, y . x = y
\Rightarrow (x + 1) = (y + 1)$, which is a theorem of Peano arithmetic.

As well as machine arithmetic differing from Peano arithmetic on the
operators they have in common, computers have several ``bitwise'' operations
that are not taken as primitive in the theory of integers.  These operations
include the Boolean operators \texttt{and, or, not, xor} applied to each
element of the bit vector.  Computer programs often make use of these
operators, which are non-linear when interpreted in the standard model of
Peano arithmetic\footnote{Some of these operators can be seen as
linear in a different algebraic structure, e.g.~\texttt{xor} corresponds to
addition in the Galois field $\mathrm{GF}(2^k)$.}.

\iffalse
\section{Combinatorics of Finite Termination}
In this section, we fix a model of computation, describe its semantics and
define the syntax of a language we will work over.

\subsection{Syntax and Semantics}

\begin{itemize}
 \item Our transition relation is $T(x, x') \subseteq X \times X$.
 \item Our loop condition is $L(x) \subseteq X$.
 \item Our ranking function is $R(x) : X \to Y$.
 \item Our state space has size $\| X \| = M = 2^k$.
 \item Our ranking co-domain has size $\| Y \| = N = 2^j$.
 \item The number of looping states is $\| L \| = l$.
 \item Our transition relation is deterministic and parititioned into chains of length $c_i$, with $l = \sum c_i$.
\end{itemize}

\subsection{Counting Programs}
\begin{itemize}
 \item There are a TON of programs (way more than you'd expect).
 \item There are a TON of terminating programs, and for our model of computation we can count
  how many (the Chaitin constant).
 \item There are a TON of ranking functions (way more than you'd expect, but not many as a
  fraction of programs).
 \item There are not many linear functions.
 \item Most terminating programs don't admit linear ranking functions.
 \item The Curry-Howard isomorphism
 \item Kolmogorov complexity is relevant for understanding termination proofs.
\end{itemize}

\begin{theorem}
 A random function $f : X \to Y$ is a ranking function for $(T, L)$ with probability

 $$N^{-l} \times \prod {{N-1} \choose c_i}$$
\end{theorem}

\begin{proof}
 Combinatorics.
\end{proof}

\begin{corollary}
 This number is really small (e.g. $10^{-193}$ for a 64-bit program with 1 variable and 10 looping states.
 Randomly sampling functions \& hoping they're ranking functions is not going to work.
\end{corollary}

\begin{conjecture}
 The probability that a random program $(T, L)$ is terminating (the Chaitin constant)
 is $0.7$.
\end{conjecture}

\begin{conjecture}
 The probability that a random program $(T, L)$ admits a linear ranking function is
 $0.1$.
\end{conjecture}

\begin{conjecture}
 The probability that a random, terminating program $(T, L)$ admits a linear ranking function
 is $0.2$.
\end{conjecture}

\begin{corollary}
 Most terminating programs do not have linear ranking functions.
\end{corollary}
\fi

\section{Termination as Second-Order Satisfaction} \label{sec:second.order}

The problem of program verification can be reduced to the problem of finding
solutions to a second-order
constraint~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08}. 
Our intention is to apply this approach to termination analysis.  In this
section we show how several variations of both the termination and the
non-termination problem can be defined in second-order logic.

In fact, we identify a fragment of second-order logic with a constrained use
of quantification that is expressive enough to encode both termination and
non-termination.  We will suggestively refer to the fragment as the
\emph{synthesis fragment}:

\begin{definition}[Synthesis Fragment]
A formula is in the \emph{synthesis fragment} iff it is of the form
%
 \[
  \exists \vec{P},~ \vec{x} . ~\forall~ \vec{y} . \sigma(\vec{P}, \vec{x}, \vec{y})
 \]
%
where $\vec{P}$ ranges over sets, while $\vec{x}$ and $\vec{y}$ range over
ground terms.  Function $\sigma: (X^n \times Y^m \to Z^k) \times X^n \times
Y^m \to \mathbb{B}$ can be seen as a specification function that returns
true iff the functions $\vec{P}$ compute appropriate outputs when fed the
inputs $\vec{x}$ and $\vec{y}$.
%
\end{definition}
%
Checking satisfiability of a formula in the synthesis fragment corresponds to program synthesis and
%Solving the synthesis problem 
amounts to finding witnesses for the unknowns $\vec{P}$ and $\vec{x}$ that meet the specification
for all $\vec{y}$. 
If a pair $(\vec{P_0}, \vec{x})$ is a solution to the synthesis problem, then we write $(\vec{P_0}, \vec{x}) \models \sigma$.
For the remainder of the presentation, we drop the $\vec{x}$ and instead just write $x$, with the understanding
that all variables range over vectors.

In the rest of this section, we show that the synthesis fragment 
is expressive enough to encode both termination and non-termination. 
Following our formulation as second-order satisfaction, 
in Section~\ref{sec:synthesis} we present a solver for the synthesis fragment that models bit-accurate semantics.

\subsection{An Isolated Loop}

We will begin our discussion by showing how to encode the
\mbox{(non-)termination} of a program consisting of a single loop in the
synthesis fragment.  For the time being, a loop is a pair ${L= \langle G,T
\rangle}$, where the states $x$ satisfying the loop's guard are given by the
predicate $G(x)$.  The body of the loop is encoded as the transition
relation $T(x, x')$, meaning that state $x'$ is reachable from state $x$ via
a single iteration of the loop body.  For example, the loop in
Figure~\ref{fig:motivation.a} is encoded as:
%
\begin{align*}
G(x) & = \{ x \mid x>0 \} \\
T(x,x') &= \{ \langle x, x' \rangle \mid x' = (x - 1) \, \& \, x \}
\end{align*}
We will abbreviate this with the notation:
\begin{align*}
G(x) & \triangleq x > 0 \\
T(x, x') & \triangleq x' = (x - 1) \, \& \, x
\end{align*}

\begin{figure*}
\begin{framed}
\begin{definition}[Unconditional Termination Formula {\bf [UT]}]
\label{def:UT}
\begin{align*}
 \exists R . \forall x, x' . & G(x) \wedge T_o(x, x') \rightarrow R(x) > 0 \wedge R(x) > R(x')
\end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula -- Open Recurrence Set  {\bf [ONT]}]
\label{def:ont}
 \begin{align*}
  \exists N, x_0 . \forall x . \exists x' . & N(x_0) ~\wedge \\ &  N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \rightarrow T_u(x, x') \wedge N(x') 
 \end{align*}
\end{definition}

\begin{definition}[Non-Termination Formula -- Closed Recurrence Set {\bf [CNT]}]
\label{def:cnt}
 \begin{align*}
  \exists N, x_0 . \forall x, x' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T_u(x, x') \rightarrow N(x') 
 \end{align*}
\end{definition}

%% \begin{definition}[Non-deterministic Non-Termination Formula - Closed Recurrence Set {\bf [Nondet-CNT]}]
%% \label{def:nondet-cnt}
%%  \begin{align*}
%%   \exists N, x_0 . \forall x, x' . \exists y_0. & N(x_0) ~ \wedge ~ N(x) \rightarrow G(x) ~ \wedge \\
%% 							& N(x) \wedge T_u(x, x') \wedge A(x') \rightarrow N(x')~
%%  \end{align*}
%% \end{definition}

\end{framed}
\end{figure*}

\noindent {\bf Unconditional termination.}
To show that $L$ is unconditionally terminating (i.e. it eventually terminates regardless of the
state it starts in), we try to find a ranking function for it. When defining $L$'s termination 
formula, we need only consider the over-approximation of its transition relation $T$, which we call $T_o$.
Note that this distinction is only important if $L$'s body contains nested loops, for which we'll have to 
consider the over-approximation of their transition relation's transitive closure
as shown in Section~\ref{sec:env}. For any other statements, $T_o$ and $T$ are equivalent.

\begin{theorem}
\label{thm:ut}
 The loop $L(G, T_o)$ terminates from every start state iff formula {\bf [UT]} (Definition~\ref{def:UT}) is
 satisfiable.
\end{theorem}

\begin{proof}
 Then the codomain of $R$ is a well founded set and $R$ is an order homomorphism.

 (Do this proof properly).
\end{proof}

As the existence of a ranking function is equivalent to the satisfiability
of the formula {\bf [UT]}, a satisfiability witness is also a ranking
function and thus a proof of $L$'s unconditional termination.

Continuing the example of the program in Figure~\ref{fig:motivation.a}, we
can see that the corresponding synthesis formula {\bf [UT]} is satisfiable,
as witnessed by the function $R(x) = x$.  Thus, $R(x) = x$ constitutes a
proof that the program in Figure~\ref{fig:motivation.a} is unconditionally
terminating.

Note that different formulations for unconditional termination are possible. 
We are aware of a proof rule based on transition invariants, i.e.  supersets
of the transition relation's transitive
closure~\cite{DBLP:conf/pldi/GrebenshchikovLPR12}.  This formulation assumes
that the second-order logic has a primitive predicate for disjunctive
well-foundedness.  Conversely, our formulation in Definition~\ref{def:UT}
does not require a transitive closure operator, nor a disjunctive
well-foundedness predicate.  \\

\noindent{\bf Non-termination.}
Dually to termination, we might want to consider the non-termination of a program.  If a program terminates,
we can prove this by finding a ranking function %and supporting invariant 
witnessing the satisfiability of formula {\bf[UT]}.  What then would a proof of non-termination look like?

Gupta et al.~\cite{DBLP:conf/popl/GuptaHMRX08} characterize non-termination
of a transition relation $T$ by the existence of an \emph{(open) recurrence
set}, i.e.  a nonempty set of states $N$ such that for each $s \in N$ there
exists a transition to some $s'\in N$.
For our case, the notion of open recurrent set is encoded by formula {\bf [ONT]}. %Definition~\ref{def:nonterm-formula}.
Dual to the termination problem, when formulating $L$'s non-termination, we only need to consider an under-approximation of its transition
relation~$T_u$. 
Again, this distinction is mostly important when dealing with nested loops,
as illustrated in Section~\ref{sec:env}.

\begin{theorem}
\label{thm:ont}
 The loop $L(G, T_u)$ has an infinite execution iff Defintion~\ref{def:ont} is satisfiable.
\end{theorem}

\begin{proof}
 Then $x_0$ is an underapproximation to the reachable states on entry to $L$, and $N$ is an open recurrence set.

 (Do this proof properly).
\end{proof}

If this formula is satisfiable, $N$ is an open recurrence set for $L$, which proves
$L$'s non-termination. The issue with this formula is the additional level of quantifier alternation as compared to the synthesis fragment
(it is an $\exists \forall \exists$ formula). %This means that we will not be able to use the solver for the synthesis fragment to solve it.

In order to formulate the non-termination problem inside the synthesis
fragment, we make use of the notion of \emph{closed recurrence set}
introduced by Chen et al.~in~\cite{DBLP:conf/tacas/ChenCFNO14}.  As opposed
to an open recurrence set, this is an $\exists \forall$ property: for each
state in the recurrence set $N$, all of its successors must be in $N$.  We
note that if $T$ is deterministic, every open recurrence set is also a
closed recurrence set (since each state has at most one successor).  Thus,
the non-termination problem for deterministic transition systems is
equivalent to the satisfiability of formula {\bf [CNT]}.

\begin{theorem}
\label{thm:cnt}
 If $T_u$ is deterministic, i.e. each state $x$ for which the guard $G(x)$ holds has exactly one successor $x'$ such that $T_u(x, x')$, then
 Definition~\ref{def:ont} is equivalent to Definition~\ref{def:cnt}.
\end{theorem}

\begin{proof}
 If $T_u$ is deterministic, then each $x$ satisfying $G(x)$ has a unique successor $S(x)$, so $\forall x, x' . G(x) \wedge T_u(x, x') \leftrightarrow x' = S(x)$.

 (Finish this proof).
\end{proof}

Note that by defining $T_u$'s determinism as the existence of exactly one
successor for any state satisfying $L$'s guard, we eliminate the possibility
of $T_u$ being $\mathit{false}$, which would otherwise constitute a
counterexample for Theorem~\ref{thm:cnt}.

Given that our program's state space is finite, a transition relation
induces a infinite execution if and only if some state is visited infinitely
often, or equivalently the state is visited twice in the course of the
execution.  This gives us an alternative way of expression a non-termination
proof: we can can try to find a witness $x$ to the satisfiability of the
formula
%
$$ \exists x . T^+(x, x)$$
%
However, deciding satisfiability of this formula directly requires a logic
that includes a transitive closure operator, which ours does not have.

\noindent {\bf Non-determinism.} A non-termination diagnosis procedure based on checking the satisfiability of formula {\bf [CNT]}
may fail to find a closed recurrence set for a loop $L=(G, T_u)$ when $T_u$ is non-deterministic.  
To understand the issue, imagine the following loop with an non-deterministic assignment:

\begin{lstlisting}[language=C]
while(x != 0) {
  y = nondet();
  x = x-y;
}
\end{lstlisting}

Only for certain values of \texttt{y} the loop is non-terminating.
In order to restrict the values that may be assigned to \texttt{y}, we need an assumption associated with the non-deterministic choice. 
However, such an assumption may cause a reachable state at the non-deterministic assignment to not have any successor. In such a case
the termination formula would be satisfiable, although the execution halts. This situation can be avoided by checking the existence of 
at least one successor for each reachable state at the non-deterministic assignment. Although we did put a reasonable amount of effort into it, 
we failed to come up with a formulation for the non-termination of such non-deterministic transition relations 
inside the synthesis fragment.
%formula {\bf [Nondet-CNT]} in Definition~\ref{def:nondet-cnt}.

\begin{figure*}
 \begin{framed}

\begin{definition}[Conditional Termination Formula {\bf [CT]}]
\label{def:ct}
 \begin{align*}
  \exists R, W . \forall x, x' . & I(x) \wedge G(x) \rightarrow W(x) ~ \wedge \\
                                 & G(x) \wedge W(x) \wedge T_o(x, x') \rightarrow W(x') \wedge R(x) > 0 
  \wedge R(x) > R(x')
 \end{align*}
\end{definition}

%%  \begin{definition}[Sequential Loops Termination Formula {\bf [ST]}]
%%   \begin{align*}
%% \label{def:multi-termination-formula}
%%   \exists R, Inv_1,..., Inv_n . \forall x_0,...,x_n, x_1',...,x_n'.  & P_0(x_0,x_1) \rightarrow Inv_1(x_1) ~  \\
%%  & \bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) ~  \\  
%%  & \bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) \rightarrow Inv_{i+1}(x_{i+1})) ~ \wedge \\
%%  & Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge G_n(x_n) \rightarrow Inv_n(x_n) ~ \wedge \\
%%  & Inv_n(x_n) \wedge G_n(x_n) \wedge T(x_n, x_n') \rightarrow Inv_n(x_n') \wedge R(x_n) > 0 \wedge R(x_n) > R(x_n')
%%  \end{align*}
%% \end{definition}

%% \begin{definition}[Nested Loops Termination Formula {\bf [NT]}]
%% \label{def:nested-term-formula}
%%  \begin{align*}
%%   \exists R, S . \forall x, x' . & G_1(x) \wedge P_1(x,x') \rightarrow S(x',x') ~ \wedge \\
%%                                 & G_2(x') \wedge S(x,x') \wedge T(x',x'')\rightarrow S(x,x'') ~ \wedge \\
%% 				& G_1(x) \wedge P_1(x,x') \wedge \neg G_2(x') \wedge S(x',x'') \wedge 
%%                                 P_2(x'',x''') \wedge 
%%                                   R(x) > 0 \wedge R(x) > R(x''') 
%%  \end{align*} 
%% \end{definition}

%% \begin{definition}[Sequential Loops Non-Termination Formula {\bf [SNT]}]
%% \label{def:multi-termination-formula}
%%  \begin{align*}
%%   \exists R, Inv_1,..., Inv_n . \forall x_0,...,x_n, x_1',...,x_n'.  & P_0(x_0,x_1) \rightarrow Inv_1(x_1) ~  \\
%%  & \bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) ~  \\  
%%  & \bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) \rightarrow Inv_{i+1}(x_{i+1})) ~ \wedge \\
%%  & Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge N(x_n)  ~ \wedge \\
%%                                                         & N(x_n) \rightarrow G_n(x_n) ~ \wedge 
%% 							 N(x_n) \wedge T_n(x_n, x_n') \rightarrow N(x_n') 
%%  \end{align*}
%% \end{definition}
 \end{framed}

\end{figure*}

\subsection{Composing a Loop with its Environment} \label{sec:env}

Sometimes the termination behaviour of a loop depends on its environment, i.e. the rest of the program.  That is to say,
the loop may not terminate if started in some particular state, but the rest of the program
ensures that such a state cannot be reached on entry to the loop.  The program as a whole
terminates, but if the loop were considered in isolation we would not be able to prove that
it terminates. We must therefore encode a loop's interaction with its environment %the rest of the program 
in order to do a sound termination analysis.\\

\iffalse
Let us assume that we have done some preprocessing of our program which has identified
loops, straight line code blocks and the control flow between these.  In particular,
the control flow analysis has determined which order these code blocks execute in,
and the nesting structure of the loops.
\fi

\noindent {\bf Conditional termination.}
Given a loop $L=(G,T_o)$, where $T_o$ is an over-approximation of its transition relation, 
if $L$'s termination depends on the state it begins
executing in, we say that $L$ is \emph{conditionally terminating}.
The information we require of the environment is a predicate $I$ which
overapproximates the set of states that $L$ may begin executing in.
That is to say, for each state $x$ that is reachable on entry to $L$,
we have $I(x)$.

Conditional termination is then equivalent to the
formula {\bf [CT]} of Defintion~\ref{def:ct}.

\begin{theorem}
\label{thm:ct}
 The loop $L(G, T_o)$ terminates when started in any state satisfying $I(x)$ iff
 Defintion~\ref{def:ct} is satisfiable.
\end{theorem}

\begin{proof}
 Do this proof.
\end{proof}

If this formula is satisfiable, two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

$W$ is called a \emph{supporting invariant} for $L$ and $R$ proves termination relative to $W$.
We require that $I \wedge G$ is strong enough to establish the base case of $W$'s inductiveness.

Conditional termination is illustrated by the program in Figure~\ref{fig:motivation.h},
which is encoded as:
\begin{align*}
            I(\langle x, y \rangle) & \triangleq y = 1 \\
            G(\langle x, y \rangle) & \triangleq x > 0 \\
            T(\langle x, y \rangle, \langle x', y' \rangle) & \triangleq x' = x - y \wedge y' = y 
\end{align*}
If the initial states $I$ are ignored this loop cannot be shown to terminate, since any state with $y = 0$ and $x > 0$
would lead to a non-terminating execution.

However, formula {\bf [CT]} is satisfiable, as witnessed by the ranking function
$R(\langle x,y\rangle) = x$ and supporting invariant $W(\langle x, y \rangle ) \triangleq y = 1$.
This constitutes a proof that the program as a whole terminates, since the loop always begins
executing in a state that guarantees its termination.\\

%synthesizing environment abstractions: underapp for termination and underapp for non-termination.
% \cite{} addresses the problem of automatically underapproximating weakest preconditions.

%\noindent {\bf Nested loops.}
%% If $L$'s body contains another loop $L'$, we cannot directly apply either formula {\bf [UT]} or
%% {\bf [CT]} or prove that $L$ terminates.  This is because we don't have direct access to the
%% single step transition relation for $L$'s body.  Instead, we consider $L'$ to be part of
%% $L$'s environment, which we model with a relation $S$.  We require that $S$ overapproximates
%% $L'$ in the sense that it satisfies the following formula:
%% \begin{align*}
%%  \forall x, x'. G(x) \wedge W(x) \wedge T'^*(x, x') \wedge \lnot G'(x') \rightarrow S(x, x')
%% \end{align*}
%% Where $G$ and $G'$ are respectively $L$ and $L'$'s guards, and $T'^*$ is the reflexive transitive closure of
%% the inner loop ($L'$)'s transition relation.

%% To give an example, if we wish to prove that the outer loop in Figure~\ref{fig:nested} terminates,
%% it suffices to summarise the inner loop with the relation:
%% \begin{align*}
%% S(\langle x, y, z \rangle, \langle x', y', z' \rangle) & \triangleq x' = x \wedge y = 1
%% \end{align*}
%% Then the ranking function $R(x, y, z) = x$ proves termination of the outer loop $L(G, S)$.\\

%% \begin{figure}
%% \begin{lstlisting}
%% while (x > 0) {
%%   z = x, y = z + 1;
%%   while (z > 0) {
%%     z--, y--;
%%   }
%%   x -= y;
%% }
%% \end{lstlisting}
%% \caption{A nested loop\label{fig:nested}}
%% \end{figure}

\noindent {\bf Over-approximating the environment for termination.}
%% While proving that loop $L$ terminates, we create constraints on the environment in the
%% form of the predicate $I$ and the over-approximating summary relation $T_o$:
%% In the presence of nested loops, 
%% for each loop $L'$ nested in the body
%% of $L$, we must generate constraints saying that $T_o$ overapproximates $L'$, i.e. it is an over-approximation of the transitive closure of $L'$'s transition relation.
%% Similarly, for each loop $L''$ appearing before $L$ in the program's control flow, we
%% must add constraints saying that $I$ over-approximates the states reachable via $L''$.
%The details of how to create these constraints are standard, so
%rather than presenting the algorithm for generating the constraints, we direct the
%reader's attention to Figure~\ref{fig:environment-model}.  
%
%\noindent {\bf Environment abstractions.}
For a loop $L=(G,T)$, the formulae {\bf [CT]} %and {\bf [CNT]} 
from Definition~\ref{def:ct} 
%and Definition~\ref{def:cnt} 
is in terms of 
$T_o$ and $I$, which over-approximate the transition relation~$T$ and the set of states that $L$ may begin executing in, respectively. 
The reason for using over-approximations [instead of exact relations] %when formulating the (non-)termination of an isolated loop 
is so that we can use the same formulae regardless of the environment the loop is executing in, or 
the complexity of its body, e.g.~if it contains nested loops.

All we need to do in order to use {\bf [CT]} in any context %environment and regardless of the complexity of the loops's body 
is generate constraints over-approximating %abstracting 
the environment and the loop's body. %: for termination, we need over-approximating abstractions, whereas for non-termination 
As the generation of such constraints is standard and covered by several other works \cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08}, 
we will not provide the full algorithm, but rather illustrate it through the example in Figure~\ref{fig:environment-model}.
Our technique proves the termination of each loop separately and in program order, 
with the exception of nested loops, whose termination we encode in the same formula.
Following the same design, let us assume that we have already proven the termination of loop $L_1$, and we are now trying to prove that $L_2$ and $L_3$ terminate, which we encode
as the conjunction of constraints on the right side of the figure:
%% =======
%% %
%% As the generation of such constraints is standard and covered by several
%% other
%% works~\cite{DBLP:conf/pldi/GrebenshchikovLPR12,DBLP:conf/pldi/GulwaniSV08},
%% we will not provide the full algorithm, but rather illustrate it through the
%% example in Figure~\ref{fig:environment-model}.  Let us assume that we have
%% already proven the termination of loops $L_1$ and $L_3$, and we are now
%% trying to prove that $L_2$ terminates, which we encode as the conjunction of
%% constraints on the right-hand side of the figure:
%% %
%% >>>>>>> ee2c21bbc574ab36e7d6739c0e82c7a03d930789
\begin{itemize}
\item $W_1$ and $W_2$ are (over-approximating) inductive invariants for $L_1$ and $L_2$, respectively.
\item $S$ is a summary of $L_3$ that over-approximates the transitive closure of its transition relation.
\item If the formula is satisfiable, then $R_2$ and $R_3$ are ranking functions for $L_2$ and $L_3$, respectively.
\end{itemize}

Note that despite its intricacy, the constraint system is still a formula in the synthesis fragment.
Since the outer loop of this program does indeed terminate, the system is satisfiable, as witnessed by:
\begin{align*}
W_1(\langle v, x, y, z \rangle ) & \triangleq x \leq y + 1 \\
W_2(\langle v, x, y, z \rangle ) & \triangleq x = y + 1 \\
S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) & \triangleq x ' = x \wedge y' = y \wedge \\
& z' + v' = x' \wedge v' \leq y' \\
R_2(x) & = x\\
R_3 & = ..
\end{align*}

\begin{figure*}
\begin{framed}
\begin{minipage}{0.17\textwidth}
\begin{lstlisting}[mathescape=true]
x = 0;
$L_1:$
while(x <= y)
  x++;

$L_2:$
while(x > 0){
  z = x;
  v = 0;

$L_3:$
  while(v < y){
    z--;
    v++;
  }

  x -= z;
  y -= z;
}
\end{lstlisting}
\end{minipage}
\vline
\begin{minipage}{0.85\textwidth}
\begin{align*}
 \exists W_1, W_2, R_2, R_3, S . \forall v, x, y, z, v', x', y', z' . \\
   x = 0 & \rightarrow W_1(v, x, y, z) \, \wedge \\
   x \leq y \wedge W_1(v, x, y, z) & \rightarrow W_1(v, x+1, x, y, z) \, \wedge \\
   x > 0 \wedge x > y \wedge W_1(v, x, y, z) & \rightarrow W_2(v, x, y, z) \, \wedge \\
   S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) \wedge v' < y' & \rightarrow S(\langle v, x, y, z \rangle, \langle v'+1, x', y', z'-1 \rangle) \, \wedge \\
   & R_3(v,x,y,z) > 0 \, \wedge \\
   & R_3(v,x,y,z) > R_3(v'+1,x',y',z'-1) \, \wedge\\
   x > 0 \wedge W_2(v, x, y, z) & \rightarrow S(\langle v, x, y, z \rangle, \langle 0, x, y, x \rangle) \, \wedge \\
   x > 0 \wedge W_2(v, x, y, z) \wedge S(\langle v, x, y, z \rangle, \langle v', x', y', z' \rangle) \wedge v' \geq y' & \rightarrow \\
   & W_2(v', x' - z', y' - z', z') \, \wedge \\
   & R_2(v, x, y, z) > 0 \, \wedge \\
   & R_2(v, x, y, z) > R_2(v', x' - z', y'-z', z')
\end{align*}
\end{minipage}
\end{framed}
\caption{A non-trivial program and its termination formula\label{fig:environment-model}}
\end{figure*}

\noindent {\bf Under-approximating the environment for non-termination.}
Dual to termination, when proving non-termination, we need to
under-approximate the environment and the loop's body and apply
formula {\bf [CNT]}.
Recall that there is one more constraint that we must obey when doing so: we
must stay inside the synthesis fragment!  This becomes more challenging now
then when dealing with over-approximations due to the existential nature of
under-approximations: they require an additional existential quantifier,
which might result in an additional quantifier alternation, i.e.~$\exists
\forall \exists$.

This issue does not arise when abstracting the environment, as its
under-approximation and the loop under consideration make use of existential
quantification at the same nestedness level.

However, when dealing with nested loops, we need to under-approximate the
inner loop, requiring a nested existential quantifier, resulting in the
$\exists \forall \exists$ alternation.  At first sight, this seems fatal for
our formulation.  Luckily, it is possible to reformulate non-termination as
an $\exists \forall$ formula by observing that, unlike a ranking function,
the defining property of a recurrence set is \emph{non relational} -- if we
end up in the recurrence set, we do not care exactly where we came from as
long as we know that it was also somewhere in the recurrence set.  This
allows us to cast non-termination of nested loops as the formula shown in
Figure~\ref{fig:nonterm-nested}.

If the formula on the right-hand side of the figure is satisfiable, then
$L_1$ is non-terminating, as witnessed by the recurrence set $N_1$ and the
initial state $x_0$ in which the program begins executing.  There are two
possible scenarios for $L_2$'s termination:
%
\begin{itemize}
%
\item If $L_2$ is terminating, then $N_2$ is an inductive invariant that
reestablished $N_1$ after $L_2$ stops executing: $\lnot G_2(x) \wedge N_2(x)
\wedge P_2(x,x') \rightarrow N_1(x') $.
%
\item If $L_2$ is non-terminating, then $N_2 \wedge G_2$ is its recurrence set.
%
\end{itemize}

\begin{figure*}
\begin{framed}
 \begin{minipage}{0.19\textwidth}
\begin{lstlisting}[mathescape=true]
$L_1$:
while ($G_1$) {
  $P_1$;

$L_2$:
  while ($G_2$) {
    $B_2$;
  }

  $P_2$;
}
\end{lstlisting}
\end{minipage}
\vline
\begin{minipage}{0.82\textwidth}
\begin{align*}
 \exists N_1, N_2, x_0 . \forall x, x' . \\
  N_1(x_0) & \, \wedge \\
  N_1(x) & \rightarrow G_1(x) \, \wedge \\
  N_1(x) \wedge P_1(x,x') & \rightarrow N_2(x') \, \wedge \\
  G_2(x) \wedge N_2(x) \wedge B_2(x,x') & \rightarrow N_2(x') \, \wedge \\
  \lnot G_2(x) \wedge N_2(x) \wedge P_2(x,x') & \rightarrow N_1(x') 
\end{align*}
\end{minipage}
\end{framed}

\caption{The non-termination formula for nested loops \label{fig:nonterm-nested}}
\end{figure*}

\iffalse

\noindent {\bf Sequential loops.}
Conditional termination allows us to
prove that programs with multiple loops terminate, even if the termination of some loop depends on the states
reachable after leaving a previous loop.
%The conditional termination formula becomes more complex in the presence of multiple loops.
For illustration, consider the following program with $n$ sequential loops $\forall i=1..n. ~L_i=(x_i,G_i,T_i)$ and
the assertions $\forall i=1..n. ~P_i$ modelling the program statements in between loops.

\begin{lstlisting}[mathescape=true]
$P_1$
while ($G_1$) { $T_1$ }
$P_2$

while ($G_2$) { $T_2$ }
...
$P_n$
while ($G_n$) { $T_n$ }
\end{lstlisting}

In order to prove that the $n^{th}$ loop terminates, we need to apply the {\bf [CT]} formula for $L_n$.
The challenge is determining the initial states  $I(x_n)$ at the loop's entry point.  
The termination problem for $L_n$ is thus equivalent to the satisfiability of
Definition~\ref{def:multi-termination-formula}.  If this formula is satisfiable:
\begin{itemize}
\item $Inv_i$ is an inductive invariant of $L_i$ for $i=1..{n-1}$ as established by the conjuncts:
$$\bigwedge_{i=1..n{-}1} (Inv_i(x_i) \wedge G_i(x_i) \wedge T_i(x_i, x_i') \rightarrow Inv_i(x_i')) $$

\item Each loop invariant $Inv_i$ with $i=1..{n-2}$ is strong enough to establish $Inv_{i+1}$:
$$\bigwedge_{i=1..n{-}2} (Inv_i(x_i) \wedge \lnot G_i(x_i) \wedge P_{i+1}(x_i, x_{i+1}) {\rightarrow} Inv_{i+1}(x_{i+1})) ~$$

\item When $L_n$'s guard holds, $Inv_{n-1}$ 
is strong enough to establish $Inv_n$, i.e. $Inv_n$ holds for the program states after $L_n$'s entry point:
$$ Inv_{n-1}(x_{n-1}) \wedge \lnot G_{n-1}(x_{n-1}) \wedge P_n(x_{n-1},x_n) \wedge G_n(x_n) $$
$$\qquad\qquad\rightarrow Inv_n(x_n)$$

\item $Inv_n$ is a supporting termination invariant of $L_n$ and $R$ is a ranking function relative to $Inv_n$:
$$Inv_n(x_n) \wedge G_n(x_n) \wedge T(x_n, x_n') \rightarrow Inv_n(x_n')$$
$$ \qquad \qquad \wedge R(x_n) > 0 \wedge R(x_n) > R(x_n')$$
\end{itemize}

\todo{add example?}\\

\noindent{\bf Nested loops.}
%
We illustrate our approach to handling nested loops for only two loops for
sake of brevity, but the approach can be easily generalised to any number of
nested loops:
%
\begin{lstlisting}[mathescape=true]
while ($G_1$) { 
  $P_1$ 
  while ($G_2$) { $T$ }
  $P_2$
}
\end{lstlisting}

The termination of the outer loop is equivalent to the formula in
Definition~\ref{def:nested-term-formula}.  The formula requires an auxiliary
assertion $S(x,x')$ representing a binary relation between the initial
states at the inner loop's entry point and their successors according to the
transition relation $T$.  If this formula is satisfiable, then $R$ is a
ranking function of the outer loop.

In order to find a termination argument for the inner loop, we use {\bf [CT]} where the initial states are given by 
$\neg G_1(x) \wedge P_1(x,x') \rightarrow I(x')$.

\noindent{\bf Sequential loops.}
Discuss {\bf [SNT]}.\\

\noindent{\bf Nested loops.}
\fi

\subsection{Generalised Termination and Non-termination Formula} 
Given a program, our algorithm attempts to prove termination or non-termination for each loop individually 
by constructing a generalised formula for each of them.
For nested loops, we generate one single formula for all the nested loops (as illustrated for termination in Figure~\ref{fig:environment-model}).

For an isolated loop $L=(I,G,T)$, the generalised termination and non-termination formula is the following:
%
 \begin{align*}
  \exists R, N, x_0 . \forall x, x' . & N(x_0) ~ \wedge \\ & N(x) \rightarrow G(x) ~ \wedge \\
							& N(x) \wedge T_u(x, x') \rightarrow N(x')\\
  & \vee G(x) \wedge T_o(x, x') \rightarrow  R(x){>}0 \wedge R(x){>}R(x')
 \end{align*}
%
%
where $T_u$ and $T_o$ under- and over-approximate the transition relation $T$, respectively.

%If this formula is satisfiable, then:

%% Our design: for each loop always search for lexicographic ranking function + supporting invariant OR recurrence set.

\section{Satisfiability of a Formula in the Synthesis Fragment} \label{sec:synthesis}

In this section we present a solver for the synthesis fragment:
 \[
  \exists P,~ x . ~\forall~ y . \sigma(P, x, y)
 \]
%
where $P$ ranges over sets, while $x$ and $y$ range over ground terms (as
mentioned before, we drop the vector notation and write $x$ for $\vec{x}$).

By viewing $\sigma: (X^n \times Y^m \to Z^k) \times X^n \times Y^m  \to
\mathbb{B}$ as a specification function that returns true iff $P$ computes
appropriate outputs for the inputs $x$ and $y$, we can regard the
satisfaction problem for the synthesis fragment as \emph{program synthesis},
i.e.  the mechanised construction of software that provably satisfies a
given specification.  Research in the area of program synthesis has been
particularly fruitful beginning with Alonzo Church's work on the
\emph{Circuit Synthesis Problem} in the sixties~\cite{church-synth}, and
continuing with works such as {\sc Brahma}~\cite{brahma} and program
sketching~\cite{lezama-thesis,sketch,modular-sketch}.

In the context of (non-)termination analysis, a witnesses to the
satisfiability of a formula in the synthesis fragment is a
\mbox{(non-)termination} proof.  As mentioned earlier in the paper, our
analysis must ensure precise treatment of bit-wise operations and arithmetic
modulo fixed widths by modelling bit-accurate semantics.  The synthesiser
must therefore be able to compute (non-termination) proofs over machine
integers and floating-point numbers from the specifications detailed in
Section~\ref{sec:second.order}.

\subsection{The Synthesis Specification}

The logic we use to express the synthesis formula, i.e.~the\\
\mbox{(non-)termination} specifications, is a subset of C that we call
\newC.  The characteristic property of a \newC program is that safety can be
decided for it using a single query to a Bounded Model Checker.  A~\newC
program is just a C program with the following syntactic restrictions:
 all loops in the program must have a constant bound;
 all recursion in the program must be limited to a constant depth;
 all arrays must be statically allocated (i.e. not using \texttt{malloc}),
 and be of constant size.
Additionally, \newC programs may use nondeterministic values, assumptions
and arbitrary-width types.

The (non-)termination proofs are arbitrary expressions over the grammar given in Figure~\ref{fig:l-language}.  
%written in a simple RISC-like language $\mathcal{L}$, whose syntax is given in Fig.~\ref{fig:l-language}.  
\todo{update the figure and make it look more like a grammar.}

\begin{algorithm*}
 \caption{Abstract refinement algorithm
 \label{alg:cegis}}
 \vspace{-1.5em}
 \begin{multicols}{2}
 \begin{algorithmic}[1]
 \Statex
\Function{synth}{inputs}
  \Let{$(i_1, \ldots, i_N)$}{inputs}
  \Let{query}{$\exists P . \sigma(i_1, P(i_1)) \land \ldots \land \sigma(i_N, P(i_N))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{unsatisfiable}
  \EndIf
\EndFunction
\Statex
\Function{verif}{P}
  \Let{query}{$\exists x . \lnot \sigma(x, P(x))$}
  \Let{result}{decide(query)}
  \If{result.satisfiable}
    \State \Return{result.model}
  \Else
    \State \Return{valid}
  \EndIf
\EndFunction
\columnbreak
\Statex
\Function{refinement loop}{}
  \Let{inputs}{$\emptyset$}
  \Loop
    \Let{candidate}{\Call{synth}{inputs}}
    \If{candidate = UNSAT}
      \State \Return{unsatisfiable}
    \EndIf
    \Let{res}{\Call{verif}{candidate}}
    \If{res = valid}
      \State \Return{candidate}
    \Else
      \Let{inputs}{inputs $\cup$ res}
    \EndIf
  \EndLoop
\EndFunction
 \end{algorithmic}
 \end{multicols}
\end{algorithm*}

\begin{figure}
 \begin{center}
 \resizebox{\linewidth}{!}{%
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,
 semithick, initial text=]

  \matrix[nodes={draw, fill=none, shape=rectangle, minimum height=1cm, minimum width=1.5cm},
          row sep=2cm, column sep=2.5cm, ampersand replacement=\&] {
   \node (synth) {Synthesise};
   \&
   \node (verif) {Verify}; %\\
   %\node[draw=none] {};
   \&
   \node[ellipse] (done) {Done}; \\
  };

   \path
    (synth) edge [bend left] node {Candidate program} (verif)
    (verif) edge [bend left] node {Counterexample input} (synth)
    (verif) edge node {Valid} (done);
 \end{tikzpicture}
 }
 \end{center}

 \caption{Abstract synthesis refinement loop
 \label{fig:abstract-refinement}}
\end{figure}

\begin{figure}
{\small
\begin{center}
\setlength{\tabcolsep}{16pt}
Integer arithmetic instructions:

\begin{tabular}{llll}
 \verb|add a b| & \verb|sub a b| & \verb|mul a b| & \verb|div a b| \\
 \verb|neg a| &   \verb|mod a b| & \verb|min a b| & \verb|max a b|
\end{tabular}

\medskip

Bitwise logical and shift instructions:

\begin{tabular}{lll}
 \verb|and  a b| & \verb|or   a b| & \verb|xor a b| \\
 \verb|lshr a b| & \verb|ashr a b| & \verb|not a|
\end{tabular}

\medskip

Unsigned and signed comparison instructions:

\begin{tabular}{lll}
 \verb|le  a b| & \verb|lt  a b| & \verb|sle  a b| \\
 \verb|slt a b| & \verb|eq  a b| & \verb|neq  a b|
\end{tabular}

Miscellaneous logical instructions:

\begin{tabular}{lll}
 \verb|implies a b| & \verb|ite a b c| & 
\end{tabular}

Floating-point arithmetic:

\begin{tabular}{llll}
 \verb|fadd a b| & \verb|fsub a b| & \verb|fmul a b| & \verb|fdiv a b| 
\end{tabular}

\end{center}
}
 \caption{The language $\mathcal{L}$}
 \label{fig:l-language}
\end{figure}

\subsection{The Synthesis Algorithm}
As safety of a \newC program is efficiently 
%the quantifier free first-order fragment of the logic is efficiently 
decidable and a termination proof is expressible as a ground term of the \newC logic, we use
Counterexample Guided Inductive Synthesis (CEGIS)~\cite{lezama-thesis,sketch} to
check the validity of the synthesis formula. 
The core of the CEGIS algorithm is the refinement loop shown in Fig.~\ref{fig:abstract-refinement} and
detailed in Algorithm~\ref{alg:cegis}.  
The algorithm is divided into two
procedures: {\sc synth} (see Figure~\ref{fig:synth-dfd}) and {\sc verif}, which interact via
a finite set of test vectors {\sc inputs}, which is initialised to $\emptyset$.

The {\sc synth} procedure tries to find existential witnesses $(P, x_0)$
that satisfy the partial specification:
%
\[
 \exists P, x_0 . \forall x \in \text{\sc inputs} . \sigma(P, x_0, x)
\]

If {\sc synth} succeeds in finding a witness $(P, x_0)$, this witness is a
candidate solution to the full synthesis formula.  We pass this candidate
solution to {\sc verif} which determines whether the candidate does satisfy
the specification on all inputs by checking satisfiability of the
verification formula:
%
\[
 \exists x . \lnot \sigma(P, x_0, x)
\]

If this formula is unsatisfiable, the candidate solution is in fact a
solution to the synthesis formula and so the algorithm terminates. 
Otherwise, the witness $x$ is an input on which the candidate solution fails
to meet the specification.  This witness $x$ is added to the {\sc inputs}
set and the loop iterates again.

Concretely, {\sc synth} is implemented as shown in
Figure~\ref{fig:synth-dfd}.  We start by taking the termination
specification and generating a C program which takes as input a candidate
$(P, x_0)$ and asserts that the candidate fails to meet the specification on
at least on of the elements of {\sc inputs}.  So this program is unsafe iff
there is some candidate $(P, x_0)$ satisfying the specification for all the
elements of {\sc inputs}.  Finding a new candidate solution is therefore
reduced to model checking this C program, which by construction is loop
free.  The model checking is done using a combination of bounded model
checking, explicit state enumeration and genetic programming (GP).  The
latter two techniques involve combining the previously generated C program,
along with code that searches for candidate solutions.  Finally, if any of
the model checkers finds a candidate solution it returns it.  The form of
this candidate solution is yet another program, this time written in the
proof language $\mathcal{L}$.

As well as efficiency, compiling the original source code of the program
under analysis has the advantage of guaranteeing that we are verifying the
exact semantics of the program as understood by a compiler, rather than some
ad-hoc abstraction of its semantics.

Similarly to {\sc synth}, the {\sc verif} procedure is implemented by
creating a C program that is unsafe iff some input exists on which the
candidate solution fails to meet the specification.  Again, symbolic and
explicit state model checking are used to decide the satisfiability of the
verification formula.

\begin{figure*}
\begin{center}
\tikzstyle{file}=[draw, text width=7.0em, text centered,
  minimum height=1.5em]
\tikzstyle{process} = [draw, minimum height=3em, circle]
\tikzstyle{line} = [draw, color=black, -latex']

\def\Divide#1#2{%
 \coordinate(a) at ($(#1.east) !.5! (#2.west)$);
 \coordinate(b) at (a |- 0,-3);
 \draw[dotted] (b) -- ++(0, 4.5);
}

%\resizebox{\linewidth}{!}{
\begin{tikzpicture}[font=\sffamily]

\node [file] (pua) {input.c};

\path (pua.east)+(2,0) node [file] (spec) {\sc Termination Specification};
\path (spec.south)+(0.0, -1) node [file] (inputs) {\sc Tracked Inputs};
%% \path (synth.south)+(0.0, -0.5) node [file] (tests) {\sc tests.c};
%% \path (tests.south)+(0.0, -0.5) node [file] (interpreter) {\sc interpreter.c};
%% \path (interpreter.south)+(0.0, -0.5) node [file] (spec) {\sc spec.c};

%% \path (spec.east)+(2.0, -0.25) node [process] (merged) {merge};
\path (spec.east)+(2.0, -0.75) node [file] (file) {synth.c};

\path (file.east)+(2.0, 1.5) node [process] (cbmc) {\sc cbmc};
\path (file.east)+(2.0, 0.0) node [process] (gcc) {\sc Search};
\path (file.east)+(2.0, -1.5) node [process] (gp) {\sc GP};

\path (gcc.east)+(2.5, 0.0) node [file] (out) {Candidate Proof};

%\path [line] (spec) -- (merged);

\path [line] (pua) -- (spec);

\path [line] (spec) -- (file);
\path [line] (inputs) -- (file);

\path [line] (file) -- (cbmc);
\path [line] (file) -- (gcc);
\path [line] (file) -- (gp);

\path [line] (cbmc) -- (out);
\path [line] (gcc) -- (out);
\path [line] (gp) -- (out);

\Divide{pua}{spec}{puaspec}
%\Divide{spec}{file}{specfile}
\Divide{file}{gcc}{filegcc}
\Divide{gcc}{out}{gccout}

\draw (pua |- 0,-3) node [align=center] {Input program \\ (written in C)};

\coordinate (midspec) at ($(spec) !.5! (file)$);
\draw (midspec |- 0,-3) node [align=center] {Automatically generated specification \\ (written in C)};
%\draw (file |- 0,-3) node {Spec};
\draw (gcc |- 0,-3) node [align=center] {Model checker};
\draw (out |- 0,-3) node [align=center] {Proof \\ (written in $\mathcal{L}$)};

\end{tikzpicture}
%}
\end{center}

\caption{Schematic diagram of {\sc synth}}
\label{fig:synth-dfd}
\end{figure*}

\subsection{Candidate Generation Strategies}
\noindent{\bf Explicit Proof Search} The simplest strategy for finding candidates
is to just exhaustively enumerate them all, starting with the shortest and
progressively increasing the number of instructions.  Since the set of
$\mathcal{L}$-programs is recursively enumerable, this procedure is complete.
\\

\noindent{\bf Symbolic Bounded Model Checking} Another complete method for generating
candidates is to simply use BMC on the {\sc synth.c} program.  As with explicit
search, we must progressively increase the length of the solution we search for
in order to get a complete search procedure.
\\

\noindent {\bf Genetic Programming and Incremental Evolution} \label{sec:gp}
To generate candidate (non-)termination proofs in the ``Synthesise'' phase
of the refinement loop, we make use of genetic programming
(GP)~\cite{langdon:fogp} as an alternative to model checking and exhaustive
search.  Genetic programming provides an intelligent and adaptive way of
searching through the space of computer programs for an individual that is
highly fit in solving the problem at hand by using the Darwinian principle
of survival and reproduction of the fittest.  Thus, the program that emerges
from the genetic programming paradigm is a consequence of fitness.

We begin with a population of random programs, then iteratively evolve this
population by applying the genetic operators {\sc crossover} and {\sc
mutate}.  The {\sc crossover} operator takes two programs and combines their
code in some way to create a new program.  The {\sc mutate} operator takes a
single program and randomly changes parts of its code, again creating a new
program.  Both these operators are applied to programs selected from the
population in proportion to their observed fitness, and a program is said to
be fitter than another if it meets the specification on more test cases.

In the setting of program synthesis, genetic programming can be very
usefully combined with the concept of incremental evolution.  The idea is
that in order to evolve a program that is adequate for some difficult task,
it is often helpful to first evolve a population that as an ensemble is good
at some simpler variation of the task.  This principle leads us to the
observation that each round of the CEGIS loop can be thought of as creating
a more difficult problem for the GP synthesiser to solve.  Therefore, if GP
finds a candidate program which is correct for the $k$ inputs so far
observed, but is not correct for the full specification, that candidate
program came from a population that was ``good'' for the specification
restrict to those $k$ inputs.  The next time round the CEGIS loop, the
problem GP must solve is slightly harder -- any solution to this new problem
must be correct for the previous $k$ inputs and also whatever new input was
added by the previous iteration of {\sc verif}.  It therefore makes sense to
begin the $k+1$st round of evolution with the population at the end of the
$k$-th round.

\section{Soundness, Completeness and Complexity}

Termination for bitvector programs is known to be
PSPACE-complete~\cite{DBLP:conf/tacas/CookKRW10}.  We first record the
expected results for soundness and completeness of our algorithm for
deciding it.
%
\begin{theorem}\label{thm:synth-sound}
Algorithm~\ref{alg:cegis} is sound -- if it terminates with witness $P$, then
$P \models \sigma$.
\end{theorem}

\begin{proof}
 The procedure {\sc RefinementLoop} terminates only if {\sc Verif} returns ``valid''.  In that
 case, $\exists x . \lnot \sigma(x, P)$ is unsatisfiable and so $\forall x . \sigma(x, P)$ holds.
\end{proof}

\begin{theorem}
 \label{thm:synth-semicomplete}
 If the existential first-order theory used to express the specification $\sigma$ is decidable,
 algorithm~\ref{alg:cegis} is semi-complete -- if a program $P \models \sigma$
 exists then algorithm~\ref{alg:cegis} will terminate.  However, if no program
 satisfies the specification, the algorithm may not terminate.  Furthermore if the domain of inputs $X$
 is finite, the algorithm is guaranteed to terminate.
\end{theorem}

\begin{proof}
 Since the {\sc ExplicitSearch} routine enumerates all programs (as can be seen by induction on
 the program length $l$), it will eventually enumerate a program that meets the specification
 in whatever set of inputs are currently being tracked.  Since the first-order theory is
 decidable, the query in {\sc Verif} will succeed for this program, causing the algorithm to terminate.
 The set of correct programs is therefore recursively enumerable and algorithm~\ref{alg:cegis}
 enumerates this set, so it is semi-complete.

 If the domain $X$ is finite then the loop in procedure {\sc RefinementLoop} can only
 iterate $\| X \|$ times, since by this time all of the elements of $X$ would have been
 added to the inputs set.  Therefore if the {\sc Synth} procedure always terminates (which it does
 by assumption of decidability), the whole algorithm terminates.
\end{proof}

\begin{theorem}
\label{thm:sound}
 {\sc Headshot} is sound -- if it returns a verdict of termination, the program under
 analysis terminates on all inputs.  Similarly if it returns a verdict of non-termination,
 the program under analysis has at least one infinite execution.
\end{theorem}

\begin{proof}
 This is a direct consequence of Theorems~\ref{thm:ut}, \ref{thm:ct}, \ref{thm:ont}, \ref{thm:cnt} and \ref{thm:synth-sound}.
\end{proof}

\begin{theorem}
\label{thm:complete-termination}
 {\sc Headshot} is complete for terminating programs -- if a program terminates, {\sc Headshot} will find a proof that it does.
\end{theorem}

\begin{proof}
 If the program under analysis $P$ has variables $v_1, \ldots, v_k$ each of which is $b$ bits wide, its state space $\mathcal{S}$ is of size $2^{bk}$.
 A ranking function $R: \mathcal{S} \to \mathcal{D}$ for $P$ exists iff $P$ terminates.  Since $R$ is injective, we have that
 $\| \mathcal{D} \| \geq \| \mathcal{S} \|$.  If $\| \mathcal{D} \| > \| \mathcal{S} \|$, we can construct a function $R': \mathcal{S} \to \mathcal{D'}$
 with $ \| \mathcal{D'} = \| \mathcal{S} \|$ by just setting $R' = R|_\mathcal{S}$, i.e. $R'$ is just the restriction of $R$ to $\mathcal{S}$
 Since $\mathcal{S}$ already comes equipped with a natural well ordering we can also construct $R'' = \iota \circ R'$
 where $\iota: \mathcal{D'} \to \mathcal{S}$ is the unique order isomorphism from $\mathcal{D'}$ to $\mathcal{S}$.
 So assuming that $P$ terminates, there is some ranking function $R''$ that is just a permutation of $\mathcal{S}$.
 If the number of variables $k > 1$ then in general the ranking function will be lexicographic with dimension $\leq k$
 and each co-ordinate of the output being a single $b$-bit value.

 It is fairly easy to see that any finite permutation is computed by a finite program in the proof language generated
 by {\sc Kalashnikov}.  A very inefficient construction which computes the permutation $R(x) = x+1$ for one variable
 is:
 \begin{verbatim}
t1 = 0
t2 = v1 == 0
t3 = ITE(t2, 1, t1)
t4 = v1 == 1
t5 = ITE(t4, 2, t3)
...
 \end{verbatim}
%
It is easy to see how this generalises to computing an arbitrary permutation.
Thus, every permutation is computed a program of size at most $b \times
2^{bk + 1}$.  Since {\sc Kalashnikov} enumerates all programs in order of
length, it will eventually check a program computing each permutation of
$\mathcal{S}$ and so will eventually check a program computing the ranking
function $R''$.  Checking that a single program satisfies the specification
is a decidable problem, and the search as a whole terminates.
%
\end{proof}

\begin{theorem}
 {\sc Headshot} is complete for deterministic programs -- {\sc Headshot} will terminate when run on
 any program with a deterministic transition relation.
\end{theorem}

\begin{proof}
 If the program under analysis is terminating, then by Theorem~\ref{thm:complete-termination} {\sc Headshot}
 will terminate having computed a ranking function.

 Conversely if the program is non-terminating, {\sc Headshot} will terminate
 having computed a recurrence set.  Since, by assumption, the program under analysis is determinstic,
 Definition~\ref{def:ont} is equivalent to Definition~\ref{def:cnt}.
 That is to say, each open recurrence set is also a closed recurrence set.  Since the specification for a predicate
 to encode a closed recurrence set is in the synthesis fragment, {\sc Kalashnikov} will find a program computing
 such a predicate if one exists.  The proof is similar to the proof of Theorem~\ref{thm:complete-termination}.
\end{proof}

We will now show that our time complexity is asymptotically a function of
the Kolmogorov complexity of the (non-)termination proof, and argue
that this gives our procedure various desirable qualities in practical
appliations.
%
\begin{theorem}
 {\sc Kalashnikov} has time complexity $NP^{NP}(K(\sigma))$.
\end{theorem}

\begin{theorem}
 {\sc Headshot} uses a single call to {\sc Kalashnikov} and uses an encoding with $O(1)$
 overhead.  It therefore has time complexity $NP^{NP}(K(\tau(P))$ where $\tau(P)$ is the
 termination specification for program $P$.
\end{theorem}

In Section~\ref{sec:synthesis} we will describe a procedure for solving the synthesis formula.
This procedure has the property that it will find the shortest program meeting the
specification, and as a consequence the runtime of the program synthesiser is dominated
by the size of this shortest program.  To frame the discussion of the runtime of our
termination proving procedure, we first recall the definition of the Kolmogorov complexity
of a function $f$:

\begin{definition}[Kolmogorov complexity]
 The Kolmogorov complexity $K(f)$ is the length of the shortest program that
 computes~$f$.
\end{definition}

We can extend this definition slightly to talk about the Kolmogorov complexity of a
synthesis problem in terms of its specification:

\begin{definition}[Kolmogorov complexity of a synthesis problem]
 The Kolmogorov complexity of a program specification $K(\sigma)$ is the length of the shortest
 program $P$ such that $P \models \sigma$.
\end{definition}

We will later show that the time complexity of our synthesis procedure is $NP^{NP}(K(\sigma))$,
which means that for even moderately large $K(\sigma)$ our procedure is intractably slow.  Conversely
if $K(\sigma)$ is small we will be able to find a termination proof, regardless of other parameters
such as the size of the program under analysis.  To put this another way, under the assumption
that $K(\sigma)$ is small for a program's termination problem, we will find a ranking function
quite rapidly.  We will now investigate the properties of functions with low Kolmogorov complexity,
and show that the assumption of a low-Kolmogorov ranking function is much weaker than the assumption
of a linear ranking function.

\begin{theorem}
 Linear functions have low Kolmogorov complexity.
\end{theorem}

\begin{proof}
 The function $f: X \to Y$ can be computed with a program consisting of
 $2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$ instructions (1 multiplication and 1 addition for
 each cell in the matrix representing $f$).  Therefore,
 $K(f) \leq 2 \cdot \dim(X) \cdot \dim(Y) - \dim(Y)$.
\end{proof}

\begin{theorem}
Assuming a reasonable program encoding, the probability that a random program of length $l$ computes
a linear function is $O(2^{-l})$.
\end{theorem}

\begin{proof}
 (Do this proof properly).

 Otherwise, we'd need to have an exponential number of programs computing the same function.
 That would be an unreasonable program encoding.
\end{proof}

Therefore the number of non-linear functions $f$ with $K(f) \leq \kappa$
grows exponentially with $\kappa$, while the number of linear functions
grows only linearly.  This shows that assuming the existence of a linear
ranking function is much stronger than assuming the existence of a short
termination proof.  In addition to being a much weaker assumption,
we argue that low Kolmogorov complexity is a more natural assumption than
linearity.  Humans tend to write programs they can understand.  It's hard
to quantify exactly what ``understandable'' means, but we think that
``having a short proof'' is closer to the mark than ``having a linear proof''.

\todo{do something with this?}
Since each loop is bounded by a constant, and each recursive function call is
limited to a constant depth, a \newC program necessarily terminates and in
fact does so in $O(1)$ time.  If we call the largest loop bound~$k$, then
a Bounded Model Checker with an unrolling bound of $k$ will be a complete
decision procedure for the safety of the program.  For a \newC program of
size $l$ and with largest loop bound~$k$, a Bounded Model Checker will
create a SAT problem of size $O(lk)$.  Conversely, a SAT problem
of size $s$ can be converted trivially into a loop-free \newC program
of size $O(s)$.  The safety problem for \newC is therefore NP-complete,
which means it can be decided fairly efficiently for many practical
instances.

\section{Experiments}

We implemented our algorithm in the {\sc Headshot} tool, which uses the {\sc
Kalashnikov} program synthesiser for termination.  We ran the tool on 40
benchmarks taken from the termination category of SVCOMP'15.  The
experiments were performed on a 4-core 3.3\,GHz i5-2500k with 8\,GB of RAM.

\begin{figure*}
\centering
\footnotesize
\begin{tabular}{|l|l||l|r||l|r|r|}
\hline
    &             & \multicolumn{2}{|c||}{\sc ARMC} & \multicolumn{3}{|c|}{\sc Headshot} \\
    & Expected &  Verdict & Time & Verdict & Time & Iterations\\
    \hline
    \hline
\input{table}
    \hline
\end{tabular}

Key: \tick = terminating, \xmark = non-terminating.

\caption{\textsc{Headshot} termination for nonlinear programs with nonlinear ranking functions\label{fig:nonlinear}}
 \end{figure*}

 \section{Conclusions and Related Work}
%\subsection{Termination analysis}
Automated program termination is a research topic that has received a fair amount of attention from the software verification community.
In order to compare our technique to the rest of the area, 
Figure~\ref{fig:handletable} summarises the related works with respect to the restrictions they impose on the transition relations as well as the form of the computed ranking functions.

Somehow expected, most of the techniques are specialised in the synthesis of linear ranking functions for linear programs over integers (or rationals) \cite{DBLP:conf/pldi/CookPR06,DBLP:conf/cav/LeeWY12,DBLP:conf/popl/Ben-AmramG13,DBLP:conf/vmcai/P04,DBLP:conf/atva/HeizmannHLP13,DBLP:conf/cav/BradleyMS05,DBLP:conf/tacas/CookSZ13}. 
Among them, 
Lee et al. make use of transition predicate abstraction, algorithmic learning, and decision procedures to compute transition
invariants as proofs for the termination of linear programs \cite{DBLP:conf/cav/LeeWY12}.
Leike and Heizmann present a new method for the constraint-based synthesis
of termination arguments for linear loop programs based on
linear ranking templates \cite{DBLP:conf/tacas/LeikeH14}.
Linear ranking functions supported by inductive linear invariants for loops with linear guards and transitions: \cite{DBLP:conf/cav/BradleyMS05}. 
Learning: \cite{DBLP:conf/cav/HeizmannHP14}

While the synthesis of termination arguments for linear programs over
integers is indeed well-covered in the literature, there is very limited
work for programs over machine integers.  Cook et al.  present a method
based on a reduction to Presburger arithmetic, and a template-matching
approach for predefined classes of ranking functions based on reduction to
SAT- and QBF-solving~\cite{DBLP:conf/tacas/CookKRW10}.  Similarly, the only
work we are aware of that can compute non-linear ranking functions for
imperative loops with polynomial guards and polynomial assignments
is~\cite{DBLP:conf/vmcai/BradleyMS05}.  However, this work extends only to
polynomials.

Given the lack of research in handling \emph{not-linear programs}, as well
as \emph{programs over bit vectors and floats}, our work focuses on covering
these areas.  One of the obvious conclusions that can be reached from
observing Figure~\ref{fig:handletable}, is that most of the works tend to
specialise on a certain aspect of termination proving that they can solve
efficiently.  Conversely to this view, we aim for \emph{generality}, as we
do not restrict the form of the synthesised ranking functions, nor the form
of the input programs.

As mentioned in Section~\ref{sec:intro}, approaches based on Ramsey's
theorem compute a set of local termination conditions that decrease as
execution follows through the loop and require expensive reachability
analyses~\cite{DBLP:conf/lpe/CodishG03,DBLP:conf/lics/PodelskiR04,DBLP:conf/pldi/CookPR06}. 
In an attempt to reduce the complexity of checking the validity of the
termination argument, Cook et al present an iterative termination proving
procedure that searches for lexicographic termination
arguments~\cite{DBLP:conf/tacas/CookSZ13}, whereas Kroening et al. 
strengthen the termination argument such that it becomes a transitive
relation~\cite{DBLP:conf/cav/KroeningSTW10}.

Proving program termination implies the simultaneous search for a
termination argument and a supporting invariant.  For this purpose, several
tools alternate between calls to a safety prover and calls to a ranking
function synthesis tool \cite{}.  In order to share more information about
the state of the termination proofs between the two tools, Brockschmidt et
al.  use a single representation of the state of the termination proof
search that both tools operate over, resulting in performance improvements
\cite{DBLP:conf/cav/BrockschmidtCF13}.  In \cite{DBLP:conf/cav/BradleyMS05},
Bradley et al.  combine the generation of ranking functions with the
generation of invariants to form one constraint solving problem such that
the necessary supporting invariants for the ranking function are discovered
on demand.  In our setting, both the ranking function and the supporting
invariant are iteratively constructed through the same refinement loop.

While program termination has been extensively studied, much less research has been conducted in the area of proving nontermination.
Gupta et al. dynamically enumerate lasso-shaped candidate
paths for counterexamples, and then statically prove their feasibility \cite{DBLP:conf/popl/GuptaHMRX08}. 
In \cite{DBLP:conf/tacas/ChenCFNO14}, Chen et al. prove nontermination via
reduction to safety proving. Their iterative algorithm uses counterexamples to a fixed safety property
to refine an underapproximation of a program.

\bibliographystyle{abbrvnat}
\bibliography{synth}{}

\end{document}