### 2014-07-06 papers/termination/table.tex

loop1.c & \tick & \tick & 0.06s & \tick & 1.3s & 1\\

loop2.c & \tick & \tick & 0.06s & \tick & 1.4s & 4\\

loop3.c & \tick & \tick & 0.06s & \tick & 1.8s & 8\\

loop4.c & \tick & \tick & 0.12s & \tick & 2.9s & 18\\

loop5.c & \tick & \tick & 0.12s & \tick & 5.3s & 18\\

loop6.c & \tick & \tick & 0.05s & \tick & 1.2s & 1\\

loop7.c~\cite{DBLP:conf/cav/BrockschmidtCF13} & \tick & ? & 0.05s & \tick & 8.3s & 17\\

loop8.c & \tick & ? & 0.06s & \tick & 1.3s & 1\\

loop9.c & \tick & \tick & 0.11s & \tick & 1.6s & 1\\

loop10.c & \tick & \xmark & 0.05s & \tick & 1.3s & 1\\

loop43.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & \tick & 0.07s & \tick & 1.5s & 1\\

loop44.c~\cite{DBLP:conf/tacas/CookSZ13} & \xmark & ? & 0.05s & \xmark & 10.5s & 9\\

loop45.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & \tick & 0.12s & \tick & 4.3s & 18\\

loop46.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & ? & 0.05s & \tick & 1.5s & 2\\

loop47.c & \tick & \tick & 0.10s & \tick & 1.8s & 1\\

loop48.c & \tick & \tick & 0.06s & \tick & 1.4s & 4\\

loop49.c & \xmark & ? & 0.05s & -- & T/O & 206\\

svcomp1.c~\cite{DBLP:conf/flops/Avery06} & \tick & \tick & 0.11s & \tick & 2.3s & 2\\

svcomp2.c & \tick & \tick & 0.05s & \tick & 1.5s & 9\\

svcomp3.c~\cite{BA:mcs} & \tick & \tick & 0.15s & \tick & 146.4s & 43\\

svcomp4.c~\cite{DBLP:conf/cav/BradleyMS05} & \xmark & \xmark & 0.09s & \xmark & 2.1s & 7\\

svcomp5.c~\cite{DBLP:conf/icalp/BradleyMS05} & ? & \tick & 0.38s & -- & T/O & 46\\

svcomp6.c~\cite{DBLP:conf/cav/BrockschmidtCF13} & \tick & -- & T/O & \tick & 29.1s & 28\\

svcomp7.c~\cite{DBLP:conf/cav/BrockschmidtCF13} & \tick & \tick & 0.09s & \tick & 5.5s & 17\\

svcomp8.c~\cite{Chen:2012:TPL:2414936.2414966} & \tick & ? & 0.05s & -- & T/O & 37\\

svcomp9.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & \tick & 0.10s & \tick & 1.5s & 8\\

svcomp10.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & \tick & 0.11s & \tick & 4.5s & 18\\

svcomp11.c~\cite{DBLP:conf/tacas/CookSZ13} & \tick & \tick & 0.20s & \tick & 14.6s & 33\\

svcomp12.c~\cite{DBLP:journals/aaecc/DershowitzLSS01} & \tick & -- & T/O & \tick & 10.9s & 24\\

svcomp13.c & \tick & ? & 0.07s & \tick & 35.1s & 62\\

svcomp14.c~\cite{DBLP:conf/pldi/GulwaniJK09} & \tick & -- & T/O & \tick & 30.8s & 26\\

svcomp15.c~\cite{DBLP:conf/sas/HarrisLNR10} & \tick & ? & 0.12s & -- & T/O & 191\\

svcomp16.c~\cite{DBLP:conf/sas/HarrisLNR10} & \tick & \tick & 0.06s & \tick & 2.2s & 3\\

svcomp17.c~\cite{DBLP:conf/cav/KroeningSTW10} & \tick & \tick & 0.05s & -- & T/O & 32\\

svcomp18.c~\cite{DBLP:conf/fmcad/LarrazORR13} & \tick & ? & 0.27s & -- & T/O & 26\\

svcomp25.c & ? & ? & 0.05s & -- & T/O & 26\\

svcomp26.c & \tick & \tick & 0.26s & \tick & 3.2s & 23\\

svcomp27.c~\cite{DBLP:conf/sigsoft/Nori013} & \xmark & \tick & 0.11s & -- & T/O & 67\\

svcomp28.c~\cite{DBLP:conf/sigsoft/Nori013} & \tick & \tick & 0.13s & -- & T/O & 79\\

svcomp29.c~\cite{DBLP:conf/vmcai/P04} & \tick & ? & 0.05s & -- & T/O & 1\\

svcomp37.c & \tick & \tick & 0.16s & \tick & 2.1s & 10\\

svcomp38.c & \tick & \tick & 0.10s & -- & T/O & 1\\

svcomp39.c & \tick & \tick & 0.25s & -- & T/O & 1\\

svcomp40.c~\cite{DBLP:conf/sas/Urban13} & \tick & ? & 0.07s & -- & T/O & 30\\

svcomp41.c~\cite{DBLP:conf/sas/Urban13} & \tick & ? & 0.07s & -- & T/O & 27\\

svcomp42.c & \tick & \tick & 0.22s & -- & T/O & 85\\

\hline  
\hline 
\multicolumn{2}{|l||}{Correct} & \multicolumn{2}{|r||}{27} & \multicolumn{3}{|r|}{31} \\
\multicolumn{2}{|l||}{Incorrect} & \multicolumn{2}{|r||}{2} & \multicolumn{3}{|r|}{0} \\
\multicolumn{2}{|l||}{Unknown} & \multicolumn{2}{|r||}{13} & \multicolumn{3}{|r|}{0} \\
\multicolumn{2}{|l||}{Timeout} & \multicolumn{2}{|r||}{3} & \multicolumn{3}{|r|}{15} \\