### 2014-09-10 papers/termination/rebuttal.txt
edits

Reviewers 2, 3 and 4 (lack of novelty over invariant generation and other
second-order techniques): Previous works on using second-order logic for
program analysis use completely different approaches for encoding
termination: [19] requires a dwf predicate in the logic to model
termination, i.e., it uses a logic with a builtin halting oracle.  This
makes the encoding trivial.  [20] reduces (non-)termination to a safety
property and uses templates to reduce the resulting second-order formula to
a first-order one.  The (non-)termination proofs take the form of
preconditions and, in contrast to our work, there is no guarantee of
completeness or even soundness.  We maintain that our encoding of
(non-)termination as a second-order fomula is novel and non-trivial.  For
example characterising termination of nested loops is not a straightforward
extension of invariant generation: it requires relational reasoning (which
is not the same class of problem as invariant generation -- an inherently
non-relational problem); our use of Skolemisation to handle nondeterminism
in non-termination arguments is novel and non-trivial; generating a
tautology for which we find a constructive proof is novel and non-trivial.

Reviewers 1, 3 and 4 (experimental performance): ARMC is very good at
handling loops with a small number of paths that have linear, rational
termination arguments.  This is a PTIME problem.  Consequently, on harder
examples ARMC performs poorly while Juggernaut excels, for example in the 5
benchmarks with nested loops, ARMC solves none while Juggernaut solves 4
correctly.  Since ARMC is solving a PTIME problem, it is not surprising that
it is faster than our more general algorithm in those cases that are
instances of the PTIME problem.  We also note that ARMC is not sound with
respect to bitvector semantics: our experiments include 3 cases (6%) where
ARMC terminates with an incorrect verdict.  Juggernaut does not give an
incorrect verdict for any benchmark.  The paper did perhaps not make it
clear that we are comparing the performance of a sound and complete
algorithm (Juggernaut) with an unsound, incomplete algorithm (ARMC); this is
the best comparison we could do, since there are no existing sound and
complete algorithms to compare with.

Reviewers 2, 3 and 4 (simplicity of our approach): our algorithm is the
first to solve many non-trivial termination problems.  It is also the first
to decide termination of bit-vector and floating-point programs, and the
first to combine both termination and non-termination analysis in one
algorithm.  We consider it the main strength of our work that we can use a
sequence of simple reductions to transform a previously intractable problem
into a simple one.

Reviewers 3 and 4 (comparison to SyGus): our specifications are C code, but
SyGus specifications are S-Expressions only.  This design decision enables
us to use explicit-state model checking and GP, as well benefitting from
compiler optimisations.  This difference is non-trivial and novel.  However
the underlying problem is clearly very similar and we appreciate it being
brought to our attention so that we may cite it.

Reviewer 4 (importance of bitvector semantics): we completely agree that
people really don't *think* about termination using bitvector arguments. 
However, they subsequently write a bitvector program, and the semantic gap
is a frequent root-cause of serious bugs.  Equally, a tool that generates
rational termination arguments and checks whether it applies for bitvector
semantics has limited utility: programmers do like termination arguments
based on integers, and will be disappointed that an analyser that uses
rationals fails to find it. The only advantage of rational termination
arguments is that they are cheap.

Reviewer 3 (novelty compared to other synthesis algorithms): Our
contribution is twofold: we first present a novel reduction from termination
to synthesis, then we present an algorithm and tool for solving the
synthesis formula.  The first part is significant and theoretically
non-trivial (see previous comments).  The second part is a significant
technical contribution in that it manages to synthesise C programs from a C
specification, and is able to combine SAT solving, explicit-state model
checking and stochastic search.  No other program synthesiser does this.

Reviewer 1: C- is distinct from L in that C- is the implementation language
of part of our synthesiser, and also the language used for specifications. 
It really does not leave out any other C features.  L is the language in
which ranking functions are written and is distinct from C-.  While section
5.1 may seem excessive, we feel that many in the POPL community are unlikely
to realise that bitvector termination has no simple relationship with
rational termination: there are programs that terminate with bitvector
semantics but don't with rationls *and vice versa*.  The ITE instruction is
"if-then-else"; we agree this figure is somewhat opaque and will add some
discussion.  The runtime is predominantly a factor of the Kolmogorov
complexity, but the problem is in fact NEXPTIME-complete in the Kolmogorov
complexity, so even modest increases in solution size may result in large
jumps in runtime.  We very much appreciate the suggestions regarding the
rest of the presentation and intend to incorporate them into the paper.

Reviewer 2 (selection of benchmarks): We omitted exactly those SVCOMP'15
benchmarks that made use of arrays or recursion.  We do not have arrays in
our logic and we had not implemented recursion in our frontend (although the
latter can be syntactically rewritten to our input format).